### 2015-09-13 doc/reports/sandia_final_impact.tex
adding reports directory in docs and adding work on final sandia fellowship report for amsi work

- Description of next steps
\label{future-work}
Currently AMSI supports only static parallel execution space discretization with non-overlapping process sets. The restriction on non-overlapping process sets is currently being worked on to allow temporary scale-reasignment while a scale is blocked waiting on an associate scale and has no work which can be conducted. However, more dynamic management of the execution space repressents the most significant current goal of the AMSI project. Implementation of this feature -- which will allow scale-tasks to grow and shrink as computational demands associated with the scale change -- is dependent on much if not all of the preceeding functionality (especially load balancing and migration features, as all relevant scale data from a process to be dynamically reassigned scales must be recovered and redistributed throughout the related scale).

Scale-balancing is the balancing of scale-tasks against eachother to minimize rendezvous time and maximize time spent in useful computation. Scale-balancing is a difficult prospect as it involves managing many dynamically-changing scales which themselves are being load-balanced using scale-sensitive load-balancing techinques, and represents a sigificant and promising area of future work. Once the ability to dynamically manage the process set associated with a scale is developed, algorithms must be developed to make the best use of this feature. Developing these algorithms repressents a major - and daunting - area of future work, as there are numerous constraints which must be taken into account in order to optimize such a scheme.

A second micro-scale simulation, where the collagen fibers are embedded in tissue is currently implemented as a standalone simulation. Work to incorporate this into the biotissue example simulation will be conducted when the correct scale-coupling terms have been developed (inclusion in the AMSI system is already accomplished, but the up- and down-scaling operations have not been developed yet so this RVE is not used). \cite{} \cite{} \cite{}

An article focusing on the implementation and use of the scale-sensitive load-balancing operations in AMSI for the biotissue example problem is being finalized and will be seeking publication soon.

- Summary of new proposals derived from work
- Follow-on funding

At present there are no current proposals associated with this work for ongoing funding, simply because funding for associated projects (Biotissue and other multi-scale projects) has already been secured. Thus work on AMSI will continue during the work on projects supported by AMSI. Several interesting multi-scale simulations have been suggested, however, particularly a combination of a Monte-Carlo code and a Mangetostatics code by a fellow researcher. At present these are in the preliminary stage of development.

- Impact on programs

### 2015-09-13 doc/reports/sandia_final_introduction.tex
adding reports directory in docs and adding work on final sandia fellowship report for amsi work

Most physical models and implementations using numerical methods operate at a single scale characteristic to the problem -- the scale at which the preponderance of attributes governing the underlying physics being modeled are measured -- and the scale at which the resolution of the solution is considered comensurate with the desired accuracy of the simulation.

Introduction of influence from scales orders of magnitude removed from the primary scale can take place in several different ways. This influence can be incorporated during the mathematical development of the physical model itself. In this case the multi-scale features are incorporated into the numerical implementation automatically, as a result of their presence in the mathematical model. Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales but opera ting on some of the same physical fields and values.

Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms from some scales and supplying them to others, typically this will involve up-/down-scaling equations to take into account the difference in scale. This is only useful in cases where terms from the various incorporated scales contribute meanignfully to the solution at the primary scale of interest. The domain at the scale of interest (for brevity the engineering scale) is sufficiently large that conducting the entire simulation with a granularity orders of magnitude smaller would make the simulation computationally intractable.

This restriction requires the development of multi-scale simulation paradigms: algorithms and operations which allow the influence of various far-removed scales to be accurately reflected in the simulation of the engineering scale. While there are several well-established multi-scale paradigms (particularly the heterogenous multi-scale method as well as the concurrent multi-scale method), this is a rapidly advancing field and as new problems are considered additional multi-scale models will undoubtedly be developed. As such developing a set of tools to optimize a single multi-scale paradigm will be less impactful then the development of general tools which can be used themselves to implement new multi-scale paradigms, while supporting extant paradigms as efficiently as possible.

### 2015-09-13 doc/reports/sandia_final_results.tex
adding reports directory in docs and adding work on final sandia fellowship report for amsi work

A multiscale simulation for the analysis of the mechanical behaviors of soft biological tissues was developed and implemented using AMSI. The biotissue simulation is currently a two-scale hierarchical multi-scale simulation coupling two continuum analyses. The primary scale (engineering scale) is a macro-scopic simulation using the finite element method on the cauchy momentum balance equation. The tertiary scale is micro-scopic and takes place at every macro-scale integration point (those points where the stress response of the material is queried during the macro-scale simulation) and used to supply the material response. The micro-scale simulation simply conducts a force-balance operation on a randomly generated collagen fiber-network residing inside of a unit cube representative volume element (RVE) which is deformed using the state of the associated macro-scale element, and has appropriate boundary conditions enforced.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{biotissue-scales.png}
  \end{center}
  \caption{\small Biotissue multi-scale domains}
  \label{biotissue_scales}
\end{figure}

The biotissue simulation was implemented by combining two single-scale simulations: a typical parallelized finite element analysis and a quasistatics code. The combination of these two simulations using AMSI was very straightforward, involving the introduction of an initialization and configuration phase for AMSI, wrapping the pre-existing main functions of each simulation in an AMSI scale-task callback, and finally introducing relavant AMSI communcation (and load-balancing) calls at locations where scale-interaction is required.

The simulation has been run many times on the Amos BlueGene/Q managed by the Center For Computational Innovations at Rennselaer Polytechnic Institute, and has scaled up to 250k elements at macro-scale with over 16k cores (with >90\% assigned to micro-scale simulations as these represent the bulk of actual computation). Scaling beyond this should be possible but has not currently been conducted. Theoretically the simulation should be able to scale up to around 300 million elements, limited by memory restrictions on the AMOS machine and the memory associated with a single RVE. Scaling to the full machine is however necessary to get a richer understanding of the performance of AMSI.

Symposia on the development and use of AMSI for the biotissue problem have been given at SIAM Parallel Processing for Scientific Computing 2014 and at SIAM Computational Science and Engineering 2015.

The Parallel Processing for Scientific Computing talk \cite{} was a discussion of the general layout and services provided by AMSI and the implementation of the biotissue example problem as well as initial performance results of the biotissue problem on a relatively modest (20k elements) mesh and a more reasonable mesh (200k elements).

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{allTimes_2macro_20k.png}
  \end{center}
  \caption{\small Multi-scale solver 1st iteration (20k elements)}
  \label{20k_first_iteration}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{dogBone_weakScaling.png}
  \end{center}
  \caption{\small Weak scaling study}
  \label{weak_scaling}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{siam_weak_scaling.png}
  \end{center}
  \caption{\small Strong scaling on 20k mesh}
  \label{strong_scaling}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{dogBone14k.png}
  \end{center}
  \caption{\small Engineering scale mesh}
  \label{dogbone_mesh}
\end{figure}

The Computational Science and Engineering talk \cite{} was a discussion of the scale-sensitive load-balancing operations provided by AMSI and their usage in the Biotissue example problem. For this talk only the 20k mesh was used as the run times for a 200k mesh proved too long for a reliable turnaround on results, however the longer the computation (in terms of incremental load steps applied) the more benefit should be seen from the scale-sensitive load balancing. Longer runs on a larger (250k element) mesh are currently being conducted, as recent work on performance and memory footprint has allowed much quicker overall execuation times making larger runs feasible.

As each RVE is a nonlinear simulation their is no a-prior estimate of the computational load each RVE represents, so the optimal initial distribution is simply treating the weight of each RVE as 1. Thus each micro-scale process is assigned either $\floor*{\frac{\text{num_rves}}{\text{num\_micro\_procs}}}$ or $\ceil*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$. Once the first load-balancing event is reached, regardless of the granularity at which load-balancing is occuring for a given case, each RVE has recorded the total number of newton iterations it has conducted since the previous load-balancing event. This number is used as a heuristic weighting metric to determine the computational load of each individual RVE. The iteration count metric was chosen as each iteration takes a similar level of work to complete, up to the difference in the number of RVEs for each unique fiber network. Further, assuming linear incremental loading, regions in the engineering-scale mesh located near stress concentrators will result in 'more interesting' boundary conditions for the micro-scale problems related to numerical integration points in that region, which may result in the RVE requiring more Newton iterations to converge to an adequate solution, since the boundary conditions may lie farther from the initial state of the RVE in the regime of convergence for the nonlinear problem.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{siam_cse_initial.png}
  \end{center}
  \caption{\small Initial load distribution (20k elements)}
  \label{initial_load}
\end{figure}

The initial distribution of weights is shown in figure \ref{initial_load}, with a 20k element macro-scale mesh and 256 processes assigned to compute micro-scale RVEs. The left graph shows the overall weight (the sum of all newton iterations for all RVEs assigned to a given process) while the right graph shows the deviation from the average. While the average deviation is under 10\% is all cases, the max deviation is what truly matters since all micro-scale RVEs must finish processesing before scale-coupling communication can take place and the macro-scale can continue to compute the elemental matrices and assemble a global tangent stiffness matrix.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{siam_cse_updated.png}
  \end{center}
  \caption{\small Load balancing (20k elements)}
  \label{load_balancing}
\end{figure}

Figure \ref{load_balancing} shows the result after scale-sensitive load-balancing (using the Zoltan libraries to plan the actual migration) is conducted. The maximal deviation after the first iteration is unfortunately much higher. But after that first iteration the max deviation is reduced by approximately 80\%. This is illustrative of the trend seen in the biotissue simulation, the number of micro-scale newton iterations conducted during the first incremental load step are not illustrative of the number of micro-scale newton iterations required during additional incremental load steps, which stay largely the same. Thus the first load balancing operation 'over-corrects' the issue, but the second and subsequent load-balancing operations result in better load balancing and run times overall. This can be seen more clearly in figure \ref{10_step_lb} which shows the weights for 64 micro-scale processes with the same mesh as above over 9 incrememental loading steps, the left graph shows the load without balancing and the right shows the load with balancing. The max deviation without load balancing stays in the 20\% range, while with load balancing it typically stays under 10\%, as low as 2.77\% in the 8th loading step.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{siam_cse_ten.png}
  \end{center}
  \caption{\small Ten incremental loading steps (20k elements)}
  \label{10_step_lb}
\end{figure}

### 2015-09-14 doc/reports/sandia_final_impact.tex
adding figures and updating document

@@
 
-An article focusing on the implementation and use of the scale-sensitive load-balancing operations in AMSI for the biotissue example problem is being finalized and will be seeking publication soon.
+At present AMSI has no meta-model of the computational domain of a problem, or of the relation between the domains of associated scales, simply that they are related.
+A richer model of the domain relation between problems would likely allow for a wider range of automation in scale-linking and dynamic scale management operations.
+This will likely not be addressed until a test case where such functionality is needed or beneficial is found.
+
+An article focusing on the implementation and use of the scale-sensitive load-balancing operations in AMSI for the biotissue example problem is being finalized and will be submited for publication soon.
+Additional articles focused primarily on the usage of the Biotissue example simulation to solve novel problems of interest is also being worked on and should see submission in the coming months.
 
@@
 Thus work on AMSI will continue during the work on projects supported by AMSI.
-Several interesting multi-scale simulations have been suggested, however, particularly a combination of a Monte-Carlo code and a Mangetostatics code by a fellow researcher.
-At present these are in the preliminary stage of development.
+Several interesting multi-scale simulations have been suggested, however, particularly a combination of a Monte-Carlo code and a Magnetostatics code by a fellow researcher.
+At present this idea is in a preliminary stage of development, though using AMSI would be relatively straightforward and achieving a working code (given two working single-scale codes, and the development of physically-sound scale-linking operations) would likely take only several days of programming and debugging.
 
 - Impact on programs
+
+There are several ongoing projects at the Scientific Computation Research Center (SCOREC) on the RPI campus which AMSI will be used with.
+Particularly SCOREC's ongoing work on incorporating our dynamic mesh operations with the Albany simulation system developed at Sandia National Labs.
+For multi-scale simulations using the Albany simulation code, we will make use of AMSI's capabilities to be minimally intrusive and only require modifications at those locations in code where multi-scale terms are needed.

### 2015-09-14 doc/reports/sandia_final/sandia_final_impact.tex
moving sandia report into its own directory, a few document updates and adding the bibtex file

--- 2015-09-14 doc/reports/sandia_final_impact.tex
@@
-- Description of next steps
 \label{future-work}
@@
 Work to incorporate this into the biotissue example simulation will be conducted when the correct scale-coupling terms have been developed (inclusion in the AMSI system is already accomplished, but the up- and down-scaling operations have not been developed yet so this RVE is not used).
-\cite{} \cite{} \cite{}
+\cite{lake2012mechanics} \cite{zhang2013cross} \cite{zhang2013coupled}
 
@@
 
-- Summary of new proposals derived from work
-- Follow-on funding
-
+\label{new-propsals-funding}
 At present there are no current proposals associated with this work for ongoing funding, simply because funding for associated projects (Biotissue and other multi-scale projects) has already been secured.
@@
 
-- Impact on programs
-
+\label{impact-on-programs}
 There are several ongoing projects at the Scientific Computation Research Center (SCOREC) on the RPI campus which AMSI will be used with.

### 2015-09-14 doc/reports/sandia_final/sandia_final_results.tex
moving sandia report into its own directory, a few document updates and adding the bibtex file

--- 2015-09-13 doc/reports/sandia_final_results.tex
@@
 The micro-scale simulation simply conducts a force-balance operation on a randomly generated collagen fiber-network residing inside of a unit cube representative volume element (RVE) which is deformed using the state of the associated macro-scale element, and has appropriate boundary conditions enforced.
+More detail into the derivation of this multi-scale system can be found in \cite{stylianopoulos2008thesis} \cite{agoram2001coupled} \cite{stylianopoulos2007multiscale} .
+Discussion of the dimensionalization of the dimensionless force terms (the upscaling operation from micro-scale to macro-scale) produced by each fiber network can be found in \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.
 
@@
 
-Symposia on the development and use of AMSI for the biotissue problem have been given at SIAM Parallel Processing for Scientific Computing 2014 and at SIAM Computational Science and Engineering 2015.
+Symposia on the development and use of AMSI for the biotissue problem have been given at the SIAM Conference on Parallel Processing for Scientific Computing 2014 and at the SIAM Conference on Computational Science and Engineering 2015.
 
-The Parallel Processing for Scientific Computing talk \cite{} was a discussion of the general layout and services provided by AMSI and the implementation of the biotissue example problem as well as initial performance results of the biotissue problem on a relatively modest (20k elements) mesh and a more reasonable mesh (200k elements).
+The Parallel Processing for Scientific Computing talk \cite{wtobin2014pp} was a discussion of the general layout and services provided by AMSI and the implementation of the biotissue example problem as well as initial performance results of the biotissue problem on a relatively modest (20k elements) mesh and a more reasonable mesh (200k elements).
 
@@
   \begin{center}
-    \includegraphics[height=2in]{dogBone_weakScaling.png}
-  \end{center}
-  \caption{\small Weak scaling study}
-  \label{weak_scaling}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[height=2in]{siam_weak_scaling.png}
+    \includegraphics[height=2in]{siam_pp_weak_scaling.png}
   \end{center}
@@
 
-The Computational Science and Engineering talk \cite{} was a discussion of the scale-sensitive load-balancing operations provided by AMSI and their usage in the Biotissue example problem.
+The Computational Science and Engineering talk \cite{wtobin2015cse} was a discussion of the scale-sensitive load-balancing operations provided by AMSI and their usage in the Biotissue example problem.
 For this talk only the 20k mesh was used as the run times for a 200k mesh proved too long for a reliable turnaround on results, however the longer the computation (in terms of incremental load steps applied) the more benefit should be seen from the scale-sensitive load balancing.
@@
 As each RVE is a nonlinear simulation their is no a-prior estimate of the computational load each RVE represents, so the optimal initial distribution is simply treating the weight of each RVE as 1.
-Thus each micro-scale process is assigned either $\floor*{\frac{\text{num_rves}}{\text{num\_micro\_procs}}}$ or $\ceil*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$.
+Thus each micro-scale process is assigned either $ \floor*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}} $ or $ \ceil*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}} $.
 Once the first load-balancing event is reached, regardless of the granularity at which load-balancing is occuring for a given case, each RVE has recorded the total number of newton iterations it has conducted since the previous load-balancing event.
@@
 The left graph shows the overall weight (the sum of all newton iterations for all RVEs assigned to a given process) while the right graph shows the deviation from the average.
-While the average deviation is under 10\% is all cases, the max deviation is what truly matters since all micro-scale RVEs must finish processesing before scale-coupling communication can take place and the macro-scale can continue to compute the elemental matrices and assemble a global tangent stiffness matrix.
+While the average deviation is under 10\% in all cases, the max deviation is what truly matters since all micro-scale RVEs must finish processesing before scale-coupling communication can take place and the macro-scale can continue to compute the elemental matrices and assemble a global tangent stiffness matrix.
 
@@
 
-Figure \ref{load_balancing} shows the result after scale-sensitive load-balancing (using the Zoltan libraries to plan the actual migration) is conducted.
+Figure \ref{load_balancing} shows the result after scale-sensitive load-balancing (using the Zoltan libraries \cite{ZoltanOverviewArticle2002} \cite{ZoltanIsorropiaOverview2012} to plan the actual migration) is conducted.
 The maximal deviation after the first iteration is unfortunately much higher.
@@
   \begin{center}
-    \includegraphics[height=2in]{siam_cse_ten.png}
+    \includegraphics[height=2.5in]{siam_cse_ten.png}
   \end{center}
-  \caption{\small Ten incremental loading steps (20k elements)}
+  \caption{\small Nine incremental loading steps (20k elements)}
   \label{10_step_lb}

### 2015-09-14 doc/reports/sandia_final/sandia_final_abstract.tex

The Adaptive Multi-scale Simulation Infrastructure (AMSI) is a set of libraries and tools developed to support the development, implementation, and execution of general multimodel simulations. Using a minimal set of simulation meta-data AMSI allows for minimally intrusive work to adapt existant single-scale simulations for use in multi-scale simulations. Support for dynamic runtime operations such as single- and multi-scale adaptive properties is a key focus of AMSI. Particular focus has been spent on the development on scale-sensitive load balancing operations to allow single-scale simulations incorporated into a multi-scale simulation using AMSI to use standard load-balancing operations without effecting the integrity of the overall multi-scale simulation.

### 2015-09-14 doc/reports/sandia_final/sandia_final_methods.tex

Currently AMSI provides interfaces for describing and managing the execution of individual scales in a multi-scale simulation, for defining and managing the scale-coupling communication required by such simulations, and for planning and enacting scale-sensitive load balancing operations on individual scales in a multi-scale system.

AMSI operates by maintaining a set of minimal simulation metadata in order to model various quantities of interest during simulation execution. A decentralized approach toward control is taken; dynamic control decisions are made and implemented at runtime during collective operations. In order to avoid the introduction of unnecessary parallel barriers into the code, AMSI control decisions are only made during operations which are already collective over the set of processes effected by a given control decision.

\label{amsi-scales}
During simulation initialization simulation scales are declared and then defined by associated them with process sets, and declaring their scale-linking relations. Process sets are simply mathematical sets of process ranks implemented so as to take advantage of any mathematical conveniences to minimize explicit storage whenever possible. At present process sets must be non-overlapping to take advantage of the simulation control portions of the AMSI libraries, but for the simulation metadata modeling, scale-coupling communication, and multi-scale load balancing operations, the process sets associated with related scales may overlap. Scales are declared to be related if some quantities of interest (such as tensor field values) are transmitted between them for scale-coupling.

\label{amsi-communication}
Using data structures describing the dynamic state of the parallel execution space the AMSI can determine which scales will need to interact, and only for these interacting scales are the required internal buffers and scale-coupling structures allocated. This approach is taken in order to reduce unneeded overhead caused by maintaining data structures specific to interacting scale-tasks, which would be the case when assuming that any two scale-tasks may interact.

Scale-linking quantities are dynamically registered with the AMSI system by the scale-task which produces the data, along with information about which interacting scale-tasks will receive how much data. This data is stored in AMSI, and when scale-linking communication is to take place, a user-configurable algorithm operating on a mapping of all pertinent scale-linking data is used to provide a full plan of scale-linking communication. This plan is the diseminated (only partially) to those processes which need to know their individual role in the communication. At this point scale-linking communication of any POD datatypes may take place at any time. If the quantity of scale-linking data changes at any point the previous steps must be repeated, but if the quantity remains the same for several instances of inter-scale communication the processes do not need to occur every time. Quantity here is a fairly abstract term, as the discrete units modeled by AMSI metadata structures is arbitrary: they can be any discrete unit of data that can be handled by the underlying communications system (MPI in most cases).

\label{amsi-load-balancing}
Multiscale load balancing processes also make use of the meta-data structures concerning the discretization of the parallel execution space. AMSI currently has separate functionalities for: data removal, data addition with load balancing, and scale-sensitive load-balancing (migration). The removal routine does not automatically call any load balancing or migration routines and should normally be paired with the migration routine. In the current hierarchical mutliscale test problem these routines are called in the previously listed order. Data is first removed which may unbalance the work loads. Next, data is added with load balancing so that differences in workload can be filled by the new data. Finally, scale-senstive load-balancing is used to rectify any remaining unbalance, though as the computational demand (weight) associated with the newly distributed data is typically not known a-priori, some semi-accurate weighting metric should be used to ensure the new data is not treated incorrectly. Currently each of these routines also assumes implicit ordering of data within AMSI since explicit ordering has not yet been implemented. This means that when using the load balancing and migration routines, the user's understanding of the data outside of AMSI needs to be reordered to maintain agreement across scales. This is a side-effect from the use of implicit (user-defined) ordering of scale-coupling data, which has the benefit of not introducing additional memory requirements to explicitly and uniquely number these pieces of data inside of AMSI. Further, AMSI provides the capability to automatically reorder the user data and provide a report on the updated ordering. Still as this may be undesirable in some cases, work is being conducted to provide an alternative option using unique numbering internal to AMSI to avoid effecting user data.

Due to the necessity of retaining inter-scale linkages through the load-balancing processes, the actual data migration undertaken to implement a specific load balancing plan is typically conducted internally by the AMSI system which is provided with buffers of data being load-balanced by each process on the appropriate scale. However, in order to maintain usability in more general situations, it is also possible to simply inform the AMSI system that the current load-balancing plan has already been accomplished by user-implemented data migration. This functionality combined with the ability to use user-designed load-balance planning algorithms allows AMSI to work with entirely external load-balancing libraries and algorithms, allowing scale-sensitive load-balancing to be used even when a specific multi-scale use case falls outside the bounds of capabilities built into the system.

### 2015-09-14 scripts/autobuild.sh
seperating release and debug builds into seperate build directories so no reconfigure is required to switch versions

@@ -1,8 +1,24 @@
 #!/bin/bash
-cd ../build
+
+if [ -z $1 ]; then
+  BUILD_TYPE=Debug
+else
+  BUILD_TYPE=$1
+fi
+
+if [ "$BUILD_TYPE" == "Debug" ]; then
+  BUILD_DIR=../build_debug
+elif [ "$BUILD_TYPE" == "Release" ] ; then
+  BUILD_DIR=../build_release
+fi
+
+if [ ! -d $BUILD_DIR ]; then
+  mkdir $BUILD_DIR
+fi
+cd $BUILD_DIR
 
 notifybuild \
-"../amsi_util/ ../amsi_meta/ ../amsi_interface/ ../amsi_control/ ../test/test_interface/linear_elasticity/ ../test/test_interface/nonlinear_elasticity/ ../test/test_interface/partitioned_linear_elastic/" \
+"../amsi_util/ ../amsi_meta/ ../amsi_interface/ ../amsi_control/ ../test/test_interface/linear_elasticity/ ../test/test_interface/nonlinear_elasticity/ ../test/test_interface/partitioned_linear_elastic/ ../test/test_interface/mesh_adaptivity/" \
 '' \
 'make install' \
-'AMSI GIT'
+"AMSI_$BUILD_TYPE"

### 2015-09-14 scripts/config.sh

@@ -1,13 +1,11 @@
 #!/bin/bash
 # CMake config for AMSI
 #
-# usage: ./configamsi.sh [build_type] [install_prefix]
+# usage: ./config.sh [build_type]
 #
 
-if [ ! -d ../build ]; then
-  mkdir ../build
-fi
-cd ../build
+
+
 rm -rf ./* #stupid and dangerous
 
 module load cmake
@@ -26,7 +24,7 @@ if [ "$HOSTNAME" == "q.ccni.rpi.edu" ]; then
 
    cmake \
     -DCMAKE_TOOLCHAIN_FILE=$CMAKE_XL_TOOLCHAIN \
-    -DCMAKE_BUILD_TYPE=Debug \
+    -DCMAKE_BUILD_TYPE=$BUILD_TYPE \
     -DBUILD_TESTS=OFF \
     -DCMAKE_INSTALL_PREFIX=$DEVROOT/install \
     -DSIM_MPI=bgmpi \
@@ -34,6 +32,7 @@ if [ "$HOSTNAME" == "q.ccni.rpi.edu" ]; then
     ..
 
 else
+  module load openmpi/1.3.3
   module load parmetis/OpenMPI-1.3.3
   module load zoltan/OpenMPI-1.3.3
   module unload parmetis/OpenMPI-1.3.3
@@ -41,7 +40,7 @@ else
   module load simmetrix/simModSuite
 
   cmake \
-    -DCMAKE_BUILD_TYPE=Debug \
+    -DCMAKE_BUILD_TYPE=$BUILD_TYPE \
     -DBUILD_TESTS=OFF \
     -DCMAKE_INSTALL_PREFIX=$DEVROOT/install \
     -DSIM_MPI=openmpi14 \

### 2015-09-16 doc/reports/sandia_final/sandia_final_conclusion.tex
final sandia report should be finished now

\label{sumup}
The development of the Adaptive Multi-scale Simulation Infrastructure has produced an initial set of tools capable of easing the implementation and execution of multi-scale simulations by combining robust legacy codes in a minimally intrusive manner. The AMSI libraries operate independently of the underlying simulation codes, both in terms of language of implementation and the numerics operations undertaken to acheive solution.

This freedom allows AMSI to be used for a wide variety of multiscale cases. However this freedom also means that it is possible to do things poorly, as with any sufficiently low-level system. Thus future developments are likely to be focused on building a hierarchy of abstractions atop the currently implemented low-level systems ease in the use of AMSI for the widest-used subset of the cases AMSI provides support for.

A set of operations for managing adaptive multi-scale simulations has been developed and implmented in AMSI, specifically support for dynamically adding and removing scale-tasks at run time. This support is being built on and enriched to allow this complex operation to happen with minimal programmer intervention, at present the support is low-level and some simulation developer intervention in code to manage the newly-initialized or removed scales is needed. Building on this operation and expanding the metadata model AMSI maintains of the simulation state will allow for the implemtation of fully dynamic simulation scale control: arbitrary initialization and destruction of non-primary scales, and allowing blocked process scales to be reassigned (temporarily or permanently) to higher-priority scales.

Additionally, support for load balancing an individual scale of a multi-scale simulation while maintaining the overall integrity of the multi-scale coupling operations has been developed. This allows more traditional load balancing operations -- vital to optimizing parallel performance and time-to-solution in traditional massively parallel simulations -- to be used on scale codes being used with the AMSI systems. This also represents an important step in the development of general adaptive capabilities for multi-scale simulations, where they greatest challenge is balancing the execution of many interacting (and possibly blocking) simulation scales.

### 2015-09-16 doc/reports/sandia_final/sandia_final_methods.tex
final sandia report should be finished now

@@
 \label{amsi-communication}
-Using data structures describing the dynamic state of the parallel execution space the AMSI can determine which scales will need to interact, and only for these interacting scales are the required internal buffers and scale-coupling structures allocated.
+By using data structures describing the dynamic state of the parallel execution space AMSI can determine which scales will need to interact, and only for these interacting scales are the required internal buffers and scale-coupling structures allocated.
 This approach is taken in order to reduce unneeded overhead caused by maintaining data structures specific to interacting scale-tasks, which would be the case when assuming that any two scale-tasks may interact.
+Dynamic redefintion of which scales interact is not currently possible, but will be implemented to support dynamic scale instantiation.
 

### 2015-09-16 doc/reports/sandia_final/sandia_final_results.tex
final sandia report should be finished now

@@
   \begin{center}
-    \includegraphics[height=2in]{biotissue-scales.png}
+    \includegraphics[height=3in]{biotissue-scales.png}
   \end{center}
@@
 
-The simulation has been run many times on the Amos BlueGene/Q managed by the Center For Computational Innovations at Rennselaer Polytechnic Institute, and has scaled up to 250k elements at macro-scale with over 16k cores (with >90\% assigned to micro-scale simulations as these represent the bulk of actual computation).
+The simulation has been run many times on the Amos BlueGene/Q managed by the Center For Computational Innovations at Rennselaer Polytechnic Institute, and has scaled up to 250k elements at macro-scale with over 16k cores (over 90\% assigned to micro-scale simulations as these represent the bulk of actual computation).
 Scaling beyond this should be possible but has not currently been conducted.
@@
   \begin{center}
-    \includegraphics[height=2in]{allTimes_2macro_20k.png}
+    \includegraphics[height=3in]{allTimes_2macro_20k.png}
   \end{center}
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[height=2in]{siam_pp_weak_scaling.png}
-  \end{center}
-  \caption{\small Strong scaling on 20k mesh}
-  \label{strong_scaling}
-\end{figure}
+Figure \ref{20k_first_iteration} is a weak scaling study conducted on a 20k element mesh, leaving the number of macro-scale processes constant and varying the number of micro-scale processes from just 32 to 8192 by powers of 2.
+The total time to assemble and solve a single tangent stiffness matrix at the engineering scale is considered (1 Newton-Raphson iteration), and different portions of the overall workload are shown.
+The actual matrix solve is unchanging as the number of macro-scale processes which compute the linear solve remains constant.
+Thus the matrix solve time represents the idealized optimal time to compute a single iteration.
+It is clear that as the number of micro-scale processes increases, the overall solve time approaches the matrix solve time, and while the communication time increases, it is still several orders of magnitude less than the overall solve time.
 
@@
   \begin{center}
-    \includegraphics[height=2in]{dogBone14k.png}
+    \includegraphics[height=3in]{siam_pp_weak_scaling.png}
+  \end{center}
+  \caption{\small Weak scaling on 20k mesh}
+  \label{weak_scaling}
+\end{figure}
+
+Figure \ref{weak_scaling} is a view of the same weak-scaling study, weighted by the total time-to-solution for each case.
+Thus it is clear that as the number of micro-scale processes increase sufficiently, the overall computational load is dominated by the matrix solve.
+Increasing the number of processes at macro-scale should result in a reduction of this computational domincation, but both the size of macro- and micro-scales (and their ratio) must be considered when trying to determine the optimal allocation size.
+This figure suggests that a 1:128 macro-to-micro ratio will result in equal time being spent on both computations, and allow local load balancing operations at individual scales to make the most impact.
+Of course overall time-to-solution can be minimzed by simply allocating as many processes as possible, but allocating in the above ratio, weighted by the efficacy of individual scale load-balancing to impact the performance of a single scale, should result in the best possible gain.
+Further discussion and analysis of this concept will be published in \cite{wtobin2015mlb}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=3in]{dogBone14k.png}
   \end{center}
@@
   \begin{center}
-    \includegraphics[height=2in]{siam_cse_initial.png}
+    \includegraphics[height=2.5in]{siam_cse_initial.png}
   \end{center}
-  \caption{\small Initial load distribution (20k elements)}
+  \caption{\small Unbalanced load distribution (20k elements)}
   \label{initial_load}
@@
   \begin{center}
-    \includegraphics[height=2in]{siam_cse_updated.png}
+    \includegraphics[height=2.5in]{siam_cse_updated.png}
   \end{center}
-  \caption{\small Load balancing (20k elements)}
+  \caption{\small Balanced load distribution (20k elements)}
   \label{load_balancing}
@@
 This is illustrative of the trend seen in the biotissue simulation, the number of micro-scale newton iterations conducted during the first incremental load step are not illustrative of the number of micro-scale newton iterations required during additional incremental load steps, which stay largely the same.
-Thus the first load balancing operation 'over-corrects' the issue, but the second and subsequent load-balancing operations result in better load balancing and run times overall.
+Thus the first load balancing operation over corrects the issue, but the second and subsequent load-balancing operations result in better load balancing and run times overall.
 This can be seen more clearly in figure \ref{10_step_lb} which shows the weights for 64 micro-scale processes with the same mesh as above over 9 incrememental loading steps, the left graph shows the load without balancing and the right shows the load with balancing.
@@
   \begin{center}
-    \includegraphics[height=2.5in]{siam_cse_ten.png}
+    \includegraphics[height=3in]{siam_cse_ten.png}
   \end{center}

### 2015-09-16 doc/reports/sandia_final/compile.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+pdflatex sandia_report_final.tex
+bibtext sandia_report_final.auf
+pdflatex sandia_report_final.tex
+bibtext sandia_report_final.auf
+pdflatex sandia_report_final.tex

### 2015-09-16 amsi_control/ControlService.cc
changes from the Q version of the code, fixes to MPI and uninitialized value stuff

@@ -96,18 +96,14 @@ namespace amsi {
 	// TODO: Change this so the user can specificy which processes to allow to output to stdout
 	// redirect std::cout to /dev/null in silent processes
 	int local_rank = task_man->getLocalTask()->LocalRank();
-	std::ofstream  dev_null("/dev/null");
-	std::streambuf * old_cout = std::cout.rdbuf();
-	
-	// local_rank != 0 ? std::cout.rdbuf(dev_null.rdbuf()) : NULL;
+	if(local_rank > 0)
+	  std::cout.setstate(std::ios_base::failbit);
 
 #       ifdef ZOLTAN
 	float version = 0.0;
 	Zoltan_Initialize(NULL,NULL,&version);
 #       endif 
 	result += task_man->Execute(argc,argv);
-	
-	std::cout.rdbuf(old_cout);
       }
       else
 	result = -1;

### 2015-09-16 amsi_interface/amsiInterface.cc

@@ -20,10 +20,15 @@ namespace amsi {
 
 #   ifdef SIM
     PMU_setCommunicator(AMSI_COMM_LOCAL);
+
+#   if HOST==BGQ
+    Sim_readLicenseFile("/gpfs/u/software/bgq/proprietary/simmetrix//license/license.txt");
+#   else
     Sim_readLicenseFile("/net/common/meshSim/license/license.txt");
+#   endif
     SimPartitionedMesh_start(NULL,NULL);
     SimMeshing_start();
-#   if HOST==BGQ
+#   if HOST!=BGQ
     SimField_start();
 #   endif
     gmi_sim_start();
@@ -35,7 +40,7 @@ namespace amsi {
   {
 #   ifdef SIM
     gmi_sim_stop();
-#   if HOST==GBQ
+#   if HOST!=BGQ
     SimField_stop();
 #   endif
     SimMeshing_stop();

### 2015-09-16 amsi_interface/apfFEA.cc

@@ -1,6 +1,5 @@
 #include "apfFEA.h"
 
-#include <cassert>
 #include <apfShape.h>
 
 namespace amsi {
@@ -21,7 +20,7 @@ namespace amsi {
 	memset(&dofs[0],0,analysis_size*sizeof(int));
 
 	dofs[analysis_rank] = local_dof_count;
-	MPI_Allgather(&dofs[analysis_rank],1,MPI_INTEGER,
+	MPI_Allgather(&local_dof_count,1,MPI_INTEGER,
 		      &dofs[0],1,MPI_INTEGER,
 		      analysis_comm);
 	

### 2015-09-16 amsi_interface/apfSimFEA.cc

@@ -6,8 +6,6 @@
 #include <apfSIMDataOf.h>
 #include <apfShape.h>
 
-#include <cassert>
-
 namespace amsi {
   namespace Analysis {
 

### 2015-09-19 doc/reports/sandia_final/sandia_final_impact.tex

@@
 Particularly SCOREC's ongoing work on incorporating our dynamic mesh operations with the Albany simulation system developed at Sandia National Labs.
-For multi-scale simulations using the Albany simulation code, we will make use of AMSI's capabilities to be minimally intrusive and only require modifications at those locations in code where multi-scale terms are needed.
+For multi-scale simulations using the Albany simulation code, we will make use of AMSI's capabilities to be minimally intrusive and only require modifications at those locations in code where scale-coupling terms are needed.

### 2015-09-19 doc/reports/sandia_final/sandia_final_introduction.tex

--- 2015-09-13 doc/reports/sandia_final_introduction.tex
@@
 In this case the multi-scale features are incorporated into the numerical implementation automatically, as a result of their presence in the mathematical model.
-Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales but opera ting on some of the same physical fields and values.
+Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales but opera ting on some of the same physical fields and values cite{shenoy97} \cite{weinan2003heterogenous}.
 

### 2015-10-30 doc/articles/amsi_overview/amsi_overview.tex
some work on an amsi-only paper, mostly commiting since im about to delete some things i may want later

\title{}
\author{W.R. Tobin, D. Fovargue, V.W.L. Chan, M.S. Shephard}
\date{\today}
\documentclass[11pt]{article}
\bibliographystyle{siam}

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}\label{introduction}

\label{single_scale_simulations}
Numerical simulation of physical systems is typically conducted using single scale models. The physical terms in these models are all situated within a narrow specturm of physical magnitudes surrounding the primary solution scale. The primary solution scale is determined primarily by the desired resolution of two classes of physical terms. The resoulution of the solution produced by simulation must be fine enough to be of utility to the end-user, and the resolution of variation of physical quantities required by the physical model must be fine enough to capture suitable variation to control the buildup of numerical error.

Single-scale models provide an averaged or generally accurate model of the overall evolution of a physical system in response to stimuli.

However all physical problems are multi-scale in nature, as the response of a system at the primary (or engineering) scale is in reality the result of actions taking place at scales many orders of magnitude removed.

\label{multi_scale_simulations}

Introduction of influence from scales orders of magnitude removed from the primary scale can take place in several different ways. This influence can be incorporated during the mathematical development of the physical model itself. In this case the multi-scale features are incorporated into the numerical implementation automatically, as a result of their presence in the mathematical model.

Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}. Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling equations to take into account the difference in scale, refered to as compression and reconstruction operators in \cite{weinan2003heterogenous} for up-scaling and down-scaling, respectively.

In order for a numerically multi-scale approach to be efficacious the domain at the scale of interest (for brevity the engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude smaller would make the simulation computationally intractable. Thus some computationally reasonable subset of the engineering scale is assigned tertiary scales to simulate. The precise choice of location of the tertiary simulations is largely dependent on the specific numerical algorithms used to implement the engineering scale simulation.

Typically the tertiary scale domain subset is related to locations in the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation.

\label{multiscale_paradigms}
Among numerically multi-scale simulations, there are two primary paradigms: concurrent and sequential multi-scale.

The sequential multi-scale paradigm involves the precomputation of parameters using a microscale simulation. These parameters are then used in a constitutive model for the macro-scale simulation \cite{garcia2008sequential}.

In the concurrent multi-scale paradigm, both scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.

\section{Adaptive Multiscale Simulation Infrastructure (amsi)}\label{amsi}

\section{Future Work}\label{future_work}

\bibliography{amsi}

\end{document}

### 2015-10-31 doc/articles/amsi_overview/amsi_overview.tex
work on a manuscript.... very rough so far

@@
-\title{}
+\title{The Adaptive Multiscale Simulation Infrastructure -- Core Features and Biological Tissue Simulation Feature Utilization}
 \author{W.R. Tobin, D. Fovargue, V.W.L. Chan, M.S. Shephard}
@@
 
-\label{single_scale_simulations}
-Numerical simulation of physical systems is typically conducted using single scale models.
-The physical terms in these models are all situated within a narrow specturm of physical magnitudes surrounding the primary solution scale.
-The primary solution scale is determined primarily by the desired resolution of two classes of physical terms.
-The resoulution of the solution produced by simulation must be fine enough to be of utility to the end-user, and the resolution of variation of physical quantities required by the physical model must be fine enough to capture suitable variation to control the buildup of numerical error.
+\section{Biotissue Multiscale Problem}\label{biotissue}
+[An overview of the biotissue problem, along with relevant references.
+Just needs to be sufficient for the the reader to understand the multi-scale structure of the problem, so they can follow along as specific sections of amsi are described, and then how biotissue uses those sections of amsi is discussed.]
 
-Single-scale models provide an averaged or generally accurate model of the overall evolution of a physical system in response to stimuli.
+\section{Adaptive Multiscale Simulation Infrastructure (AMSI)}\label{amsi}
+The Adaptive Multi-scale Simulation Infrastructure (AMSI) a set of libraries providing support for the implementation and execution of dynamic numerically multi-scale simulations on massively-parallel HPC machines.
+Currently AMSI provides interfaces for describing and managing the execution of individual scales in multi-scale simulations, for defining and managing the scale-coupling communication required by such simulations, and for planning and enacting scale-sensitive load balancing operations on individual scales in a multi-scale system.
+AMSI is designed to ease the incorporation of legacy single-scale simulation codes into multi-scale simulations.
+This is accomplished by minimizing the insertion of calls to the AMSI systems in a legacy simulation code to only those locations associated with the desired multi-scale interaction.
 
-However all physical problems are multi-scale in nature, as the response of a system at the primary (or engineering) scale is in reality the result of actions taking place at scales many orders of magnitude removed.
+Centralized control, while simple, is not possible in massively parallel simulation codes -- both single- and multi-scale -- as centralized dynamic control decisions require global barrier operations to effectively implement, resulting in lowered process usage and an increased time-to-solution.
+Decentralized control is thus required to achieve performant operation with the desired level of dynamic control flexibility.
+An initial approach taken to implement decentralized control decisions is a combination of deterministic control algorithms and decision implementation deferment.
+Deterministic algorithms operating independently on the same set of data will all arrive at the same result, and thus no further communication with respect to local implementation of such a control decision is necesarry.
+Decision implementation deferement is simply delaying the implementation of a specific control decision - which might otherwise introduce a parallel barier - until a suitable barier operation at which the control decision may be implemented is reached.
 
-\label{multi_scale_simulations}
+\section{Simulation Metamodel}\label{meta_model}
+AMSI runtime operations are predicated on a carefully designed set of simulation metadata.
+Minimal memory and computational footprints are a primary focus in the design and implementation of this metadata, in order to curtail any impact on underlying simulation codes used in the multi-scale simulation.
 
-Introduction of influence from scales orders of magnitude removed from the primary scale can take place in several different ways.
-This influence can be incorporated during the mathematical development of the physical model itself.
-In this case the multi-scale features are incorporated into the numerical implementation automatically, as a result of their presence in the mathematical model.
+The metadata retained by an individual rank is limited only to the metadata that rank will need to operate on during the execution of a simulation.
+There are four levels of the metadata hierarchy: rank-local, scale-local, scale-coupling, and global metadata.
+The determination of the minimal set of required metadata from each level of the hierarchy for an individual rank to possess during the progression of the multi-scale simulation is what concerns a large portion of the metadata management algorithms implemented in AMSI.
 
-Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
-Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling equations to take into account the difference in scale, refered to as compression and reconstruction operators in \cite{weinan2003heterogenous} for up-scaling and down-scaling, respectively.
+Local metadata currently consists primarily of simple information such as the set of scales assigned to the rank (currently limited to just a single scale, but overlapped scale assignment is a very desireable future development, see discussion in \ref{future_work}) as well as the scale-ranks of the local rank.
+The most important local metadata is the discrete quantity of locally-owned inter-scale data terms called a DataDistribution.
+The actual type(s) of the user data are unimportant to the metadata operations offered by AMSI, so long as the data can be serialized for parallel communication.
+This DataDistribution quantity is specific to a single scale assigned to the rank, and any number of DataDistributions can be created on each scale.
+The DataDistribution is converted to scale-local metadata using an AMSI process called assembly.
+This allows scale-local control decisions to be made using independently on each rank in a scale as opposed to globally.
 
-In order for a numerically multi-scale approach to be efficacious the domain at the scale of interest (for brevity the engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude smaller would make the simulation computationally intractable.
-Thus some computationally reasonable subset of the engineering scale is assigned tertiary scales to simulate.
-The precise choice of location of the tertiary simulations is largely dependent on the specific numerical algorithms used to implement the engineering scale simulation.
+Scale-local metadata is data modeling terms distributed across an entire scale, these are typically generated via assembly.
+Scale-local metadata such as an assembled DataDistribution are used to plan inter-scale communication, producing a piece of scale-local metadata called the CouplingPattern.
+The CouplingPattern models the parallel communication of coupling terms from the local scale to some coupled foreign scale.
+The CouplingPattern is transformed into a piece of scale-coupling metadata through a process called reconciliation.
+Naturally not all ranks in the foreign scale need to know the entire CouplingPattern, only the rank-local information which will guide individual parallel communication.
+Thus reconciliation is not communication of the entire CouplingPattern, but only individual terms of the coupling pattern, producing rank-local metadata on the foreign scale.
 
-Typically the tertiary scale domain subset is related to locations in the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation.
+Another key peice of scale-local metadata is the RankSet associated with non-local but coupled scales.
+The MPI_Comm of the foreign scale is necessarily MPI_COMM_NULL on the local rank set (as scales are currently non-overlapping), so explicit modeling of the ranks associated with coupled foreign scales is required.
+Explicit storage of individual ranks assigned to foreign (and local) scales is avoided wherever possible in order to reduce single-rank memory impact as the simulation grows.
 
-\label{multiscale_paradigms}
-Among numerically multi-scale simulations, there are two primary paradigms: concurrent and sequential multi-scale.
+Global metadata is currently nonexistent, maintaining zero global metadata is important to the design philosophy of AMSI, though it may be required at some point.
+Though global metadata determined essential to AMSI will be designed to be exteremely minimal.
 
-The sequential multi-scale paradigm involves the precomputation of parameters using a microscale simulation.
-These parameters are then used in a constitutive model for the macro-scale simulation \cite{garcia2008sequential}.
+\label{amsi_scales}
+Individual scales in a simulation are declared during the initialization of the simulation, follow by the determination and asignment of RankSets to scales, guided by a supplied configuration file or user-implemented algorithm.
+Scale-coupling relations are also declared at initialization time, though this decleration is limited to the indication that two scale may interact at some point during the simulation, and nothing more.
 
-In the concurrent multi-scale paradigm, both scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.
+RankSets are simply mathematical sets of MPI ranks implemented so as to take advantage of any mathematical conveniences to minimize explicit storage whenever possible.
 
-\section{Adaptive Multiscale Simulation Infrastructure (amsi)}\label{amsi}
+Scales are declared to be related if some quantities of interest (such as tensor field values) are transmitted between them for scale-coupling.
+This coupling relationship is not assumed to be symmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be seperately declared.
+
+\label{amsi_communication}
+Scale coupling communication is handled by the AMSI system using DataDistributions and CommuncationPatterns.
+A data distribution is a meta representation of individual units of scale-coupling data distributed across a single scale-task.
+The implementation of the coupling data is arbitrary (so long as it can be serialized for communication).
+A data distribution represents the smallest unit of data important to the scale-coupling communication, which is necessarily distinct for various multi-scale use cases.
+
+\label{amsi_data_migration_balancing}
+If a local installation of the Zoltan \cite{ZoltanOverviewArticle2002} \cite{ZoltanIsorropiaOverview2012} load balancing library is available, the load balancing planning algorithms of Zoltan can be used underlying the scale-sensitive load balancing planning of AMSI.
+A minimal set of load-balancing algorithms are provided with AMSI.
+Since there is no general solution to the dynamic load balancing problem any set of provided load-balancing algorithms will prove insufficient for some use case.
+Thus AMSI provides users with the ability to create and register their own load-balancing algorithms specific to the requirements of a given numerical implementation.
+
+Due to the necessity of retaining inter-scale linkages through the load-balancing processes, the actual data migration undertaken to implement a specific load balancing plan is typically conducted internally by the AMSI system which is provided with buffers of data being load-balanced by each process on the appropriate scale.
+However, in order to maintain usability in more general situations, it is also possible to simply inform the AMSI system that the current load-balancing plan has already been accomplished by user-implemented data migration.
+This functionality combined with the ability to use user-designed load-balance planning algorithms allows AMSI to work with entirely external load-balancing libraries and algorithms, allowing scale-sensitive load-balancing to be used even when a specific multi-scale use case falls outside the bounds of capabilities built into the system.
+
+\section{Biotissue Metadata}
 

### 2015-10-31 amsi_interface/apfSimFEA.cc

@@ -66,10 +66,10 @@ namespace amsi {
 	for(GFIter gfiter = GM_faceIter(model); (entity = GFIter_next(gfiter));)
 	  fixed_dofs += Entity_ApplyBC_Dirichlet(entity,2);
       }
-    for(GEIter geiter = GM_edgeIter(model); (entity = GEIter_next(geiter));)
+      for(GEIter geiter = GM_edgeIter(model); (entity = GEIter_next(geiter));)
 	fixed_dofs += Entity_ApplyBC_Dirichlet(entity,1);
-
-  for(GVIter gviter = GM_vertexIter(model); (entity = GVIter_next(gviter));)
+      
+      for(GVIter gviter = GM_vertexIter(model); (entity = GVIter_next(gviter));)
 	fixed_dofs += Entity_ApplyBC_Dirichlet(entity,0);
       
       std::cout << "There are " << fixed_dofs << " dofs fixed by essential boundary conditions." << std::endl;

### 2016-01-03 doc/articles/amsi_biotissue/paper.tex
adding article work for amsi capabilities and utilization in biotissue

\title{The Adaptive Multi-scale Simulation Infrastructure - Design and Implementation of a soft tissue simulation}
\author{W.R. Tobin, D. Fovargue, V.W.L. Chan, M.S. Shephard}
\date{\today}
\documentclass[11pt]{siamltex1213}
\bibliographystyle{siam}

\usepackage{graphicx}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

\section{Note}\label{note}
This article predominantly discusses multi-scale simulations - wherein two or more separate physical scales are involved in the simulation - the discussion is equally applicable to multi-model simulations - wherein two or more numerical or physical models operating on the same domain (possibly with differing discretizations) are involved in the simulation - as well as to multi-scale-multi-model simulations.

\section{Introduction}\label{introduction}
\label{single_scale_simulations}
Most physical models and implementations using numerical methods operate at a single scale characteristic to the problem. The scale at which the preponderance of attributes governing the underlying physics being modelled are measured and the resolution of the numerical solution is considered acceptable for the desired accuracy of the simulation.

\label{multi_scale_simulations}
Introduction of influence from scales orders of magnitude removed from the primary scale can take place in several different ways. This influence can be incorporated during the mathematical development of the physical model itself. In this case the multi-scale features are incorporated into the numerical implementation automatically, as a result of their presence in the mathematical model.

Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}. Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling equations to take into account the difference in scale, referred to as compression and reconstruction operators in \cite{weinan2003heterogenous} for up-scaling and down-scaling, respectively.

In order for a numerically multi-scale approach to be efficacious the domain at the scale of interest (for brevity the engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude smaller would make the simulation computationally intractable. Thus some computationally reasonable subset of the engineering scale is assigned tertiary scales to simulate. The precise choice of location of the tertiary simulations is largely dependent on the specific numerical algorithms used to implement the engineering scale simulation.

Typically the tertiary scale domain subset is related to locations in the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation.

\label{multiscale_paradigms}
Among numerically multi-scale simulations, there are two primary paradigms: concurrent and sequential multi-scale.

The sequential multi-scale paradigm involves the precomputation of parameters using a microscale simulation. These parameters are then used in a constitutive model for the macro-scale simulation \cite{garcia2008sequential}.

In the concurrent multi-scale paradigm, both scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.

\subsection{Adaptive Multi-scale Simulation Infrastructure}\label{amsi}
The Adaptive Multi-scale Simulation Infrastructure (amsi) is a set of libraries and tools developed at the Scientific Computation Research Center (SCOREC) at Rensselaer Polytechnic Institute. Amsi is designed to support the implementation and execution of dynamic numerically multi-scale simulations on massively-parallel HPC machines. Currently amsi provides interfaces for describing and managing the execution of individual scales in a multi-scale simulation, for defining and managing the scale-coupling communication required by such simulations, and for planning and enacting scale-sensitive load balancing operations on individual scales in a multi-scale system.

The interfaces provided by the amsi libraries are intended to facilitate the coupling of functional (possibly legacy) simulation codes already in usage for single-scale problems. The currently implemented interfaces are physics-agnostic, operating only on computational quantities with no model of a term's relation to the mathematical derivation of the multi-scale interaction. Incorporation of meta-information and services targeted toward specific multi-scale physical couplings represents a desirable extension of the amsi libraries, constituting a rich area of future work.

Amsi operates by maintaining a set of minimal simulation metadata in order to model various quantities of interest during simulation execution. A decentralized approach toward control is taken; dynamic control decisions are made and implemented at runtime during collective operations. In order to avoid the introduction of unnecessary parallel barriers into the code, amsi control decisions are only made during operations which are already collective over the set of processes effected by a given control decision.

\label{amsi_scales}
A simple configuration file is supplied via command-line parameter containing the initial configuration of the process sets and their assignment to individual scales, as well as declaring the coupling between scales in the simulation. The deceleration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.

Process sets are mathematical sets of MPI ranks implemented so as to take advantage of any mathematical conveniences to minimize explicit storage whenever possible. Process sets are required to be non-overlapping in order for amsi control mechanisms to operate correctly. Ongoing developments are targeted at removing this requirement to allow a richer set of dynamic control options and increase overall utilization in the case of coupled scales blocking on one another.

Deceleration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling deceleration during simulation initialization is used to instantiate data structures specific to scale-coupling operations. This limits memory overhead by not allocating space for unneeded coupling operations.

Scales are declared to be related if some quantities of interest (such as physical tensor field values) are transmitted between them for scale-coupling. This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared. For brevity this may be written as A~B for A is coupled to B.

While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the amsi implementation. Any scale is capable of informing the process set associated with a coupled scale of the addition or deletion of individual scale-tasks. At present it is dependent on the developer to implement the capability of a scale to accept additional scale-tasks, though algorithms and hooks for developer-designed distribution algorithms are supplied to ease the implementation.

\label{amsi_communication}
Scale coupling communication is handled by the amsi system using data distributions and communication patterns. A data distribution is a meta representation of individual units of scale-coupling data distributed across a single scale-task. The implementation of the coupling data is arbitrary (so long as it can be serialized for communication). A data distribution represents the smallest unit of data important to the scale-coupling communication, which is necessarily distinct for various multi-scale use cases.

All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, essentially denoted by the term's index in a serialized communication buffer.

\label{amsi_load_balancing}
If a local installation of the Zoltan \cite{ZoltanOverviewArticle2002} \cite{ZoltanIsorropiaOverview2012} load balancing library is available, the load balancing planning algorithms of Zoltan can be used underlying the scale-sensitive load balancing planning of amsi. A minimal set of load-balancing algorithms are provided with amsi. Since there is no general solution to the dynamic load balancing problem any set of provided load-balancing algorithms will prove insufficient for some use case. Thus amsi provides users with the ability to create and register their own load-balancing algorithms specific to the requirements of a given numerical implementation.

Due to the necessity of retaining inter-scale linkages through the load-balancing processes, the actual data migration undertaken to implement a specific load balancing plan is typically conducted internally by the amsi system which is provided with buffers of data being load-balanced by each process on the appropriate scale. However, in order to maintain usability in more general situations, it is also possible to simply inform the amsi system that the current load-balancing plan has already been accomplished by user-implemented data migration. This functionality combined with the ability to use user-designed load-balance planning algorithms allows amsi to work with entirely external load-balancing libraries and algorithms, allowing scale-sensitive load-balancing to be used even when a specific multi-scale use case falls outside the bounds of capabilities built into the system.

\subsection{Soft Tissue Simulation}\label{biotissue}

The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation. The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the global simulation. Specific quantities needed to compute the elemental tangent stiffness matrices are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh, see figure \ref{biotissue_hierarchy}.

\begin{figure}
  \begin{center}
    \includegraphics[height=2in]{biotissue_visual_hierarchy.eps}
  \end{center}

  \caption{\small biotissue multi-scale problem}
  \label{biotissue_hierarchy}
\end{figure}

The incremental displacements (the displacement deltas generated by the previous Newton iteration) on each engineering-scale finite element node are used to distort the a dimensionless representative volume element (RVE) resulting in each micro-scale node being displaced. Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation. The micro-scale boundary condition necessitates that the force exerted by each fiber network node located on the boundary of the RVE be zero. The micro-scale global Jacobian is formulated along with the force vector, wherein the boundary condition is enforced by setting specific rows of the force vector associated with boundary nodes to zero.

Another Newton-Raphson iterative process occurs at micro-scale to converge to the set of displacements resulting in force equilibrium for all the fibers within the fiber network constituting the RVE. After a micro-scale simulation converges to a solution, the Jacobian and force vector are then evaluated again at the solution position, the force exerted along the boundary of the RVE is summed for each of the principal and shear components of a stress tensor, which is then dimensionalized and sent back to the engineering-scale for use in the elemental tangent stiffness matrix formulation. More detail into the derivation of this multi-scale system can be found in \cite{stylianopoulos2008thesis} \cite{agoram2001coupled} \cite{stylianopoulos2007multiscale} . Discussion of the dimensionalization of the dimensionless force terms produced by each fiber network can be found in \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.

%Typically in the displacement-based finite element method, a constitutive model depending on some set of material properties is derived from a set of governing physical partial differential equations. In the case of the Biotissue multi-scale code, the constitutive model governing the systemic response to the boundary conditions is supplemented by the micro-scale force-balance simulation. The Newton-Raphson method is used to iterate to the solution for a particular loading state, and at each Newton iteration the global tangent stiffness matrix is assembled from elemental systems generated from each element in the engineering-scale mesh. In typically FEM the elemental system is derived from the governing material model used for the specific type of material being simulated; for the Biotissue code the necessary values are instead provided as

Thus the simulation as a whole is essentially doubly nonlinear, as during each engineering-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.

\subsection{Biotissue Scale-Sensitive Load Balancing}\label{biotissue_load_balancing}

Scale-sensitive load balancing is conducted on the micro-scale RVE distribution at locations in the code where communication bottlenecks are already present. This prevents the introduction of new bottlenecks which could result in reduced performance. Macro-scale Newton-iterations are the most frequently occuring multi-scale communication bottleneck at which we chose to conduct the load balancing operation, and macro-scale incremental load steps the least-frequent. These represent the two extreme cases considered in our analysis on the efficacy of the scale-sensitive load-balancing scheme adopted for the biotissue simulation.

Each micro-scale RVE simulation involves the assembly and solve of a linear system of equations on the order of 10,000 unknowns for each micro-scale Newton iteration. While the total solve time is dependent on the number of nonlinear iterations and the hardware executing the code, this system is sufficiently small that parallelization of the RVE code itself is unwarranted, as the introduction of communication overhead into the RVE solution process would result in decreased performance, despite the modest gain to computation time. However, as each RVE represents an independent task to be accomplished by the simulation, they may be freely distributed across the processes assigned to the micro-scale computation. The only restriction to their (re)distribution for load balancing purposes is that the relation to a specific macro-scale numerical integration point must be maintained.

As each RVE is a nonlinear simulation their is no a-prior estimate of the computational load each RVE represents, so the optimal initial distribution is simply treating the weight of each RVE as 1. Thus each micro-scale process is assigned either $\floor*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$ or $\ceil*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$. Once the first load-balancing event is reached, regardless of the granularity at which load-balancing is occuring for a given case, each RVE has recorded the total number of newton iterations it has conducted since the previous load-balancing event. This number is used as a heuristic weighting metric to determine the computational load of each individual RVE. The iteration count metric was chosen as each iteration takes a similar level of work to complete, up to the difference in the number of RVEs for each unique fiber network. Further, assuming linear incremental loading, regions in the engineering-scale mesh located near stress concentrators will result in 'more interesting' boundary conditions for the micro-scale problems related to numerical integration points in that region, which may result in the RVE requiring more Newton iterations to converge to an adequate solution, since the boundary conditions may lie farther from the initial state of the RVE in the regime of convergence for the nonlinear problem.

\section{Results}\label{results}

\section{Future Work}\label{future_work}

Work on scale-balancing is ongoing for the Biotissue test problem. Using this work as a basis, a set of algorithms to provide support for more general scale-balancing operations will be developed and packaged into amsi.

Embedded fiber-matrix RVEs separately parallelized and incorporated into multi-scale model, used at most error-sensitive locations \cite{lake2012mechanics} \cite{zhang2013cross} \cite{zhang2013coupled}.

\bibliography{citations}

\end{document}

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
working on the paper

@@
 Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
-Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling equations to take into account the difference in scale, referred to as compression and reconstruction operators in \cite{weinan2003heterogenous} for up-scaling and down-scaling, respectively.
+Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling operations to account for differences in dimension as well as possibly dimensionalizing terms computed in a non-dimensional domain.
 
@@
 \label{multiscale_paradigms}
-Among numerically multi-scale simulations, there are two primary paradigms: concurrent and sequential multi-scale.
-
-The sequential multi-scale paradigm involves the precomputation of parameters using a microscale simulation.
-These parameters are then used in a constitutive model for the macro-scale simulation \cite{garcia2008sequential}.
+Multi-scale simulations fall broadly into two distinct paradigms concurrent and sequential multi-scale.
 
 In the concurrent multi-scale paradigm, both scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.
+
+In the sequential multi-scale paradigm, tertiary scales compute parameters which are converted to quantities of utility to a primary scale.
+Each scale must be processed sequentially and is essentially blocked on the completion of the previous scale \cite{garcia2008sequential}.
+This multi-scale model maps intuitively to phased communication models related to bulk synchronous parallel computation \cite{valiant1990bridging}.
 
@@
 
-The interfaces provided by the amsi libraries are intended to facilitate the coupling of functional (possibly legacy) simulation codes already in usage for single-scale problems.
+The interfaces provided by the amsi libraries are intended to facilitate the coupling of functional (possibly legacy) simulation codes already in usage for single-scale problems, in order to avoid development overhead associated with casting a simulation into a set of framework-specific components.
 The currently implemented interfaces are physics-agnostic, operating only on computational quantities with no model of a term's relation to the mathematical derivation of the multi-scale interaction.
@@
 In order to avoid the introduction of unnecessary parallel barriers into the code, amsi control decisions are only made during operations which are already collective over the set of processes effected by a given control decision.
+
+\subsection{Novelty} %rename
+Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{parker2006component} \cite{} \cite{} and these frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah}.
+
+These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) in order to take full advantage of such frameworks.
+Leveraging a component-based design is valuable from a software engineering perspective, but also introduces new complexities involved in defining the set of components and how they interoperate in the framework shifting some complexity up a level of abstraction.
+The core features of amsi used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.
+
+simulation flow control is determined solely by the runtime system.
+A framework may also limit specific features of the numerical implementation to those provided by the framework.
+Domain discretization models and associated adaptive processes -- structured and unstructured meshing and associated refinement algorithms, particle-in-cell methods, etc.
+-- convergence algorithms for both the minimization of the relevant residuals (Newton-Raphson, etc.) as well as iterative solution of the resulting linear system (KSP, GMRES, ILU, etc.).
+Amsi supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations built on top of other software supplied by SCOREC \cite{core} and the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}.
+Usage of the supplied numerical components is not required to make use of the multi-scale meta-modeling, communication, dynamic instantiation, and scale-sensitive load balancing facilities supplied by amsi.
+
+While implementation of desired numerical operations into a framework is of course possible, it may prove an additional undesirable overhead for an application programmer interested only in utilizing the multi-scale framework to implement simulation, rather than implementing the various numerical operations often perfected in widely available packages and libraries \cite{} \cite{} \cite{}.
 
@@
 \label{amsi_communication}
-Scale coupling communication is handled by the amsi system using data distributions and communication patterns.
-A data distribution is a meta representation of individual units of scale-coupling data distributed across a single scale-task.
-The implementation of the coupling data is arbitrary (so long as it can be serialized for communication).
+Scale coupling communication is guided by data distributions and communication patterns.
+A data distribution is a representation of individual units of scale-coupling data distributed across a single scale-task.
+The format of the coupling data is arbitrary (so long as it can be serialized and provided for use by MPI communication primitives).
 A data distribution represents the smallest unit of data important to the scale-coupling communication, which is necessarily distinct for various multi-scale use cases.

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
working on the paper

@@
 \label{amsi_scales}
-A simple configuration file is supplied via command-line parameter containing the initial configuration of the process sets and their assignment to individual scales, as well as declaring the coupling between scales in the simulation.
+A configuration file is supplied during command invocation containing the initial configuration of the process sets and their assignment to individual scales, as well as declaring the coupling between scales in the simulation.
 The deceleration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
bit more work

@@
 \section{Note}\label{note}
-This article predominantly discusses multi-scale simulations - wherein two or more separate physical scales are involved in the simulation - the discussion is equally applicable to multi-model simulations - wherein two or more numerical or physical models operating on the same domain (possibly with differing discretizations) are involved in the simulation - as well as to multi-scale-multi-model simulations.
+This article predominantly discusses multi-scale simulations - wherein two or more separate physical scales are involved in the simulation - the discussion is equally applicable to multi-model simulations - wherein two or more numerical or physical models operating on the same domain (possibly with differing discretizations) are involved in the simulation - as well as to simulations which are both.
 
@@
 Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
-Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact in a meaningful way by passing key physical terms between scales, typically this will involve up-/down-scaling operations to account for differences in dimension as well as possibly dimensionalizing terms computed in a non-dimensional domain.
+Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact through passing key physical terms between scales.
+This scale-coupling will involve up-/down-scaling operations to account for differences in dimension, as well as possibly dimensionalizing terms computed in a non-dimensional domain.
 
-In order for a numerically multi-scale approach to be efficacious the domain at the scale of interest (for brevity the engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude smaller would make the simulation computationally intractable.
+In order for a numerically multi-scale approach to be effective the domain at the scale of interest (or engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude smaller would make the simulation computationally intractable.
 Thus some computationally reasonable subset of the engineering scale is assigned tertiary scales to simulate.
@@
 
-Typically the tertiary scale domain subset is related to locations in the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation.
+Often the subset(s) of the domain associated with the tertiary scale(s) are locations in the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation, in order to leverage the most power from the multi-scale approach.
 
-\label{multiscale_paradigms}
-Multi-scale simulations fall broadly into two distinct paradigms concurrent and sequential multi-scale.
-
-In the concurrent multi-scale paradigm, both scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.
+In the concurrent multi-scale paradigm, all scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.
 
 In the sequential multi-scale paradigm, tertiary scales compute parameters which are converted to quantities of utility to a primary scale.
-Each scale must be processed sequentially and is essentially blocked on the completion of the previous scale \cite{garcia2008sequential}.
-This multi-scale model maps intuitively to phased communication models related to bulk synchronous parallel computation \cite{valiant1990bridging}.
+Each scale (though not necessarily each scale-task, representing an individual operation or simulation occurring in a scale) must be processed sequentially \cite{garcia2008sequential} and may be blocked on the completion of the previous scale.
+This multi-scale model maps intuitively to phased parallel communication models found in the commonly used bulk synchronous parallel abstraction \cite{valiant1990bridging}.
 

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
removed load balancing section

@@
 \subsection{Soft Tissue Simulation}\label{biotissue}
-
 The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation.
 The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the global simulation.
-Specific quantities needed to compute the elemental tangent stiffness matrices are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh, see figure \ref{biotissue_hierarchy}.
+Specific quantities needed to compute the elemental tangent stiffness matrices are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}.
 
@@
   \end{center}
-
   \caption{\small biotissue multi-scale problem}
@@
 Thus the simulation as a whole is essentially doubly nonlinear, as during each engineering-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process. 
-
-\subsection{Biotissue Scale-Sensitive Load Balancing}\label{biotissue_load_balancing}
-
-Scale-sensitive load balancing is conducted on the micro-scale RVE distribution at locations in the code where communication bottlenecks are already present.
-This prevents the introduction of new bottlenecks which could result in reduced performance.
-Macro-scale Newton-iterations are the most frequently occuring multi-scale communication bottleneck at which we chose to conduct the load balancing operation, and macro-scale incremental load steps the least-frequent.
-These represent the two extreme cases considered in our analysis on the efficacy of the scale-sensitive load-balancing scheme adopted for the biotissue simulation.
-
-Each micro-scale RVE simulation involves the assembly and solve of a linear system of equations on the order of 10,000 unknowns for each micro-scale Newton iteration.
-While the total solve time is dependent on the number of nonlinear iterations and the hardware executing the code, this system is sufficiently small that parallelization of the RVE code itself is unwarranted, as the introduction of communication overhead into the RVE solution process would result in decreased performance, despite the modest gain to computation time.
-However, as each RVE represents an independent task to be accomplished by the simulation, they may be freely distributed across the processes assigned to the micro-scale computation.
-The only restriction to their (re)distribution for load balancing purposes is that the relation to a specific macro-scale numerical integration point must be maintained.
-
-As each RVE is a nonlinear simulation their is no a-prior estimate of the computational load each RVE represents, so the optimal initial distribution is simply treating the weight of each RVE as 1.
-Thus each micro-scale process is assigned either $\floor*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$ or $\ceil*{\frac{\text{num\_rves}}{\text{num\_micro\_procs}}}$.
-Once the first load-balancing event is reached, regardless of the granularity at which load-balancing is occuring for a given case, each RVE has recorded the total number of newton iterations it has conducted since the previous load-balancing event.
-This number is used as a heuristic weighting metric to determine the computational load of each individual RVE.
-The iteration count metric was chosen as each iteration takes a similar level of work to complete, up to the difference in the number of RVEs for each unique fiber network.
-Further, assuming linear incremental loading, regions in the engineering-scale mesh located near stress concentrators will result in 'more interesting' boundary conditions for the micro-scale problems related to numerical integration points in that region, which may result in the RVE requiring more Newton iterations to converge to an adequate solution, since the boundary conditions may lie farther from the initial state of the RVE in the regime of convergence for the nonlinear problem.
-
 \section{Results}\label{results}

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
working on paper

@@
 
-While implementation of desired numerical operations into a framework is of course possible, it may prove an additional undesirable overhead for an application programmer interested only in utilizing the multi-scale framework to implement simulation, rather than implementing the various numerical operations often perfected in widely available packages and libraries \cite{} \cite{} \cite{}.
+Implementing required numerical operations inside an existing framework is of course possible, though it may prove an additional -- undesirable overhead for an application programmer interested only in utilizing the multi-scale framework to implement a simulation, rather than implementing the various numerical operations often optimized in widely available packages and libraries.
 
@@
 
-All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, essentially denoted by the term's index in a serialized communication buffer.
+All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer.
 

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
adding a couple of good references, need to dig into them more however

@@
 Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact through passing key physical terms between scales.
-This scale-coupling will involve up-/down-scaling operations to account for differences in dimension, as well as possibly dimensionalizing terms computed in a non-dimensional domain.
+This scale-coupling will involve up-/down-scaling operations to account for differences in scale dimension, as well as possibly dimensionalizing terms computed in a non-dimensional domain.
 
@@
 \subsection{Novelty} %rename
-Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{parker2006component} \cite{} \cite{} and these frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah}.
+Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{parker2006component} \cite{chopard2011framework} \cite{} and these frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah}.
 
-These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) in order to take full advantage of such frameworks.
+These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or high-level modelling language in order to take full advantage of such frameworks.
 Leveraging a component-based design is valuable from a software engineering perspective, but also introduces new complexities involved in defining the set of components and how they interoperate in the framework shifting some complexity up a level of abstraction.
@@
 \label{amsi_communication}
-Scale coupling communication is guided by data distributions and communication patterns.
+Scale-coupling communication is guided by data distributions and communication patterns.
 A data distribution is a representation of individual units of scale-coupling data distributed across a single scale-task.
@@
 All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer.
+At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in a future release to automate application of such operations during scale-coupling.
 

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
more small work

@@
 
-The incremental displacements (the displacement deltas generated by the previous Newton iteration) on each engineering-scale finite element node are used to distort the a dimensionless representative volume element (RVE) resulting in each micro-scale node being displaced.
+The incremental displacements (deltas generated by the previous Newton-Raphson iteration) on each engineering-scale finite element node are used to distort the a dimensionless representative volume element (RVE) through a dimensionalized abstraction of the RVE embedded in the engineering domain and an associated displacement mapping term $\frac{d\mathbf{RVE}}{d\mathbf{FE}}$, .
+This results in each micro-scale node being displaced.
 Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation.

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
working on the paper

--- 2016-01-05 doc/articles/amsi_biotissue/paper.tex
@@
-\title{The Adaptive Multi-scale Simulation Infrastructure - Design and Implementation of a soft tissue simulation}
-\author{W.R. Tobin, D. Fovargue, V.W.L. Chan, M.S. Shephard}
+\title{The Adaptive Multi-scale Simulation Infrastructure - Design and Application}
+\author{W.R. Tobin, V.W.L. Chan, M.S. Shephard}
 \date{\today}
@@
 
-The incremental displacements (the displacement deltas generated by the previous Newton iteration) on each engineering-scale finite element node are used to distort the a dimensionless representative volume element (RVE) resulting in each micro-scale node being displaced.
+The incremental displacements (deltas generated by the previous Newton-Raphson iteration) on each engineering-scale finite element node are used to distort the a dimensionless representative volume element (RVE) through a dimensionalized abstraction of the RVE embedded in the engineering domain and an associated displacement mapping term $\frac{d\mathbf{RVE}}{d\mathbf{FE}}$, the first order relation between the incremental displacement of an RVE with respect to the displacement of the associated mesh element.
+This results in each micro-scale node being displaced.
 Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation.
@@
 Thus the simulation as a whole is essentially doubly nonlinear, as during each engineering-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
+
 \section{Results}\label{results}
+The biotissue simulation was run on the AMOS BG/Q machine housed at the CCI facilities.
+Both strong and weak scaling studies were conducted up to an eighth of the total machine core count, or 16384 cores.
+The ratio of macro-scale to micro-scale assigned processes was held near 1:128, though the total core count was kept as an integer power of 2 in order for the SLURM job scheduler to submit the jobs to the compute nodes in a timely manner.
+The precise calculation used to determine macro and micro ratio is given by
+$$ 
+num_macro = N 
+num_cores = N*128
+num_micro = num_cores - N
+$$
+thus as the number of cores assigned to the macro-scale computation scales, the ratio approaches 1:128.
+
+Separate timings were taken for various sections of the simulation code, isolating the micro-scale computation, macro-scale linear system assembly and calculation of elemental linear systems, and macro-scale linear system solution.
+In addition, the time for spent in the amsi subroutines for communication was also measured.
+Thus the behavior of each section of the code is independently evaluated in our scaling study, as well as the overall execution time.
+
+Weak scaling.
+
+Strong scaling.
+
+Amsi-specific results of some sort...?
 
@@
 
-Work on scale-balancing is ongoing for the Biotissue test problem.
+Work on scale-balancing is ongoing for the biotissue test problem.
 Using this work as a basis, a set of algorithms to provide support for more general scale-balancing operations will be developed and packaged into amsi.
+
+Scale-sensitive load-balancing operations have been developed for the biotissue multi-scale test problem.
+Unfortunately the problem does not exhibit load persistence in the current parallel task discretization, so an alternative parallel discretization such as overdecomposition with work-stealing to enforce more local balancing operations and occasional global balancing operations is being explored.
 

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
more work on the paper

@@
 
-These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or high-level modelling language in order to take full advantage of such frameworks.
+These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or high-level modelling language in order to take full advantage the services provided by the framework.
 Leveraging a component-based design is valuable from a software engineering perspective, but also introduces new complexities involved in defining the set of components and how they interoperate in the framework shifting some complexity up a level of abstraction.
@@
 
-simulation flow control is determined solely by the runtime system.
+Typically simulation flow control is determined solely by the runtime system.
 A framework may also limit specific features of the numerical implementation to those provided by the framework.
-Domain discretization models and associated adaptive processes -- structured and unstructured meshing and associated refinement algorithms, particle-in-cell methods, etc.
+This includes domain discretization models and associated adaptive processes -- structured and unstructured meshing and associated refinement algorithms, particle-in-cell methods, etc.
 -- convergence algorithms for both the minimization of the relevant residuals (Newton-Raphson, etc.) as well as iterative solution of the resulting linear system (KSP, GMRES, ILU, etc.).
-Amsi supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations built on top of other software supplied by SCOREC \cite{core} and the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}.
+
+Amsi supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations of these interfaces built on top of other software supplied by SCOREC \cite{core} for finite element analysis operations as well as the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}.
 Usage of the supplied numerical components is not required to make use of the multi-scale meta-modeling, communication, dynamic instantiation, and scale-sensitive load balancing facilities supplied by amsi.

### 2016-01-05 doc/articles/amsi_biotissue/paper.tex
paper work

@@
 
-Often the subset(s) of the domain associated with the tertiary scale(s) are locations in the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation, in order to leverage the most power from the multi-scale approach.
+Often the subset(s) of the domain associated with the tertiary scale(s) are areas of the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation, in order to leverage the most power from the multi-scale approach.
 
@@
 The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation.
-The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the global simulation.
-Specific quantities needed to compute the elemental tangent stiffness matrices are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}.
+The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the macro-scale simulation.
+%
+\begin{align}
+\nabla \cdot \pmb{\sigma} =& \; 0 \text{ in } \Omega \label{momentum_balance} \\
+\pmb{\sigma} \cdot \pmb{n} =& \; \pmb{t} \text{ on } \partial \Omega \nonumber
+\end{align}
+%
+Specific quantities needed to compute the elemental tangent stiffness matrices and force vectors are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}.
 
@@
 
-Thus the simulation as a whole is essentially doubly nonlinear, as during each engineering-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
+Thus the simulation as a whole has two Newton-Raphson convergence operations occurring.
+During each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
 
@@
 The precise calculation used to determine macro and micro ratio is given by
-$$ 
+%
 num_macro = N 
@@
 num_micro = num_cores - N
-$$
-thus as the number of cores assigned to the macro-scale computation scales, the ratio approaches 1:128.
+%
+thus as the number of cores assigned to the macro-scale computation increases, the ratio approaches 1:128.
 
@@
 
-Weak scaling.
+Weak scaling results...
 
-Strong scaling.
+Strong scaling results...
 
@@
 
-\section{Future Work}\label{future_work}
+Could include utilization/idle time metrics, but those are far more relevant to the load balancing paper...
 
-Work on scale-balancing is ongoing for the biotissue test problem.
-Using this work as a basis, a set of algorithms to provide support for more general scale-balancing operations will be developed and packaged into amsi.
+\section{Conclusions and Future Work}\label{future_work}
+
+The Adaptive Multi-scale Simulation Infrastructure provides a valuable set of tools to facilitate the implementation and execution of multi-scale simulations without imposing restrictions on the application developer.
+Usage of the core operations of the amsi libraries does not require adherence to a specific set of interfaces nor usage of an abstract modeling language, which complicate inclusion of legacy components into multi-scale simulations.
 

### 2016-01-10 doc/articles/amsi_biotissue/paper.tex
more work, adding a couple of initial figures which should be replaced once additional scaling results are available

@@
 
-\section{Introduction}\label{introduction}
+\section{Introduction and Motivation}\label{introduction}
 \label{single_scale_simulations}
@@
 \label{multi_scale_simulations}
-Introduction of influence from scales orders of magnitude removed from the primary scale can take place in several different ways.
-This influence can be incorporated during the mathematical development of the physical model itself.
+Introducing multi-scale influences on the scale of interest in a simulation can occur in several places during the development of a simulation.
+Most fundamentally, the influence can be incorporated during the mathematical development of the physical model itself.
 In this case the multi-scale features are incorporated into the numerical implementation automatically, as a result of their presence in the mathematical model.
 
-Multi-scale influences can also be incorporated in a more strictly numerical sense, combining physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
-Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact through passing key physical terms between scales.
-This scale-coupling will involve up-/down-scaling operations to account for differences in scale dimension, as well as possibly dimensionalizing terms computed in a non-dimensional domain.
+Multi-scale influences can also be incorporated in a numerical sense, combining established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
+Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact by passing key physical terms between scales.
+This operation, referred to as scale-coupling, involves up-/down-scaling operations to account for differences in scale dimensionality (and possibly dimensionalizing terms computed in a non-dimensional domain).
 
-In order for a numerically multi-scale approach to be effective the domain at the scale of interest (or engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude smaller would make the simulation computationally intractable.
-Thus some computationally reasonable subset of the engineering scale is assigned tertiary scales to simulate.
+In order for a numerically multi-scale approach to be effective the domain associated with the scale of interest (or engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude removed would make the simulation computationally intractable.
+Thus some computationally reasonable subset of the engineering scale is assigned tertiary scale(s) to simulate.
 The precise choice of location of the tertiary simulations is largely dependent on the specific numerical algorithms used to implement the engineering scale simulation.
 
-Often the subset(s) of the domain associated with the tertiary scale(s) are areas of the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation, in order to leverage the most power from the multi-scale approach.
+Often the subset of the domain associated with the tertiary scales are areas of the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation, in order to leverage the most utility from the multi-scale approach.
 
+Numerical multi-scale simulations fall broadly into two categories.
 In the concurrent multi-scale paradigm, all scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.
-
 In the sequential multi-scale paradigm, tertiary scales compute parameters which are converted to quantities of utility to a primary scale.
@@
 Amsi is designed to support the implementation and execution of dynamic numerically multi-scale simulations on massively-parallel HPC machines.
-Currently amsi provides interfaces for describing and managing the execution of individual scales in a multi-scale simulation, for defining and managing the scale-coupling communication required by such simulations, and for planning and enacting scale-sensitive load balancing operations on individual scales in a multi-scale system.
+Currently amsi provides interfaces for describing and managing the execution of individual scales in a multi-scale simulation, for defining, managing and enacting the scale-coupling communication required by such simulations, and for planning and enacting scale-sensitive load balancing operations on individual scales in a multi-scale system.
 
@@
 
-\subsection{Novelty} %rename
 Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{parker2006component} \cite{chopard2011framework} \cite{} and these frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah}.
 
-These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or high-level modelling language in order to take full advantage the services provided by the framework.
-Leveraging a component-based design is valuable from a software engineering perspective, but also introduces new complexities involved in defining the set of components and how they interoperate in the framework shifting some complexity up a level of abstraction.
+These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or specify the problem through a high-level modelling language in order to take full advantage the services provided by the framework.
+Leveraging a component-based design is valuable from a software engineering perspective, but also introduces new complexities involved in defining the set of components and how they interoperate in the framework, shifting some complexity up the abstraction hierarchy.
 The core features of amsi used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.
@@
 
-\label{amsi_load_balancing}
-If a local installation of the Zoltan \cite{ZoltanOverviewArticle2002} \cite{ZoltanIsorropiaOverview2012} load balancing library is available, the load balancing planning algorithms of Zoltan can be used underlying the scale-sensitive load balancing planning of amsi.
-A minimal set of load-balancing algorithms are provided with amsi.
-Since there is no general solution to the dynamic load balancing problem any set of provided load-balancing algorithms will prove insufficient for some use case.
-Thus amsi provides users with the ability to create and register their own load-balancing algorithms specific to the requirements of a given numerical implementation.
-
-Due to the necessity of retaining inter-scale linkages through the load-balancing processes, the actual data migration undertaken to implement a specific load balancing plan is typically conducted internally by the amsi system which is provided with buffers of data being load-balanced by each process on the appropriate scale.
-However, in order to maintain usability in more general situations, it is also possible to simply inform the amsi system that the current load-balancing plan has already been accomplished by user-implemented data migration.
-This functionality combined with the ability to use user-designed load-balance planning algorithms allows amsi to work with entirely external load-balancing libraries and algorithms, allowing scale-sensitive load-balancing to be used even when a specific multi-scale use case falls outside the bounds of capabilities built into the system.
-
 \subsection{Soft Tissue Simulation}\label{biotissue}
@@
 
-%Typically in the displacement-based finite element method, a constitutive model depending on some set of material properties is derived from a set of governing physical partial differential equations.
-In the case of the Biotissue multi-scale code, the constitutive model governing the systemic response to the boundary conditions is supplemented by the micro-scale force-balance simulation.
-The Newton-Raphson method is used to iterate to the solution for a particular loading state, and at each Newton iteration the global tangent stiffness matrix is assembled from elemental systems generated from each element in the engineering-scale mesh.
-In typically FEM the elemental system is derived from the governing material model used for the specific type of material being simulated; for the Biotissue code the necessary values are instead provided as
-
 Thus the simulation as a whole has two Newton-Raphson convergence operations occurring.
@@
 
-\section{Results}\label{results}
+\section{Results and Discussion}\label{results}
 The biotissue simulation was run on the AMOS BG/Q machine housed at the CCI facilities.
@@
 
-Weak scaling results...
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=2in]{ws_phil.eps}
+  \end{center}
+  \caption{\small weak scaling}
+  \label{biotissue_hierarchy}
+\end{figure
 
-Strong scaling results...
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=2in]{ss_phil.eps}
+  \end{center}
+  \caption{\small strong scaling}
+  \label{biotissue_hierarchy}
+\end{figure
+As the problem size is held constant with respect to the number of individual elements in the engineering scale domain discretization the simulation displays near optimal convergence to the macro-scale solve time alone, which is also reducing in time cost.
+The limiting factor in the speedup of the overall simulation is the potential speedup of the macro-scale simulation rather than the micro-scale simulation with fiber-only RVEs.
+As indicated in the weak scaling results this is related primarily to the macro-scale linear solve communication time increasing.
+At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU_dist solver \cite{li05superlu} \cite{lishao10superluilu} \cite{lidemmel03superludist} ).
+Using instead an iterative solution method to reach a desired level of accuracy may result in stronger scaling at macro-scale, but as yet this approach has been unneeded due to the preponderance of time spent computing micro-scale results in use cases.
+
+Many problems of interest will possess finite elements numbered in the tens to hundreds of millions of degrees of freedom, such that a suitable partitioning of the mesh to cores on the machine in terms of elements per partition (and directly the macro-scale time-to-solution) may require a substantial fraction of the total cores on the machine for macro-scale to execute in reasonable time.
+Amsi provides services for dynamically adding and removing individual scale-tasks to a specific scale, which can be utilized to limit RVE usage to only those locations in the multi-scale mesh most influential on the overall evolution of the simulation.
 
@@
 Scale-sensitive load-balancing operations have been developed for the biotissue multi-scale test problem.
-Unfortunately the problem does not exhibit load persistence in the current parallel task discretization, so an alternative parallel discretization such as overdecomposition with work-stealing to enforce more local balancing operations and occasional global balancing operations is being explored.
+Unfortunately the problem does not exhibit load persistence in the current parallel execution space task discretization, so an alternative parallel discretization such as overdecomposition with work-stealing to enforce more local balancing operations and occasional global balancing operations is being explored.
 
-Embedded fiber-matrix RVEs separately parallelized and incorporated into multi-scale model, used at most error-sensitive locations \cite{lake2012mechanics} \cite{zhang2013cross} \cite{zhang2013coupled}.
+Development of libraries concerned with higher levels of abstraction of the multi-scale problem will also be valuable in easing the development of new multi-scale simulations from the perspective of the application programmer.
+High level abstractions such as the Scale Separation Map in \cite{chopard2011framework} could provide mechanisms to provide initial configuration information to the amsi system during the initialization phase, as well as assist in the implementation of (or usage of existing library-provided) up-/down-scaling operations.
 
+An RVE simulation taking into account inter-fibril matrix material has been developed \cite{lake2012mechanics} \cite{zhang2013cross} \cite{zhang2013coupled}.
+This alternative RVE will be incorporated into the multi-scale model as an alternative to the fiber-only RVE, to be used in areas of the macro-scale domain determined to be most critical to the overall evolution of the global system, or where error accumulation through model inaccuracy is to be minimized.
+As this fiber-matrix RVE represents a more significant source of computational complexity, it is independently parallelized -- in contrast to the fiber-only RVE -- and will represent a compelling use case for the parallel execution space discretization and scale-coupling communication management capabilities of the amsi libraries.
 \bibliography{citations}

### 2016-01-10 doc/articles/amsi_biotissue/paper.tex
markup fix

@@
   \label{biotissue_hierarchy}
-\end{figure
+\end{figure}
 
@@
   \label{biotissue_hierarchy}
-\end{figure
+\end{figure}
+
 As the problem size is held constant with respect to the number of individual elements in the engineering scale domain discretization the simulation displays near optimal convergence to the macro-scale solve time alone, which is also reducing in time cost.

### 2016-01-10 doc/articles/amsi_biotissue/paper.tex
work on the paper

@@
 
+\section{Methods}
+
 \label{amsi_scales}
@@
 %
-num_macro = N 
-num_cores = N*128
-num_micro = num_cores - N
+num\_macro = N 
+num\_cores = N*128
+num\_micro = num\_cores - N
 %
@@
   \caption{\small weak scaling}
-  \label{biotissue_hierarchy}
+  \label{weak scaling}
 \end{figure}
@@
   \caption{\small strong scaling}
-  \label{biotissue_hierarchy}
+  \label{strong scaling}
 \end{figure}
-
 As the problem size is held constant with respect to the number of individual elements in the engineering scale domain discretization the simulation displays near optimal convergence to the macro-scale solve time alone, which is also reducing in time cost.
@@
 As indicated in the weak scaling results this is related primarily to the macro-scale linear solve communication time increasing.
-At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU_dist solver \cite{li05superlu} \cite{lishao10superluilu} \cite{lidemmel03superludist} ).
+At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU\_dist solver \cite{li05superlu} \cite{lishao10superluilu} \cite{lidemmel03superludist} ).
 Using instead an iterative solution method to reach a desired level of accuracy may result in stronger scaling at macro-scale, but as yet this approach has been unneeded due to the preponderance of time spent computing micro-scale results in use cases.

### 2016-01-10 doc/articles/amsi_biotissue/paper.tex
working on paper

@@
 \usepackage{graphicx}
+\usepackage{listings}
 \usepackage{mathtools}
@@
 A configuration file is supplied during command invocation containing the initial configuration of the process sets and their assignment to individual scales, as well as declaring the coupling between scales in the simulation.
-The deceleration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.
+The declaration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.
+A simple configuration file can be seen in listing \ref{amsi config file} which declares the existence of two scales, macro and micro\_fo, and assigns them 64 and 8128 processes respectively (for a total of $2^13$ processes).
+The file also specifies the scale-coupling relations, with macro coupled to micro and micro coupled to macro.
+Note that the order of specification is not relevant to the operation of amsi.
+%
+\begin{lstlisting}
+  macro 64
+  micro_fo 8128
+  #
+  macro micro_fo
+  micro_fo macro_fo
+\end{lstlisting}
+\label{amsi config file}
+%
 
@@
 
-Deceleration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling deceleration during simulation initialization is used to instantiate data structures specific to scale-coupling operations.
+Declaration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations.
 This limits memory overhead by not allocating space for unneeded coupling operations.
@@
 This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared.
-For brevity this may be written as A~B for A is coupled to B.
 
@@
 At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in a future release to automate application of such operations during scale-coupling.
+The current implementation includes on blocking versions of the scale-coupling communication operations.
+Asynchronous versions have been designed but are not currently being implemented due to lack of immediate need for the primary test amsi use case.
 
@@
 \begin{align}
-\nabla \cdot \pmb{\sigma} =& \; 0 \text{ in } \Omega \label{momentum_balance} \\
+\nabla \cdot \pmb{\sigma} =& \; 0 \text{ in } \Omega \label{momentum_balance} \nonumber \\
 \pmb{\sigma} \cdot \pmb{n} =& \; \pmb{t} \text{ on } \partial \Omega \nonumber
@@
 Specific quantities needed to compute the elemental tangent stiffness matrices and force vectors are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}.
-
+%
 \begin{figure}
@@
 \end{figure}
+%
 
@@
 
-Thus the simulation as a whole has two Newton-Raphson convergence operations occurring.
-During each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
+The simulation two nested Newton-Raphson operations occurring; during each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
 
@@
 %
-num\_macro = N 
-num\_cores = N*128
-num\_micro = num\_cores - N
+\begin{lstlisting}[Language=C]
+  num_cores = 128 * num_macro;
+  num_micro = num_core - num_macro;
+\end{lstliting}
+\label{macro micro ratio}
 %
@@
 
+%
 \begin{figure}
@@
 \end{figure}
+%
+Keeping the work per process approximately constant (1k elements per macro-scale process, with the macro:micro ratio as specified in listing \ref{macro micro ratio}) it can be seen that the time to compute every RVE in the system stays relatively consistent, while the time to solve the macro-scale linear system increases, causing a modest increase in the overall time per macro-scale Newton iteration.
 
+%
 \begin{figure}
@@
 \end{figure}
+%
 As the problem size is held constant with respect to the number of individual elements in the engineering scale domain discretization the simulation displays near optimal convergence to the macro-scale solve time alone, which is also reducing in time cost.

### 2016-01-10 doc/articles/amsi_biotissue/paper.tex
more paper work

@@
 \DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
+
+\lstset{basicstyle=\small, 
+        stringstyle=\ttfamily, 
+        frame=top, frame=bottom,
+        captionpos=t}
 
@@
 The declaration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.
-A simple configuration file can be seen in listing \ref{amsi config file} which declares the existence of two scales, macro and micro\_fo, and assigns them 64 and 8128 processes respectively (for a total of $2^13$ processes).
+A simple configuration file can be seen in listing \ref{amsi_config_file} which declares the existence of two scales, macro and micro\_fo, and tells the amsi intialization procedures to assigne them process sets containig 64 and 8128 processes, respectively (for a total of $2^{13}$ processes).
 The file also specifies the scale-coupling relations, with macro coupled to micro and micro coupled to macro.
@@
 %
-\begin{lstlisting}
+\begin{lstlisting}[caption={amsi\_config file},label=amsi_config_file]
   macro 64
@@
 \end{lstlisting}
-\label{amsi config file}
 %
+Typically creation of this file will be automated and added to the shell script used to invoke the simulation, so ensure that process counts passed in throught the configuration file are valid.
+Alternative specification mechanisms such as ratio specification (equivalently percentage specification) are being considered for future use to prevent the possbility of misallocation of processes (the user not specifying assignment for all processes available to the simulation).
 
@@
 This results in each micro-scale node being displaced.
-Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation.
+Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation, resulting in a high coordination number at each node and consequently a numerically stable fiber network.
 The micro-scale boundary condition necessitates that the force exerted by each fiber network node located on the boundary of the RVE be zero.
@@
 The ratio of macro-scale to micro-scale assigned processes was held near 1:128, though the total core count was kept as an integer power of 2 in order for the SLURM job scheduler to submit the jobs to the compute nodes in a timely manner.
-The precise calculation used to determine macro and micro ratio is given by
+The precise calculation used to determine macro and micro process counts is given by
 %
-\begin{lstlisting}[Language=C]
-  num_cores = 128 * num_macro;
-  num_micro = num_core - num_macro;
-\end{lstliting}
+\begin{lstlisting}[language=Bash,caption={process assignment calculation}]
+  MICRO_RATIO=128
+  NP=$(( $NUM_MACRO * $MICRO_RATIO))
+  NUM_MICRO=$(( $NP - $NUM_MACRO))
+\end{lstlisting}
 \label{macro micro ratio}

### 2016-01-10 doc/articles/amsi_biotissue/paper.tex
more paper work

@@
 The declaration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.
+
+Process sets are mathematical sets of MPI ranks implemented so as to take advantage of any mathematical conveniences to minimize explicit storage whenever possible.
+Process sets are required to be non-overlapping in order for amsi control mechanisms to operate correctly.
+Ongoing developments are targeted at removing this requirement to allow a richer set of dynamic control options and increase overall utilization in the case of coupled scales blocking on one another.
+
+Scales are an abstraction grouping a set of standalone scale-tasks operating on a single scale, the set may be as minimal as a single scale-task as in the case of a primary scale with many contributing tertiary scales.
+The set may also contain many individual scale-tasks which all operate within a single scale, such as when a primary scale-task has an associated tertiary scale which constitutes many individual simulations, as in the biotissue example case discussed in section \ref{biotissue}.
+Precisely what defines a scale-task is left up to the user, the intent is for each scale-task to represent a standalone single-scale simulation incorporated into the multi-scale system.
+
+Declaration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations.
+This limits memory overhead by not allocating space for unneeded coupling operations.
+
+Scales are declared to be related if some quantities of interest (such as physical tensor field values) are transmitted between them for scale-coupling.
+This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared.
+
 A simple configuration file can be seen in listing \ref{amsi_config_file} which declares the existence of two scales, macro and micro\_fo, and tells the amsi intialization procedures to assigne them process sets containig 64 and 8128 processes, respectively (for a total of $2^{13}$ processes).
@@
 
-Process sets are mathematical sets of MPI ranks implemented so as to take advantage of any mathematical conveniences to minimize explicit storage whenever possible.
-Process sets are required to be non-overlapping in order for amsi control mechanisms to operate correctly.
-Ongoing developments are targeted at removing this requirement to allow a richer set of dynamic control options and increase overall utilization in the case of coupled scales blocking on one another.
-
-Declaration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations.
-This limits memory overhead by not allocating space for unneeded coupling operations.
-
-Scales are declared to be related if some quantities of interest (such as physical tensor field values) are transmitted between them for scale-coupling.
-This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared.
+The assignment of processes to a process set is currently static, and processes cannot be reassigned during a simulation.
+Additionally the assignment of a scale to a process set is static as well.
+However, scale-tasks are capable of being dynamically created and destroyed during runtime.
+Additionally, scale-tasks associated with specific subsets of a process set (either a single process or several processes) may be migrated in part or in whole.
+Migrating a scale-task in part amounts to using load balancing facilities on the associated processes to redistribute all data scale-task from a scale-task to a subset of the previously associated processes, allowing the freed processes to be assigned data from other scale-tasks.
+Migrating a scale task in whole constitutes using load balancing operations on the process set associated with the entire scale to move all pertinent data between subset of the managed process set.
+In the current implementation these operations are both collective on the process set associated with the managed scale.
 

### 2016-01-11 doc/articles/amsi_biotissue/paper.tex
yet again, more paper work

@@
 
+\usepackage{amsfonts}
 \usepackage{graphicx}
@@
 
-Declaration of a coupling relation between two scales is required if any communication between the scales is to occur during the simulation (through the amsi libraries), as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations.
+A scale coupling is a relation denoting that scale-coupling communication will take place between two scales at some point during the simulation.
+Declaration of a coupling relation between two scales is required for communicating scales, as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations.
 This limits memory overhead by not allocating space for unneeded coupling operations.
+This coupling relationship is asymmetric, thus if scale A is related to scale B (ie.
+A produces terms required for B to operate), scale B is not implicitly related to scale A, though this may be separately declared.
 
-Scales are declared to be related if some quantities of interest (such as physical tensor field values) are transmitted between them for scale-coupling.
-This coupling relationship is assumed to be asymmetric, thus if scale A is related to scale B, scale B is not implicitly related to scale A, though this may be separately declared.
-
-A simple configuration file can be seen in listing \ref{amsi_config_file} which declares the existence of two scales, macro and micro\_fo, and tells the amsi intialization procedures to assigne them process sets containig 64 and 8128 processes, respectively (for a total of $2^{13}$ processes).
+A simple configuration file can be seen in listing \ref{amsi_config_file} which declares the existence of two scales, macro and micro\_fo, and tells the amsi intialization procedures to assign them process sets containig 64 and 8128 processes, respectively (for a total of $2^{13}$ processes).
 The file also specifies the scale-coupling relations, with macro coupled to micro and micro coupled to macro.
@@
 %
-Typically creation of this file will be automated and added to the shell script used to invoke the simulation, so ensure that process counts passed in throught the configuration file are valid.
-Alternative specification mechanisms such as ratio specification (equivalently percentage specification) are being considered for future use to prevent the possbility of misallocation of processes (the user not specifying assignment for all processes available to the simulation).
 
-The assignment of processes to a process set is currently static, and processes cannot be reassigned during a simulation.
+Dynamically creating the required configuration file as part of a shell script invoking the simulation is straightforward.
+Alternative specification mechanisms such as ratio specification (equivalently percentage specification) are being considered for future implementation to prevent the possibility of misallocating processes (the user not specifying assignment for all processes available to the simulation, or attempting to overallocate processes).
+
+The assignment of processes to a process set is static, and processes cannot be reassigned during a simulation.
 Additionally the assignment of a scale to a process set is static as well.
-However, scale-tasks are capable of being dynamically created and destroyed during runtime.
+The application programmer may provide capabilities for scale-tasks to be dynamically created and destroyed at runtime, amsi provides functions to inform coupled scales of such dynamic modifications.
+Since scale-tasks are an abstraction the amsi systems cannot provide direct operations on scale-tasks, but provide mechanisms for the application to inform amsi of such operations when they occur.
+
 Additionally, scale-tasks associated with specific subsets of a process set (either a single process or several processes) may be migrated in part or in whole.
@@
 Migrating a scale task in whole constitutes using load balancing operations on the process set associated with the entire scale to move all pertinent data between subset of the managed process set.
-In the current implementation these operations are both collective on the process set associated with the managed scale.
+Amsi provides operations to enact load-balancing operations from plans such as those generated by Zoltan algorithms.
+In the current implementation these operations are collective on the process set associated with the managed scale.
 
@@
 Any scale is capable of informing the process set associated with a coupled scale of the addition or deletion of individual scale-tasks.
-At present it is dependent on the developer to implement the capability of a scale to accept additional scale-tasks, though algorithms and hooks for developer-designed distribution algorithms are supplied to ease the implementation.
+At present it is dependent on the developer to implement the capability of a scale to accept additional scale-tasks.
 
@@
 
+Scale-linking quantities are dynamically registered with the amsi system using a data distribution defined on a scale by the scale-tasks operating at that scale.
+Production of a communication pattern is required prior to executing a scale-coupling communication, and the pattern must be reconciled (synchronized across the coupling relationship) prior to being used to guide a coupling communication operation (and when the communication pattern is altered by the scale providing the coupling data).
+Several simple communication pattern production algorithms are provided by amsi, but these are relatively trivial due to the nature of scale-coupling communications and the general nature of the amsi libraries.
+Hooks are provided for users to implement algorithms for the construction of communication patterns tailored to the specifics of their problem.
+
+A communication pattern based on the relationship of scale $A$ coupled to scale $B$ is thought of most simply as a matrix $C \in \mathbb{N}^{|A|\times|B|}$ (where $|X|$ is the cardinality of the process set currently assigned to scale X).
+The implementation of the communication pattern uses a sparse format in order to reduce the memory impact as the simulation scales.
+The quantity $C_{ij}$ represents the quantity of scale-coupling terms (of arbitrary type) to be sent from process $i$ in the process set assigned to $A$ to process $j$ in the process set assigned to $B$.
+
 All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer.
-At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in a future release to automate application of such operations during scale-coupling.
-The current implementation includes on blocking versions of the scale-coupling communication operations.
-Asynchronous versions have been designed but are not currently being implemented due to lack of immediate need for the primary test amsi use case.
+Usage only of implicit ordering was decided upon in order to reduce the memory impact as the simulation scales, as any meta-data which scales with the simulation will necessarily pose a memory issue once scaling hits some threshold.
+
+At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in future releases to automate application of such operations during scale-coupling.
+The current implementation includes blocking versions of the scale-coupling communication operations.
+Asynchronous versions have been designed but are not currently implemented due to lack of immediate need in the primary amsi use case.
 

### 2016-01-13 amsi_control/ControlService.h
changing amsiMPI.cc to accommodate a more general method of calculating distributed values, also work on simplifying the commpattern and migration operations, though that work should be exteremely minimal

@@ -285,7 +285,7 @@ namespace amsi {
 
 	  // iterate over the recvers for this rank
 	  std::vector<int> send_to;
-	  send_pattern->GetSendTo(task_rank,send_to);
+	  send_pattern->getSentTo(task_rank,send_to);
 
 	  size_t offset = 0;
 #         ifndef CORE
@@ -323,7 +323,7 @@ namespace amsi {
 	  // assuming that the commpattern is reconciled here...
 	  CommPattern * pattern = comm_man->CommPattern_Get(rdd_id);
 	  std::vector<int> recv_from;
-	  pattern->GetRecvFrom(task_rank,recv_from);
+	  pattern->getRecvedFrom(task_rank,recv_from);
 	
 #         ifdef CORE
           PCU_Comm_Send();

### 2016-01-13 amsi_meta/AddData_LoadBalancing.cc

@@ -25,8 +25,8 @@ namespace amsi {
   void CommPattern_LoadBalance_Spread(CommPattern * pattern, CommPattern * init, int * num)
   {
 
-    int s1 = pattern->getSendTaskSize();
-    int s2 = pattern->getRecvTaskSize();
+    int s1 = pattern->getNumSenders();
+    int s2 = pattern->getNumRecvers();
     int total_data = 0;
     for(int ii = 0; ii < s1; ii++)
     {
@@ -75,8 +75,8 @@ namespace amsi {
   void CommPattern_LoadBalance_LeastFirst(CommPattern * pattern, CommPattern * init, int * num)
   {
 
-    int s1 = pattern->getSendTaskSize();
-    int s2 = pattern->getRecvTaskSize();
+    int s1 = pattern->getNumSenders();
+    int s2 = pattern->getNumRecvers();
     int tempsum;
     int sums[s2];
 
@@ -199,7 +199,7 @@ namespace amsi {
     assert(init);
     assert(num);
 
-    int s2 = pattern->getRecvTaskSize();
+    int s2 = pattern->getNumRecvers();
     srand(time(NULL));
     for(int ii=0;ii<num[rank];ii++)
     {

### 2016-01-13 amsi_meta/Assemblable.h

@@ -9,8 +9,9 @@ namespace amsi {
    {
    public:
      virtual int Assemble(MPI_Comm,int) = 0;
-     virtual bool isAssembled() const = 0;
-   private:
+     bool isAssembled() {return assembled;}
+   protected:
+     void unassembled() {assembled = false;}
      bool assembled;
    };
 
@@ -18,8 +19,9 @@ namespace amsi {
    {
    public:
      virtual void Reconcile() = 0;
-     virtual bool isReconciled() const = 0;
-   private:
+     bool isReconciled() {return reconciled;}
+   protected:
+     void unreconciled() {reconciled = false;}
      bool reconciled;
    };
 

### 2016-01-13 amsi_util/amsiMPI.cc

@@ -1,10 +1,5 @@
 #include "amsiMPI.h"
-
-double distributedAverage(double partial_sum, int partial_count, MPI_Comm cm)
-{
-  double sum = 0.0;
-  MPI_Allreduce(&partial_sum,&sum,1,MPI_DOUBLE,MPI_SUM,cm);
-  int count = 0;
-  MPI_Allreduce(&partial_count,&count,1,MPI_INTEGER,MPI_SUM,cm);
-  return sum / count;
-}
+template <>
+MPI_Datatype mpi_type<double>() {return MPI_DOUBLE;}
+template <>
+MPI_Datatype mpi_type<int>() {return MPI_INT;}

### 2016-01-13 amsi_util/amsiMPI.h

@@ -1,16 +1,33 @@
 #ifndef MPI_T_H_
 #define MPI_T_H_
-
 #include <mpi.h>
-
 #include <cstring> //memcpy
 #include <iostream>
-
 extern MPI_Comm AMSI_COMM_WORLD;
 extern MPI_Comm AMSI_COMM_LOCAL;
-
-double distributedAverage(double partial_sum, int partial_count, MPI_Comm cm);
-
+template <typename T>
+MPI_Datatype mpi_type() {return MPI_BYTE;}
+template<typename T>
+T comm_sum(T v, MPI_Comm cm = AMSI_COMM_LOCAL)
+{
+  T sm = 0;
+  MPI_Allreduce(&v,&sm,1,mpi_type<T>(),MPI_SUM,cm);
+  return sm;
+}
+template <typename T>
+T comm_min(T v, MPI_Comm cm = AMSI_COMM_LOCAL)
+{
+  T mn = 0;
+  MPI_Allreduce(&v,&mn,1,mpi_type<T>(),MPI_MIN,cm);
+  return mn;
+}
+template <typename T>
+T comm_max(T v, MPI_Comm cm = AMSI_COMM_LOCAL)
+{
+  T mx = 0;
+  MPI_Allreduce(&v,&mx,1,mpi_type<T>(),MPI_MAX,cm);
+  return mx;
+}
 template <typename T>
 void t_ssend(T & msg, MPI_Datatype type, int to)
 {

### 2016-01-13 scripts/config.sh
accomodating new method of adding petsc to the build

@@ -49,16 +49,19 @@ if [ "$HOSTNAME" == "q.ccni.rpi.edu" ]; then
     ..
 
 else
+    module load openmpi-1.3.3
+    module load simmetrix/simModSuite
+    module load $DEVROOT/petsc/petsc-3.6.3/ompi133-debug/lib/petsc/conf/modules/petsc/3.6.3
 
-  cmake \
-    -DCMAKE_BUILD_TYPE=$BUILD_TYPE \
-    -DBUILD_TESTS=ON \
-    -DCMAKE_INSTALL_PREFIX=$DEVROOT/install/amsi/sim/openmpi-1.3.3/ \
-    -DCMAKE_C_COMPILER="mpicc" \
-    -DCMAKE_CXX_COMPILER="mpicxx" \
-    -DCORE_DIR=$DEVROOT/install/core-sim/openmpi-1.3.3 \
-    -DSIM_MPI=openmpi14 \
-    ..
+    cmake \
+	-DCMAKE_BUILD_TYPE=$BUILD_TYPE \
+	-DBUILD_TESTS=ON \
+	-DCMAKE_INSTALL_PREFIX=$DEVROOT/install/amsi/sim/openmpi-1.3.3/ \
+	-DCMAKE_C_COMPILER="mpicc" \
+	-DCMAKE_CXX_COMPILER="mpicxx" \
+	-DCORE_DIR=$DEVROOT/install/core-sim/openmpi-1.3.3 \
+	-DSIM_MPI=openmpi14 \
+	..
 
 #    -DSIMMETRIX_LIB_DIR=$DEVROOT/simPartitionWrapper/PartitionWrapper/lib \
 #    --debug-output \

### 2016-01-14 doc/articles/amsi_biotissue/paper.tex
im a moron

@@
 Both strong and weak scaling studies were conducted up to an eighth of the total machine core count, or 16384 cores.
-The ratio of macro-scale to micro-scale assigned processes was held near 1:128, though the total core count was kept as an integer power of 2 in order for the SLURM job scheduler to submit the jobs to the compute nodes in a timely manner.
+The ratio of macro-scale to micro-scale assigned processes was held at 1:127, and the total core count was kept as an integer power of 2 in order for the SLURM job scheduler to submit the jobs to the compute nodes in a timely manner.
 The precise calculation used to determine macro and micro process counts is given by
@@
 \begin{lstlisting}[language=Bash,caption={process assignment calculation}]
-  MICRO_RATIO=128
-  NP=$(( $NUM_MACRO * $MICRO_RATIO))
+  NP=$(( $NUM_MACRO * 128))
   NUM_MICRO=$(( $NP - $NUM_MACRO))
@@
 %
-thus as the number of cores assigned to the macro-scale computation increases, the ratio approaches 1:128.
 
@@
 %
-Keeping the work per process approximately constant (1k elements per macro-scale process, with the macro:micro ratio as specified in listing \ref{macro micro ratio}) it can be seen that the time to compute every RVE in the system stays relatively consistent, while the time to solve the macro-scale linear system increases, causing a modest increase in the overall time per macro-scale Newton iteration.
+Keeping the work per process approximately constant (1k elements per macro-scale process, with the 1:127 macro-micro ratio as specified in listing \ref{macro micro ratio}) it can be seen that the time to compute every RVE in the system stays relatively consistent, while the time to solve the macro-scale linear system increases, causing a modest increase in the overall time per macro-scale Newton iteration.
 

### 2016-01-15 amsi_interface/NonFiniteElement.cc
start of day file revisions to correct formatting and naming conventions

@@ -1,181 +1,161 @@
 #include "NonFiniteElement.h"
-
-namespace amsi {
-  namespace Analysis {
-
-  void LinearStiffnessK0(int num_nodes,
-			 apf::DynamicMatrix & B0, 
-			 apf::Matrix<6,6> & hookes,
-			 apf::DynamicMatrix & K0)
+namespace amsi
 {
-  apf::DynamicMatrix B0T;
-  apf::transpose(B0,B0T);
-
-  apf::DynamicMatrix T(6,6);
-  T *= 0.0;
-
-  for(int ii = 0; ii < num_nodes; ii++)
-    for(int jj = 0; jj < 6; jj++)
-      for(int kk = 0; kk < 6; kk++)
-	T(ii,jj) += B0T(ii,kk) * hookes[kk][jj];
-  
-  apf::multiply(T,B0,K0);
-}
-
-  void NonLinearStiffness(int num_nodes,
-			  apf::DynamicMatrix & BNL, 
-			  apf::Matrix3x3 & pk2, 
-			  apf::DynamicMatrix & KNL)
+  namespace Analysis
   {
-    apf::DynamicMatrix BNLT; //12x9 num_nodes * 3 (dofs_per_node) x 9
-    apf::transpose(BNL,BNLT); 
-    
-    apf::DynamicMatrix expanded_pk2(9,9);
-    expanded_pk2 *= 0.0;
-    
-    for(int ii = 0; ii < 3; ii++)
-      for(int jj = 0; jj < 3; jj++)
-      {
-	expanded_pk2(ii,jj)     = pk2[ii][jj];
-	expanded_pk2(ii+3,jj+3) = pk2[ii][jj];
-	expanded_pk2(ii+6,jj+6) = pk2[ii][jj];
-      }
-
-    KNL.setSize(BNLT.getRows(), BNL.getColumns());
-    apf::DynamicMatrix T;
-    apf::multiply(BNLT,expanded_pk2,T);
-    apf::multiply(T,BNL,KNL);
-  }
-  
-  void LinearStrainDispB0(int num_nodes,
-			  apf::NewArray<apf::Vector3> & shape_deriv, 
-			  apf::DynamicMatrix & B) 
-{
-  B.setSize(6,3*num_nodes);
-  B *= 0.0;
-  
-  for(int ii = 0; ii < num_nodes; ii++)
-  {
[…]

### 2016-01-15 amsi_interface/NonFiniteElement.h

@@ -1,48 +1,37 @@
 #ifndef H_NONLINFINITEELEMENT
 #define H_NONLINFINITEELEMENT
-
 #include "apf.h"
 #include "apfDynamicMatrix.h"
-
 namespace amsi {
   namespace Analysis {
-
-  void LinearStiffnessK0(int num_nodes,
-			 apf::DynamicMatrix & B0, 
-			 apf::Matrix<6,6> & hookes, 
-			 apf::DynamicMatrix & K0);
-
-  void NonLinearStiffness(int num_nodes,
-			  apf::DynamicMatrix & BNL, 
-			  apf::Matrix3x3 & pk2, 
-			  apf::DynamicMatrix & KNL);
-
-  void LinearStrainDisp(int num_nodes,
-			apf::NewArray<apf::Vector3> & shape_deriv, 
-			apf::NewArray<apf::Vector3> & node_disp, 
-			apf::DynamicMatrix & BL);
-
-  void LinearStrainDispB0(int num_nodes,
-			  apf::NewArray<apf::Vector3> & shape_deriv, 
-			  apf::DynamicMatrix & B);
-
-  void LinearStrainDispB1(int num_nodes, 
-			  apf::NewArray<apf::Vector3> & shape_deriv, 
-			  apf::NewArray<apf::Vector3> & node_disp,
-			  apf::DynamicMatrix & B); 
-
-  void ComputeGreenStrain(apf::DynamicMatrix & H, 
-			  apf::DynamicVector & y, 
-			  apf::Matrix3x3 E);
-
-  void ComputePK2Stress(apf::DynamicMatrix & H, 
-			apf::DynamicVector & S, 
-			apf::DynamicMatrix & C);
-
-  void NonLinearStrainDisp(int num_nodes,
-			   apf::NewArray<apf::Vector3> & shape_deriv, 
-			   apf::DynamicMatrix & B);
-
+  void linearStiffnessK0(int num_nodes,
+                         apf::DynamicMatrix & B0,
+                         apf::Matrix<6,6> & hookes,
+                         apf::DynamicMatrix & K0);
+  void nonLinearStiffness(int num_nodes,
+                          apf::DynamicMatrix & BNL,
+                          apf::Matrix3x3 & pk2,
+                          apf::DynamicMatrix & KNL);
+  void linearStrainDisp(int num_nodes,
+                        apf::NewArray<apf::Vector3> & shape_deriv,
[…]

### 2016-01-15 doc/articles/amsi_biotissue/paper.tex
fixed usage of refactored functions from start-of-day commit, reworked commpattern interface and usage in key functions Communication and Reconcile, needs to be tested more, which will be done in the next few hours

@@
 
-\section{Introduction and Motivation}\label{introduction}
+\section{Introduction}\label{introduction}
 \label{single_scale_simulations}
@@
 
-\subsection{Adaptive Multi-scale Simulation Infrastructure}\label{amsi}
+\section{Adaptive Multi-scale Simulation Infrastructure}\label{amsi}
 The Adaptive Multi-scale Simulation Infrastructure (amsi) is a set of libraries and tools developed at the Scientific Computation Research Center (SCOREC) at Rensselaer Polytechnic Institute.
@@
 
-\subsection{Soft Tissue Simulation}\label{biotissue}
+\section{Soft Tissue Simulation}\label{biotissue}
 The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation.

### 2016-01-15 amsi_control/ControlService.cc

@@ -1,115 +1,104 @@
 #include "ControlService.h"
-
-// amsi_meta
 #include "Assemblable.h"
 #include "PlanMigration.h"
 #include "DataDistribution.h"
-
-
 #ifdef ZOLTAN
 #include <zoltan.h>
 #endif
-
 #include <fstream>
 #include <algorithm>
+namespace amsi
+{
+  ControlService * ControlService::instance = NULL;
+  size_t ControlService::rel_dist_id_gen = 0;
+  ControlService::ControlService()
+    : comm_man()
+    , task_man()
+    , rdd_map()
+  {}
+  /// @brief Set the CommunicationManager to be used by the AMSI ControlService
+  /// @param cm A pointer to the CommunicationManager
+  void ControlService::SetCommunicationManager(CommunicationManager * cm)
+  {
+    comm_man = cm;
+  }
+  /// @brief Set the TaskManager to be used by the AMSI ControlService
+  /// @param tm A pointer to the TaskManager
+  void ControlService::SetTaskManager(TaskManager * tm)
+  {
+    task_man = tm;
+  }
+  /// @brief Define a CommRelation between two named Tasks
+  /// @param nm1 A string containing the name of the sending task
+  /// @param nm2 A string containing the name of the recving task
+  /// @return size_t An identifier for the CommRelation or 0 if the relation could
+  ///                not be created (typically if one of the Tasks doesn't exist).
+  size_t ControlService::CommRelation_Define(const std::string & nm1,
+                                             const std::string & nm2)
+  {
+    size_t result = 0;
+    size_t t1_id = task_man->getTaskID(nm1);
+    size_t t2_id = task_man->getTaskID(nm2);
 
-namespace amsi {
+    if(t1_id != 0 && t2_id != 0)
+      result = comm_man->defineRelation(t1_id,t2_id);
 
-    ControlService * ControlService::instance = NULL;
-    size_t ControlService::rel_dist_id_gen = 0;
+    return result;
+  }
 
-    //private
-    ControlService::ControlService() :
-      comm_man(),
-      task_man(),
-      rdd_map()
-    {}
+  /// @brief Retrieve the ID of a CommRelation
+  /// @param nm1 A string containing the name of the sending task
+  /// @param nm2 A string containing the name of the recving task
+  /// @return size_t An identifier for the CommRelation or 0 if the relation could
[…]

### 2016-01-15 amsi_control/ControlService.h

@@ -14,7 +14,7 @@
 #include <string>
 namespace amsi
 {
-    /// Class decleration of the ControlService class (set to be renamed as MultiscaleManager or ScaleCouplingManager or...), 
+    /// Class decleration of the ControlService class (set to be renamed as MultiscaleManager or ScaleCouplingManager or...),
     ///  used as the top-level interface to the task managing portions of the AMSI infrastructure.
     class ControlService
     {
@@ -30,56 +30,56 @@ namespace amsi
       int Execute(int& argc,char **& argv);
 
       size_t CommRelation_Define(const std::string & nm1,
-				 const std::string & nm2);
+                                 const std::string & nm2);
 
       size_t CommRelation_GetID(const std::string & nm1,
                                 const std::string & nm2);
 
       size_t CreateCommPattern(const std::string & t1_dd,
-			       const std::string & t1,
-			       const std::string & t2);
+                               const std::string & t1,
+                               const std::string & t2);
 
       void getPatternInfo(size_t rdd_id,int &,int &,CommPattern *&);
 
       // Remove and add data from comm pattern
       void RemoveData(size_t rdd_id, std::vector<int> & data);
-      
+
       template <typename D, template <typename T, typename All = std::allocator<T> > class Container>
       size_t AddData(size_t rdd_id,
                      Container<D> & objects,
-		     std::vector<int> & data,
+                     std::vector<int> & data,
                      int option = 0,
                      void (*userFunc)() = NULL );
 
       // Migration functions
       void planMigration(std::vector<int> & migration_indices,
-			 size_t rdd_id,
+                         size_t rdd_id,
                          int option = 0,
                          void (*userFunc)() = NULL);
-      
+
       template <typename D, template<typename T, typename All = std::allocator<T> > class Container>
       void migration(std::vector< std::vector<char> > & migration_data,
                      size_t rdd_id,
                      Container<D> & objects);
 
-      
-      
+
+
[…]

### 2016-01-15 amsi_interface/NonFiniteElement.cc

@@ -131,8 +131,8 @@ namespace amsi
                           apf::DynamicMatrix & BL)
     {
       apf::DynamicMatrix B0, B1;
-      LinearStrainDispB0(num_nodes,shape_deriv,B0);
-      LinearStrainDispB1(num_nodes,shape_deriv,node_disp,B1);
+      linearStrainDispB0(num_nodes,shape_deriv,B0);
+      linearStrainDispB1(num_nodes,shape_deriv,node_disp,B1);
       BL.setSize(B1.getRows(),B1.getColumns());
       BL.zero();
       BL += B0;

### 2016-01-15 amsi_interface/NonLinearElasticIntegrator.cc

@@ -1,32 +1,28 @@
 #include "NonLinearElasticIntegrator.h"
-
 #include "apfFunctions.h"
 #include "ConvenienceFunctions.h"
 #include "NonFiniteElement.h"
-
 #include <apfMesh.h>
 #include <apfShape.h>
-
 namespace amsi {
   namespace Analysis {
-
     NonLinearElasticIntegrator::NonLinearElasticIntegrator(apf::Field * field,
-							   int o,
-							   double pr,
-							   double sm) :
-      apf::Integrator(o),
-      ke(),
-      fe(),
-      e(NULL),
-      f(field),
-      fs(NULL),
-      me(NULL),
-      node_values(),
-      poisson_ratio(pr),
-      shear_modulus(sm),
-      num_element_eqs(0),
-      num_field_components(0),
-      num_element_nodes(0)
+                                                           int o,
+                                                           double pr,
+                                                           double sm)
+      : apf::Integrator(o)
+      , ke()
+      , fe()
+      , e(NULL)
+      , f(field)
+      , fs(NULL)
+      , me(NULL)
+      , node_values()
+      , poisson_ratio(pr)
+      , shear_modulus(sm)
+      , num_element_eqs(0)
+      , num_field_components(0)
+      , num_element_nodes(0)
     {
       fs = apf::getShape(f);
       num_field_components = apf::countComponents(f);
@@ -46,14 +42,14 @@ namespace amsi {
 
       if(reallocate)
       {
-	ke.setSize(num_element_eqs,num_element_eqs);
-	fe.setSize(num_element_eqs);
+        ke.setSize(num_element_eqs,num_element_eqs);
+        fe.setSize(num_element_eqs);
       }
-      
+
       ke *= 0.0;
       fe.zero();
     }
-    
+
     void NonLinearElasticIntegrator::atPoint(apf::Vector3 const &p, double w, double dV)
     {
       apf::Matrix3x3 J;
@@ -76,27 +72,27 @@ namespace amsi {
       PK2Stress(C,dV,poisson_ratio,shear_modulus,pk2);
 
       apf::DynamicMatrix BL;
-      LinearStrainDisp(num_element_nodes,
-		       shape_derivs,
-		       node_values,
-		       BL);
+      linearStrainDisp(num_element_nodes,
+                       shape_derivs,
+                       node_values,
+                       BL);
 
[…]

### 2016-01-15 amsi_meta/CommPattern.cc

@@ -1,13 +1,14 @@
 #include "CommPattern.h"
+#include <cassert>
 #include <cstring> // for memset
 #include <iomanip>
 #include <iostream>
 namespace amsi
 {
   std::pair<int,int> coupledInfoByIndex(CommPattern * cp,
-					Role role,
-					int rank,
-					int index)
+                                        Role role,
+                                        int rank,
+                                        int index)
   {
     std::pair<int,int> result;
     int remainder = 0;
@@ -19,19 +20,19 @@ namespace amsi
       int count = 0;
       for(int ii = 0; ii < num_recvers; ii++)
       {
-	int other_count = (*cp)(rank,ii);
-	if(count + other_count >= index)
-	{
-	  remainder = index - count;
-	  coupled_rank = ii;
-	  break;
-	}
-	else
-	  count += other_count;
+        int other_count = (*cp)(rank,ii);
+        if(count + other_count >= index)
+        {
+          remainder = index - count;
+          coupled_rank = ii;
+          break;
+        }
+        else
+          count += other_count;
       }
       count = 0;
       for(int ii = 0; ii < rank; ii++)
-	count += (*cp)(ii,coupled_rank);
+        count += (*cp)(ii,coupled_rank);
       result.first = coupled_rank;
       result.second = count + remainder;
     }
@@ -40,126 +41,147 @@ namespace amsi
       int count = 0;
       for(int ii = 0; ii < num_senders; ii++)
       {
-	int other_count = (*cp)(ii,rank);
-	if(count + other_count >= index)
-	{
-	  remainder = index - count;
-	  coupled_rank = ii;
-	  break;
-	}
-	else
-	  count += other_count;
+        int other_count = (*cp)(ii,rank);
+        if(count + other_count >= index)
+        {
+          remainder = index - count;
+          coupled_rank = ii;
+          break;
+        }
+        else
+          count += other_count;
       }
       count = 0;
       for(int ii = 0; ii < rank; ii++)
-	count += (*cp)(coupled_rank,ii);
+        count += (*cp)(coupled_rank,ii);
       result.first = coupled_rank;
       result.second = count + remainder;
     }
     return result;
   }
-
-    /// @brief Default constructor.
[…]

### 2016-01-15 amsi_meta/CommPattern.h

@@ -14,23 +14,18 @@ namespace amsi
   class CommPattern : public Assemblable
   {
   public:
-    CommPattern(int so, int st)
-      : s1(so)
+  CommPattern(int so, int st)
+    : s1(so)
       , s2(st)
     {}
     virtual int& operator()(int,int) = 0;
     virtual int operator()(int,int) const = 0;
-
-    void getSentTo(int rank, std::vector<int>& sent_to) const;
-    void getRecvedFrom(int rank, std::vector<int>& recv_from) const;
-
     int getNumSenders() const {return s1;}
     int getNumRecvers() const {return s2;}
   protected:
     int s1;
     int s2;
   };
-  
   class FullCommPattern : public CommPattern
   {
   public:
@@ -42,16 +37,25 @@ namespace amsi
     FullCommPattern();
     int * pattern;
   };
-
+  /// count the number of processes the specified sending process is sending to
+  int countRanksSentTo(const CommPattern * cp, int rnk);
+  /// count the number of processes the specified recving process is recving from
+  int countRanksRecvFrom(const CommPattern * cp, int rnk);
+  /// get the task ranks of the processes the specified sending process is sending to
+  void getRanksSentTo(const CommPattern * cp, int rnk, int * rnks);
+  /// get the task ranks of the processes the specified recving process is recving from
+  void getRanksRecvFrom(const CommPattern * cp, int rnk, int * rnks);
+  /// get the unit count for the specified rank to send to the ranks given by getRanksSentTo()
+  void getUnitsSentTo(const CommPattern * cp, int rnk, int * sent_to);
+  /// get the unit count for the specified rank to recv from the ranks given by getRanksRecvFrom()
+  void getUnitsRecvFrom(const CommPattern * cp, int rnk, int * recv_from);;
   std::ostream& operator<<(std::ostream& os, const CommPattern& obj);
-
   /// Create a new CommPattern which is an inversion of the old one (ie all sending processes in the old CommPattern are recvers in the produced CommPattern and vice-versa, essentially a matrix transposition)
   CommPattern * CommPattern_CreateInverted(const CommPattern * pattern);
-
-    // retrieve information about the rank and remote index of a piece of local data by index
+  // retrieve information about the rank and remote index of a piece of local data by index
   std::pair<int,int> coupledInfoByIndex(CommPattern * cp,
-					Role role,
-					int rank,
-					int index);
+                                        Role role,
+                                        int rank,
+                                        int index);
 } // namespace amsi
 #endif

### 2016-01-15 amsi_control/ControlService.cc
quick debugging fix

@@ -174,10 +174,10 @@ namespace amsi
     // determine the task rank
     int task_rank = tl->localRank();
     // switch to the correct communicator for this operation
-#     ifndef CORE
+#   ifdef CORE
     PCU_Switch_Comm(comm_man->CommRelation_GetInterComm(r_dd_id.first));
     PCU_Comm_Begin();
-#     endif
+#   endif
     if(tl == t1) // sending task
     {
       // send pattern info to t2_per_t1 + -many ranks in t2
@@ -196,22 +196,22 @@ namespace amsi
         getUnitsRecvFrom(send_pattern,snd_to,&rcv_cnts[0]);
         // intercomm rank of the recving task (hacky)
         int inter_rnk = t1s+snd_to;
-#         ifdef CORE
+#       ifdef CORE
         PCU_Comm_Write(inter_rnk, &num_rcv_rnks, sizeof(int));
         PCU_Comm_Write(inter_rnk, &rcv_rnks, sizeof(int)*num_rcv_rnks);
         PCU_Comm_Write(inter_rnk, &rcv_cnts, sizeof(int)*num_rcv_rnks);
-#         else
+#       else
         t_isend(recvfrom,MPI_INTEGER,t2->localToGlobalRank(send_to));
-#         endif
+#       endif
       }
-#       ifdef CORE
+#     ifdef CORE
       // All processes must call PCU Send and Read, makes this blocking which is bad
       PCU_Comm_Send();
       int rcv_frm = -1;
       void * rcv = NULL;
       size_t rcv_sz = 0;
       while(PCU_Comm_Read(&rcv_frm,&rcv,&rcv_sz)) {}
-#       endif
+#     endif
     }
     else //recving task
     {
@@ -240,7 +240,7 @@ namespace amsi
       t_recv(recv_count,MPI_INTEGER,t1->localToGlobalRank(recv_from));
 #     endif
     }
-#   ifndef CORE
+#   ifdef CORE
     // Switch pcu comms back to task
     PCU_Switch_Comm(tl->comm());
 #   endif

### 2016-01-24 doc/articles/amsi_biotissue/paper.tex
lots of work on the paper, still a lot to do

@@
 \date{\today}
-\documentclass[11pt]{siamltex1213}
+\documentclass[11pt]{article}
 \bibliographystyle{siam}
 
+\usepackage{comments}
 \usepackage{amsfonts}
@@
 
-\section{Note}\label{note}
-This article predominantly discusses multi-scale simulations - wherein two or more separate physical scales are involved in the simulation - the discussion is equally applicable to multi-model simulations - wherein two or more numerical or physical models operating on the same domain (possibly with differing discretizations) are involved in the simulation - as well as to simulations which are both.
+\section{Introduction}\label{introduction}
+[Need a couple sentences saying that multi-scale is important]
 
-\section{Introduction}\label{introduction}
-\label{single_scale_simulations}
-Most physical models and implementations using numerical methods operate at a single scale characteristic to the problem.
-The scale at which the preponderance of attributes governing the underlying physics being modelled are measured and the resolution of the numerical solution is considered acceptable for the desired accuracy of the simulation.
+Multi-scale influences can also be incorporated in a more strictly numerical or computational sense, combining implementations of established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
 
-\label{multi_scale_simulations}
-Introducing multi-scale influences on the scale of interest in a simulation can occur in several places during the development of a simulation.
-Most fundamentally, the influence can be incorporated during the mathematical development of the physical model itself.
-In this case the multi-scale features are incorporated into the numerical implementation automatically, as a result of their presence in the mathematical model.
+% classes of multi-scale problems
+In order for a numerically multi-scale approach to be effective the domain associated with the scale of interest (the engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude removed would make the simulation computationally intractable.
+Thus some computationally reasonable subset of the engineering scale is assigned significantly finer scale(s) to simulate. 
+The choice of location of supporting fine-scale simulations is largely dependent on the specific mathematical models and numerical methods used to implement the engineering scale simulation.
+Two general classes of fine-scale simulation locality have been noted \cite{weinan2011principles}, 'A' type problems where local granularity requirements such as boundary layers or crack propagation necessitate use of localized fine-scale simulation, and 'B' type problems where fine-scale simulation is used to supplement or replace a continuum model at the engineering scale.
+For either case the subset of the engineering-scale domain associated with finer scales are areas of the domain where the fine-scale contributions have the greatest impact on the engineering scale simulation, in order to leverage the most utility from the multi-scale approach.
 
-Multi-scale influences can also be incorporated in a numerical sense, combining established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
-Numerically multi-scale simulations associate many numerical implementations of single-scale physical models and cause them to interact by passing key physical terms between scales.
-This operation, referred to as scale-coupling, involves up-/down-scaling operations to account for differences in scale dimensionality (and possibly dimensionalizing terms computed in a non-dimensional domain).
+% types of numerical paradigms related to multi-scale, maybe mention that a problem using the concurrent multi-scale paradigm is discussed later in the paper?
+Numerical multi-scale simulations fall broadly into two categories.
+In the sequential multi-scale paradigm, fine-scale computation is used to precompute quantities of interest for establishing an engineering-scale physical model, possibly using a lookup table or some other method to incorporate the multi-scale influences into the engineering scale \cite{weinan2011principles} \cite{garcia2008sequential}.
+In the concurrent multi-scale paradigm, all scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{weinan2011principles} \cite{zeng2010concurrent}.
 
-In order for a numerically multi-scale approach to be effective the domain associated with the scale of interest (or engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude removed would make the simulation computationally intractable.
-Thus some computationally reasonable subset of the engineering scale is assigned tertiary scale(s) to simulate.
-The precise choice of location of the tertiary simulations is largely dependent on the specific numerical algorithms used to implement the engineering scale simulation.
+% introduce scale-coupling topic to be discussed further in AMSI and methods sections
+Numerically multi-scale simulations associate the computational implementations of single-scale physical models and cause them to interact by passing key physical terms between scales. 
+This scale-coupling operation and may require the application of transformer functions to account for key physical differences between disparate scales. 
+The primary purpose of such transformer functions is to convert physical coupling values produced by simulations operations at one scale to be meaningfully applied to simulation operations at the coupled, consuming scale.
+These transformer functions are typified by conversions between the unit metrics used by coupled scales. 
+For example a micro-scale simulation will typically be implemented with units of $\mu$m while a macro-scale simulation might be implemented using units of cm.
+Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales, such as dimensionalizing terms calculated in a dimensionless domain.
 
-Often the subset of the domain associated with the tertiary scales are areas of the engineering domain where the impact of the tertiary scales has the most impact on the engineering scale simulation, in order to leverage the most utility from the multi-scale approach.
+% introduce legacy component concept which should be discussed in AMSI and methods sections
+Decades of numerical simulation development have resulted in a large and diverse array of widely used codes for single-scale simulations. 
+This set of legacy codes represents a massive investment of development efforts, and thus leveraging the capabilities of such legacy codes for use in multi-scale simulations is desirable. 
+Incorporating such codes into a functional numerical implementation of a multi-scale model has been a largely ad-hoc, per-code affair. 
+Providing tools to ease the incorporation of legacy components into multi-scale simulations using general procedures while requiring minimal intervention in the legacy codebase results in less fracturing of the legacy codebases when incorporated into varied multi-scale simulations, and avoids the option of complete reimplementation of legacy code functionality in specialized multi-scale ad-hoc codes, or structured frameworks or runtimes.
 
-Numerical multi-scale simulations fall broadly into two categories.
-In the concurrent multi-scale paradigm, all scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{zeng2010concurrent}.
-In the sequential multi-scale paradigm, tertiary scales compute parameters which are converted to quantities of utility to a primary scale.
-Each scale (though not necessarily each scale-task, representing an individual operation or simulation occurring in a scale) must be processed sequentially \cite{garcia2008sequential} and may be blocked on the completion of the previous scale.
-This multi-scale model maps intuitively to phased parallel communication models found in the commonly used bulk synchronous parallel abstraction \cite{valiant1990bridging}.
+Implementation of multi-scale simulations has fallen widely into two categories: purely ad-hoc implementations, and implementations supported by a structured framework or runtime. 
+In the case of an ad-hoc implementation all support for multi-scale interaction is hard-coded into the simulation code itself. 
+Frameworks and runtimes provide a robust set of scale-coupling operations, scale management,  and overall simulation management. 
+The application programmer must tailor the numerical operations occurring at a discrete scale to operate using a set of interfaces or high-level modelling language required by the framework/runtime.
+Ad-hoc implementations allow for overall optimizations at the cost of development time (and potentially complexity). 
+Framework/runtime implementations often have a component-based approach that breaks the implementation of a simulation (single- or multi-scale) into discrete interoperable components. 
+This approach requires that operations are implemented to work specifically within the set of interfaces understood by the framework/runtime.
 
-\section{Adaptive Multi-scale Simulation Infrastructure}\label{amsi}
-The Adaptive Multi-scale Simulation Infrastructure (amsi) is a set of libraries and tools developed at the Scientific Computation Research Center (SCOREC) at Rensselaer Polytechnic Institute.
-Amsi is designed to support the implementation and execution of dynamic numerically multi-scale simulations on massively-parallel HPC machines.
-Currently amsi provides interfaces for describing and managing the execution of individual scales in a multi-scale simulation, for defining, managing and enacting the scale-coupling communication required by such simulations, and for planning and enacting scale-sensitive load balancing operations on individual scales in a multi-scale system.
+\section{Adaptive Multi-scale Simulation Infrastructure}\label{AMSI}
+The Adaptive Multi-scale Simulation Infrastructure (AMSI) is a set of libraries and tools developed by the Scientific Computation Research Center (SCOREC) at Rensselaer Polytechnic Institute. 
+AMSI is designed to support the implementation and execution of dynamic concurrent multi-scale simulations on massively-parallel HPC machines.
+AMSI provides support for several key operations needed for concurrent multi-scale simulations: modelling and managing scales, scale-coupling parallel communication, adaptively adding/removing scale-tasks, and load-balancing coupled scales.
+A key goal in the design and implementation of these operations is a focus on general applicability and avoiding adherence to specific interfaces.
 
-The interfaces provided by the amsi libraries are intended to facilitate the coupling of functional (possibly legacy) simulation codes already in usage for single-scale problems, in order to avoid development overhead associated with casting a simulation into a set of framework-specific components.
-The currently implemented interfaces are physics-agnostic, operating only on computational quantities with no model of a term's relation to the mathematical derivation of the multi-scale interaction.
-Incorporation of meta-information and services targeted toward specific multi-scale physical couplings represents a desirable extension of the amsi libraries, constituting a rich area of future work.
+The interfaces provided by the AMSI libraries are intended to facilitate the coupling of functional single-scale simulation codes.
+This approach is used in order to mitigate development overhead associated with casting a simulation into a set of framework-specific components, or developing an entire ad-hoc set of multi-scale operations each time a new multi-scale code is being developed.
+Several key multi-scale operations are planned and enacted by the AMSI libraries, allowing the application programmer (a programmer using the AMSI libraries to implement a new multi-scale code or interaction) to determine the most effective use of the tools provided.
+All interfaces are designed to be physics-agnostic, operating only on computational quantities with no meta-model of the relation between a coupling term and the mathematical derivation of the multi-scale interaction.
 
-Amsi operates by maintaining a set of minimal simulation metadata in order to model various quantities of interest during simulation execution.
-A decentralized approach toward control is taken; dynamic control decisions are made and implemented at runtime during collective operations.
-In order to avoid the introduction of unnecessary parallel barriers into the code, amsi control decisions are only made during operations which are already collective over the set of processes effected by a given control decision.
+AMSI operates by maintaining a set of minimal simulation metadata in order to model various quantities of interest during simulation execution. 
+AMSI provides mechanisms to assist in the implementation of run-time decision making and control decisions related to multi-scale execution, with a focus on allowing such adaptive procedures to be planned and implemented in a decentralized manner.
+This design approach allows application developers to avoid centralized control decision bottlenecks.
+Control decisions and associated parallel communication (as with most adaptive procedures) are best implemented at locations in code already requiring parallel barriers for synchronization.
 
+In order to avoid the introduction of unnecessary parallel barriers into the code, AMSI control decisions are only  during operations which are already collective over the set of processes effected by a given control decision.
+
+% need to expand this substantially
 Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{parker2006component} \cite{chopard2011framework} \cite{} and these frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah}.
 
-These systems suffer from the requirement that an application developer must cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or specify the problem through a high-level modelling language in order to take full advantage the services provided by the framework.
-Leveraging a component-based design is valuable from a software engineering perspective, but also introduces new complexities involved in defining the set of components and how they interoperate in the framework, shifting some complexity up the abstraction hierarchy.
-The core features of amsi used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.
+Such systems require that the application developer cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or specify the problem through a high-level modelling language in order to take full advantage the services provided by the framework.
+Leveraging a component-based design is valuable from a software engineering perspective, as it allows composition of disparate components into a cohesive application. 
+This approach also introduces new complexities involved in defining the set of components and how they interoperate in the framework, shifting some complexity up the abstraction hierarchy.
+The core features of AMSI used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.
 
-Typically simulation flow control is determined solely by the runtime system.
+Typically simulation flow control is determined solely by a runtime system. 
 A framework may also limit specific features of the numerical implementation to those provided by the framework. 
@@
 
-Amsi supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations of these interfaces built on top of other software supplied by SCOREC \cite{core} for finite element analysis operations as well as the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}.
-Usage of the supplied numerical components is not required to make use of the multi-scale meta-modeling, communication, dynamic instantiation, and scale-sensitive load balancing facilities supplied by amsi.
+AMSI supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations of these interfaces built on top of other software supplied by SCOREC \cite{core} for finite element analysis operations, Simmetrix mesh database software \cite{simmetrix}, and the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}.
+Usage of the supplied numerical components is not required to make use of the multi-scale capabilities of the AMSI libraries, but provide a jumping-off point for application developers needing to implement a new single-scale code to incorporate into a multi-scale simulation.
 
@@
 
-\section{Methods}
+\section{Methods}\label{methods}
 
-\label{amsi_scales}
-A configuration file is supplied during command invocation containing the initial configuration of the process sets and their assignment to individual scales, as well as declaring the coupling between scales in the simulation.
-The declaration of a scale and the initialization of the process set associated with the scale are considered a single atomic operation in the current implementation.
+\subsection{Simulation Construction with AMSI}
+In order to construct a multi-scale simulation using the AMSI libraries the following must be specified: the Scales in the problem, the ScaleCouplings between individual scales, the DataDistribution of coupling data produced by each Scale in a coupling, and the CouplingPattern used to communicate coupling data during coupling operations.
 
-Process sets are mathematical sets of MPI ranks implemented so as to take advantage of any mathematical conveniences to minimize explicit storage whenever possible.
-Process sets are required to be non-overlapping in order for amsi control mechanisms to operate correctly.
-Ongoing developments are targeted at removing this requirement to allow a richer set of dynamic control options and increase overall utilization in the case of coupled scales blocking on one another.
+A Scale is a data structure modelling the parallel state of a discrete simulation scale and all operations (or tasks) occurring at that scale. 
+All Scales have a set of MPI_Ranks associated with them through an intermediary structure called the RankSet. 
+In the current implementation the RankSet of any two Scales in a simulation must be non-overlapping.  
+A Scale is required for the specification of all other data structures in the AMSI libraries.
 
-Scales are an abstraction grouping a set of standalone scale-tasks operating on a single scale, the set may be as minimal as a single scale-task as in the case of a primary scale with many contributing tertiary scales.
-The set may also contain many individual scale-tasks which all operate within a single scale, such as when a primary scale-task has an associated tertiary scale which constitutes many individual simulations, as in the biotissue example case discussed in section \ref{biotissue}.
-Precisely what defines a scale-task is left up to the user, the intent is for each scale-task to represent a standalone single-scale simulation incorporated into the multi-scale system.
+A ScaleCoupling is an object denoting that scale-coupling communication will take place between two scales at some point during the simulation. 
+All interacting Scales in a simulation must declare a ScaleCoupling between them.
+For generality the ScaleCoupling relationship is non-commutative, thus Scale A is related to Scale B if and only if A produces coupling data which will be sent to B.
+Symmetrical coupling relations therefore require two ScaleCouplings, one denoting the production of data by A for consumption by B and one denoting the reverse.
+Scales not related through a ScaleCoupling cannot create CouplingPatterns which are required for multi-scale communication and therefore cannot pass scale-coupling data through the AMSI libraries.
 
-A scale coupling is a relation denoting that scale-coupling communication will take place between two scales at some point during the simulation.
-Declaration of a coupling relation between two scales is required for communicating scales, as the coupling declaration during simulation initialization is used to instantiate data structures specific to scale-coupling operations.
-This limits memory overhead by not allocating space for unneeded coupling operations.
-This coupling relationship is asymmetric, thus if scale A is related to scale B (ie.
-A produces terms required for B to operate), scale B is not implicitly related to scale A, though this may be separately declared.
+A DataDistribution is a structure denoting the parallel distribution of terms produced by a scale for use in a ScaleCoupling relationship for a specific CouplingPattern. 
+The DataDistribution is minimally described as a parallel array of integers on the RankSet associated with some Scale A, thus the length of the array is $|A|$ (the cardinality of the RankSet currently assigned to Scale A).
+The DataDistribution may be dynamically modified during the simulation to reflect single-scale adaptive procedures or specifically multi-scale adaptive features (such as selectively adding/removing specific fine-scale simulations).
+Each Scale may have many DataDistributions, and a DataDistribution is not tied to a particular ScaleCoupling, in the event that the same distribution characterizes the scale-coupling data produced by a Scale needed by multiple coupled scales.
 
-A simple configuration file can be seen in listing \ref{amsi_config_file} which declares the existence of two scales, macro and micro\_fo, and tells the amsi intialization procedures to assign them process sets containig 64 and 8128 processes, respectively (for a total of $2^{13}$ processes).
-The file also specifies the scale-coupling relations, with macro coupled to micro and micro coupled to macro.
-Note that the order of specification is not relevant to the operation of amsi.
-%
-\begin{lstlisting}[caption={amsi\_config file},label=amsi_config_file]
-  macro 64
-  micro_fo 8128
-  #
-  macro micro_fo
-  micro_fo macro_fo
-\end{lstlisting}
-%
+A CouplingPattern describes the parallel communication required to send scale-coupling data produced by the the sending Scale in a ScaleCoupling and described by a DataDistribution. 
+A CouplingPattern based on the ScaleCoupling of scale $A$ coupled to scale $B$ is described most simply as a matrix $C \in \mathbb{N}^{|A|\times|B|}$ (where $|X|$ is the cardinality of the RankSet currently assigned to scale X).
+The quantity $C_{ij}$ represents the quantity of scale-coupling terms (of arbitrary type) to be sent from process $i$ in the process set assigned to $A$ to process $j$ in the process set assigned to $B$.
+As the DataDistribution upon which the CouplingPattern is based is allowed to change due to adaptive procedures, all CouplingPatterns related to a DataDistribution are updated when adding/removing terms from a DataDistribution.
+Finalization of the creation of a CouplingPattern causes the creation of a DataDistribution on the receiving Scale, as modifications to a DataDistribution may happen on either side of the ScaleCoupling, but the initial CouplingPattern must be created by the producing Scale.
+Determination of how a CouplingPattern should change based on changes to the underlying DataDistribution may differ for potentially each multi-scale use-case, so hooks are provided so the application program may correctly update CouplingPatterns for their specific use case.
+As both Scales in the ScaleCoupling must know about the modification in order to appropriately handle the modified scale-coupling data being communicated, all modifications to the CouplingPattern and DataDistribution are communicated from the producing Scale to the consuming Scale.
+Due to this requirement it is best to update DataDistributions completely in a single operation following any adaptive procedures, since these are often parallel barriers already.
 
-Dynamically creating the required configuration file as part of a shell script invoking the simulation is straightforward.
-Alternative specification mechanisms such as ratio specification (equivalently percentage specification) are being considered for future implementation to prevent the possibility of misallocating processes (the user not specifying assignment for all processes available to the simulation, or attempting to overallocate processes).
+Similar operations are provided for scale-sensitive load-balancing operations taking place on a sending or receiving scale in a ScaleCoupling, as the associated scale must be informed of the changes to the relevant DataDistribution.
 
-The assignment of processes to a process set is static, and processes cannot be reassigned during a simulation.
-Additionally the assignment of a scale to a process set is static as well.
-The application programmer may provide capabilities for scale-tasks to be dynamically created and destroyed at runtime, amsi provides functions to inform coupled scales of such dynamic modifications.
-Since scale-tasks are an abstraction the amsi systems cannot provide direct operations on scale-tasks, but provide mechanisms for the application to inform amsi of such operations when they occur.
+\subsection{Parallel Operations}\label{paralle_ops}
+
+All communication and planning operations in the AMSI libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer. 
+Usage only of implicit ordering was decided upon in order to reduce the memory impact as the simulation scales, as any meta-data which scales with the simulation will necessarily pose a memory issue once scaling hits some threshold.
+
+Since scale-tasks are an abstraction the AMSI systems cannot provide direct operations on scale-tasks, but provide mechanisms for the application to inform AMSI of such operations when they occur.
 
@@
 Migrating a scale task in whole constitutes using load balancing operations on the process set associated with the entire scale to move all pertinent data between subset of the managed process set.
-Amsi provides operations to enact load-balancing operations from plans such as those generated by Zoltan algorithms.
+AMSI provides operations to enact load-balancing operations from plans such as those generated by Zoltan algorithms.
 In the current implementation these operations are collective on the process set associated with the managed scale.
 
-While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the amsi implementation.
+While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the AMSI implementation.
 Any scale is capable of informing the process set associated with a coupled scale of the addition or deletion of individual scale-tasks.
@@
 
-\label{amsi_communication}
 Scale-coupling communication is guided by data distributions and communication patterns.
@@
 
-Scale-linking quantities are dynamically registered with the amsi system using a data distribution defined on a scale by the scale-tasks operating at that scale.
+Scale-linking quantities are dynamically registered with the AMSI system using a data distribution defined on a scale by the scale-tasks operating at that scale.
 Production of a communication pattern is required prior to executing a scale-coupling communication, and the pattern must be reconciled (synchronized across the coupling relationship) prior to being used to guide a coupling communication operation (and when the communication pattern is altered by the scale providing the coupling data).
-Several simple communication pattern production algorithms are provided by amsi, but these are relatively trivial due to the nature of scale-coupling communications and the general nature of the amsi libraries.
+Several simple communication pattern production algorithms are provided by AMSI, but these are relatively trivial due to the nature of scale-coupling communications and the general nature of the AMSI libraries.
 Hooks are provided for users to implement algorithms for the construction of communication patterns tailored to the specifics of their problem.
-
-A communication pattern based on the relationship of scale $A$ coupled to scale $B$ is thought of most simply as a matrix $C \in \mathbb{N}^{|A|\times|B|}$ (where $|X|$ is the cardinality of the process set currently assigned to scale X).
-The implementation of the communication pattern uses a sparse format in order to reduce the memory impact as the simulation scales.
-The quantity $C_{ij}$ represents the quantity of scale-coupling terms (of arbitrary type) to be sent from process $i$ in the process set assigned to $A$ to process $j$ in the process set assigned to $B$.
-
-All communication and planning operations in the amsi libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer.
-Usage only of implicit ordering was decided upon in order to reduce the memory impact as the simulation scales, as any meta-data which scales with the simulation will necessarily pose a memory issue once scaling hits some threshold.
 
@@
 The current implementation includes blocking versions of the scale-coupling communication operations.
-Asynchronous versions have been designed but are not currently implemented due to lack of immediate need in the primary amsi use case.
+Asynchronous versions have been designed but are not currently implemented due to lack of immediate need in the primary AMSI use case.
 
@@
 Separate timings were taken for various sections of the simulation code, isolating the micro-scale computation, macro-scale linear system assembly and calculation of elemental linear systems, and macro-scale linear system solution.
-In addition, the time for spent in the amsi subroutines for communication was also measured.
+In addition, the time for spent in the AMSI subroutines for communication was also measured.
 Thus the behavior of each section of the code is independently evaluated in our scaling study, as well as the overall execution time.
@@
 Many problems of interest will possess finite elements numbered in the tens to hundreds of millions of degrees of freedom, such that a suitable partitioning of the mesh to cores on the machine in terms of elements per partition (and directly the macro-scale time-to-solution) may require a substantial fraction of the total cores on the machine for macro-scale to execute in reasonable time.
-Amsi provides services for dynamically adding and removing individual scale-tasks to a specific scale, which can be utilized to limit RVE usage to only those locations in the multi-scale mesh most influential on the overall evolution of the simulation.
+AMSI provides services for dynamically adding and removing individual scale-tasks to a specific scale, which can be utilized to limit RVE usage to only those locations in the multi-scale mesh most influential on the overall evolution of the simulation.
 
-Amsi-specific results of some sort...?
+AMSI-specific results of some sort...?
 
@@
 The Adaptive Multi-scale Simulation Infrastructure provides a valuable set of tools to facilitate the implementation and execution of multi-scale simulations without imposing restrictions on the application developer.
-Usage of the core operations of the amsi libraries does not require adherence to a specific set of interfaces nor usage of an abstract modeling language, which complicate inclusion of legacy components into multi-scale simulations.
+Usage of the core operations of the AMSI libraries does not require adherence to a specific set of interfaces nor usage of an abstract modeling language, which complicate inclusion of legacy components into multi-scale simulations.
 
@@
 Development of libraries concerned with higher levels of abstraction of the multi-scale problem will also be valuable in easing the development of new multi-scale simulations from the perspective of the application programmer.
-High level abstractions such as the Scale Separation Map in \cite{chopard2011framework} could provide mechanisms to provide initial configuration information to the amsi system during the initialization phase, as well as assist in the implementation of (or usage of existing library-provided) up-/down-scaling operations.
+High level abstractions such as the Scale Separation Map in \cite{chopard2011framework} could provide mechanisms to provide initial configuration information to the AMSI system during the initialization phase, as well as assist in the implementation of (or usage of existing library-provided) up-/down-scaling operations.
 
@@
 This alternative RVE will be incorporated into the multi-scale model as an alternative to the fiber-only RVE, to be used in areas of the macro-scale domain determined to be most critical to the overall evolution of the global system, or where error accumulation through model inaccuracy is to be minimized.
-As this fiber-matrix RVE represents a more significant source of computational complexity, it is independently parallelized -- in contrast to the fiber-only RVE -- and will represent a compelling use case for the parallel execution space discretization and scale-coupling communication management capabilities of the amsi libraries.
+As this fiber-matrix RVE represents a more significant source of computational complexity, it is independently parallelized -- in contrast to the fiber-only RVE -- and will represent a compelling use case for the parallel execution space discretization and scale-coupling communication management capabilities of the AMSI libraries.
 \bibliography{citations}

### 2016-01-24 doc/articles/amsi_biotissue/paper.tex
more paper work

@@
 \section{Introduction}\label{introduction}
-[Need a couple sentences saying that multi-scale is important]
+[Need a couple sentences to talk about how multi-scale is important]
+Implementation of multi-scale numerical simulations typically rely on either an ad-hoc set of algorithms designed for a particular problem, or on large frameworks and runtimes. 
+Ad-hoc approaches, by their nature, are targeted to a particular multi-scale problem for a specific set of single-scale numerical codes, and thus offer no wider utility beyond their immediate usage as a simulation code.
+Framework and runtime approaches are more flexible and admit a broader range of reuse, but typically necessitate that single-scale operations be adapted to work in the ecosystem they create.
 
-Multi-scale influences can also be incorporated in a more strictly numerical or computational sense, combining implementations of established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
+Development of multi-scale simulations and methods to support the implementation of such simulations has been ongoing for several decades. 
+Some approaches begin with the specification of multi-scale contributions during the specification of mathematical model.
+However, multi-scale influences can be incorporated in a strictly numerical or computational sense, by combining implementations of established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
 
 % classes of multi-scale problems
-In order for a numerically multi-scale approach to be effective the domain associated with the scale of interest (the engineering scale) must be sufficiently large that conducting the entire simulation with terms and granularity orders of magnitude removed would make the simulation computationally intractable.
+In order for a numerically multi-scale approach to be effective the domain associated with the scale of interest (the engineering scale) must be of sufficient size that conducting the entire simulation with terms and granularity orders of magnitude removed would make the simulation computationally intractable.
 Thus some computationally reasonable subset of the engineering scale is assigned significantly finer scale(s) to simulate. 
@@
 Two general classes of fine-scale simulation locality have been noted \cite{weinan2011principles}, 'A' type problems where local granularity requirements such as boundary layers or crack propagation necessitate use of localized fine-scale simulation, and 'B' type problems where fine-scale simulation is used to supplement or replace a continuum model at the engineering scale.
-For either case the subset of the engineering-scale domain associated with finer scales are areas of the domain where the fine-scale contributions have the greatest impact on the engineering scale simulation, in order to leverage the most utility from the multi-scale approach.
+For either case the subset of the engineering-scale domain associated with finer scales are areas of the domain where the fine-scale contributions have the greatest impact on the evolution of the engineering scale simulation, in order to leverage the most utility from the multi-scale approach.
 
 % types of numerical paradigms related to multi-scale, maybe mention that a problem using the concurrent multi-scale paradigm is discussed later in the paper?
-Numerical multi-scale simulations fall broadly into two categories.
+Numerical multi-scale simulations fall broadly into two parallel paradigms, distinct from the choice of subdomain assigned to fine scale simulation.
 In the sequential multi-scale paradigm, fine-scale computation is used to precompute quantities of interest for establishing an engineering-scale physical model, possibly using a lookup table or some other method to incorporate the multi-scale influences into the engineering scale \cite{weinan2011principles} \cite{garcia2008sequential}.
@@
 Due to this requirement it is best to update DataDistributions completely in a single operation following any adaptive procedures, since these are often parallel barriers already.
-
 Similar operations are provided for scale-sensitive load-balancing operations taking place on a sending or receiving scale in a ScaleCoupling, as the associated scale must be informed of the changes to the relevant DataDistribution.
 
+ScaleCoupling transformation functions are not currently supported by the AMSI libraries and must be implemented by the application programmer for the each specific multi-scale interaction.
+
 \subsection{Parallel Operations}\label{paralle_ops}
+
+Each Scale in AMSI represents some number of individual scale-tasks. 
+Scale-tasks are not an AMSI construct, merely an abstraction to allow grouping of multiple copies of the same single-scale simulation into a single Scale. 
+Minimally a Scale possess a single scale-task. 
+For example the biotissue simulation (discussed in Section \ref{biotissue}) the Macro Scale has a single scale-task, a parallelized finite element analysis simulation.
+However a Scale may also possess many scale-tasks, each a duplicate of a single-scale simulation.
+Again in the biotissue example case, the Micro_FO Scale has a number of scale-tasks (individual Representative Volume Element (RVE) simulations) related to the number of Macro mesh elements.
+The management of these scale-tasks the responsibility of the application programmer, though changes made that redistributed terms modelled in any DataDistributions should be noted and used to update appropriately once dynamic/adaptive operations have ceased.
+Adding/removing terms from a DataDistribution may result in the requirement that scale-tasks be instantiated/deleted to accommodate the modification.
+
+When DataDistributions are modified the modifications must be reflected across the RankSet associated with the Scale on which the DataDistribution is defined. 
+This procedure is called Assembly, where each rank contributes the local (and hence valid) portion of the control data to arrive at a valid Scale-global value.
+In practice this reduces to an MPI_Allgather operation for the DataDistribution, but for other types of AMSI control data (most specifically the CouplingPattern) the implementation of the Assembly operation requires additional operations.
+
+When CouplingPatterns are modified (as a result of a changing DataDistribution) the Scale informing the modification must also ensure that the coupled Scale receives the corrected CouplingPattern. 
+This process is called Reconciliation, and has two phases the inter-scale phase and the local-scale phase.
+In the inter-scale phase a sparse representation of the relevant CouplingPattern data (relevant meaning related to the rank receiving the data, informing that rank fully of only the role it plays in the communication) is communicated from the modifying scale to the modified scale.
+In the scale-local phase a specialized Assembly operation is performed on the received partial updates to update the full CouplingPattern representation on all processes in the modified Scale.
+
+%% revised up to here, with sporadic edits after
 

### 2016-01-24 doc/articles/amsi_biotissue/paper.tex
getting paper to compile

@@
 
-\usepackage{comments}
 \usepackage{amsfonts}
@@
 \usepackage{mathtools}
+\usepackage{url}
 \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
@@
 A Scale is a data structure modelling the parallel state of a discrete simulation scale and all operations (or tasks) occurring at that scale. 
-All Scales have a set of MPI_Ranks associated with them through an intermediary structure called the RankSet. 
+All Scales have a set of MPI\_Ranks associated with them through an intermediary structure called the RankSet. 
 In the current implementation the RankSet of any two Scales in a simulation must be non-overlapping.  
@@
 However a Scale may also possess many scale-tasks, each a duplicate of a single-scale simulation.
-Again in the biotissue example case, the Micro_FO Scale has a number of scale-tasks (individual Representative Volume Element (RVE) simulations) related to the number of Macro mesh elements.
+Again in the biotissue example case, the Micro\_FO Scale has a number of scale-tasks (individual Representative Volume Element (RVE) simulations) related to the number of Macro mesh elements.
 The management of these scale-tasks the responsibility of the application programmer, though changes made that redistributed terms modelled in any DataDistributions should be noted and used to update appropriately once dynamic/adaptive operations have ceased.
@@
 This procedure is called Assembly, where each rank contributes the local (and hence valid) portion of the control data to arrive at a valid Scale-global value.
-In practice this reduces to an MPI_Allgather operation for the DataDistribution, but for other types of AMSI control data (most specifically the CouplingPattern) the implementation of the Assembly operation requires additional operations.
+In practice this reduces to an MPI\_Allgather operation for the DataDistribution, but for other types of AMSI control data (most specifically the CouplingPattern) the implementation of the Assembly operation requires additional operations.
 

### 2016-01-25 doc/articles/amsi_biotissue/paper.tex
paper

@@
 For example a micro-scale simulation will typically be implemented with units of $\mu$m while a macro-scale simulation might be implemented using units of cm.
-Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales, such as dimensionalizing terms calculated in a dimensionless domain.
-
-% introduce legacy component concept which should be discussed in AMSI and methods sections
-Decades of numerical simulation development have resulted in a large and diverse array of widely used codes for single-scale simulations. 
-This set of legacy codes represents a massive investment of development efforts, and thus leveraging the capabilities of such legacy codes for use in multi-scale simulations is desirable. 
-Incorporating such codes into a functional numerical implementation of a multi-scale model has been a largely ad-hoc, per-code affair. 
-Providing tools to ease the incorporation of legacy components into multi-scale simulations using general procedures while requiring minimal intervention in the legacy codebase results in less fracturing of the legacy codebases when incorporated into varied multi-scale simulations, and avoids the option of complete reimplementation of legacy code functionality in specialized multi-scale ad-hoc codes, or structured frameworks or runtimes.
+Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales \cite{weinan2003heterogenous}, such as dimensionalizing terms calculated in a dimensionless domain \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.
 
@@
 This approach requires that operations are implemented to work specifically within the set of interfaces understood by the framework/runtime.
+
+% introduce legacy component concept which should be discussed in AMSI and methods sections
+Decades of numerical simulation development have resulted in a large and diverse array of widely used codes for single-scale simulations. 
+This set of legacy codes represents a massive investment of development efforts, and thus leveraging the capabilities of such legacy codes for use in multi-scale simulations is desirable. 
+Incorporating such codes into a functional numerical implementation of a multi-scale model has been a largely ad-hoc, per-code affair. 
+Providing tools to ease the incorporation of legacy components into multi-scale simulations using general procedures while requiring minimal intervention in the legacy codebase results in less fracturing of the legacy codebases when incorporated into varied multi-scale simulations, and avoids the option of complete reimplementation of legacy code functionality in specialized multi-scale ad-hoc codes, or structured frameworks or runtimes.
 
@@
 All Scales have a set of MPI\_Ranks associated with them through an intermediary structure called the RankSet. 
-In the current implementation the RankSet of any two Scales in a simulation must be non-overlapping.  
+In the current implementation the RankSet of all Scales in a simulation must be non-overlapping.  
 A Scale is required for the specification of all other data structures in the AMSI libraries.
@@
 
+Constructing a new multi-scale simulation from a set of legacy single-scale simulation codes involves several steps.
+Each Scale in the simulation must have a 'main-like' function with the signature \verb|int (int *, char *[], MPI_Comm)| which initializes the various supporting libraries used by that single-scale code on the provided MPI\_Comm and then begins conducting the simulation at that Scale.
+Typically this function can simply be adapted from a main function already used by the single-scale simulation code, but initializing on the provided MPI\_Comm rather than MPI\_Comm\_World. 
+Other MPI calls in the code operating on MPI\_COMM\_WORLD should be replaced with AMSI\_COMM\_LOCAL which is the current MPI\_Comm associated with the AMSI Scale associated with the single-scale code.
+Multi-scale interaction points are determined and as wells as calls to create the appropriate DataDistributions and CouplingPatterns, which may be created during initialization for simplicity and modified later during execution.
+Calls to appropriately set DataDistribution values and modify CouplingPatterns should be added after adaptive procedures are conducted at the desired scale.
+AMSI does not treat any Scale as controlling any other Scale, thus care must be taken to avoid two Scales conducting adaptive procedures on the same CouplingPattern at the same time, though both Scales may conduct such procedures as long as all operations from adaptive procedures at one Scale are completed prior to starting adaptive procedures at the second Scale.
+This allows the CouplingPattern to reach a globally valid state prior to undergoing additional modification. 
+When a Scale has many scale-tasks operating it is necessary to write new code to manage the instantiation and management of the various scale-tasks, and update the relevant DataDistributions as required.
+
 \subsection{Parallel Operations}\label{paralle_ops}
@@
 
-%% revised up to here, with sporadic edits after
-
 All communication and planning operations in the AMSI libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer. 
 Usage only of implicit ordering was decided upon in order to reduce the memory impact as the simulation scales, as any meta-data which scales with the simulation will necessarily pose a memory issue once scaling hits some threshold.
+
+[major revisions end here, though there are sporadic edits herafter]
+%% revised up to here, with sporadic edits after
 

### 2016-01-25 amsi_util/amsiPermute.h
forgot to add files to a previous commit

@@ -1,0 +1,30 @@
+#ifndef AMSI_PERMUTE_H_
+#define AMSI_PERMUTE_H_
+#include <algorithm>
+#include <vector>
+namespace amsi
+{
+  template <typename T, typename OP>
+    struct PermuteComp
+  {
+  private:
+    T frst;
+    OP op;
+  public:
+    PermuteComp(T f, OP p)
+      : frst(f)
+      , op(p)
+    {}
+    bool operator()(unsigned a, unsigned b)
+    {
+      return op((*frst+a),(*frst+b));
+    }
+  };
+  template <typename T, typename OP>
+    PermuteComp<T,OP> permuteComp(T f, OP p)
+  {
+    return PermuteComp<T,OP>(f,p);
+  }
+  void combinePermutations(int * p1, int * p2, int sz);
+}
+#endif

### 2016-01-25 doc/articles/amsi_biotissue/paper.tex
modifying latex packages used to compiler paper, need to use 'pdflatex --shell-escape' in order for the eps figures to be converted to pdf, just add a .bash_profile alias for pdflatex to save the hastle

@@
 \usepackage{amsfonts}
+\usepackage{amsmath}
 \usepackage{graphicx}
+\usepackage[outdir=./]{epstopdf}
 \usepackage{listings}
-\usepackage{mathtools}
+%\usepackage{mathtools}
 \usepackage{url}
-\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
-\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
+%\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
+%\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
 
@@
 [Need a couple sentences to talk about how multi-scale is important]
+
 Implementation of multi-scale numerical simulations typically rely on either an ad-hoc set of algorithms designed for a particular problem, or on large frameworks and runtimes.

### 2016-01-25 amsi_meta/AddData_LoadBalancing.cc
renaming a bunch of stuff to simplify things

@@ -1,177 +1,0 @@
-#include "AddData_LoadBalancing.h"
-#include "amsiCoupling.h"
-#include "amsiPermute.h"
-#include <algorithm>
-#include <cassert>
-#include <iostream>
-#include <limits.h>
-#include <cstdlib>
-namespace amsi
-{
-  // Load balancing with added data (no migration)
-  // Updates pattern (current comm pattern) and init (initialization comm pattern)
-  //   by distributing the data counts specified in num
-  // This load balancing uses the default algorithm in CommPattern_CreateDistro to
-  //   essentially spew new data accross the comm pattern without regard for filling
-  //   lighter loaded processes. So essentially it only load balances the new data.
-  // This algorithm is efficient but doesn't really load balance
-  void addData_evenSpread(CommPattern * cp, CommPattern * dlta, int * num)
-  {
-    int s1 = cp->getNumSenders();
-    int s2 = cp->getNumRecvers();
-    int total_data = 0;
-    for(int ii = 0; ii < s1; ii++)
-      total_data += num[ii];
-    int even_distro = total_data / s2;
-    int extra_distro = total_data % s2;
-    int dst = 0;
-    int dst_rcv = even_distro + (dst < extra_distro);
-    for(int src = 0; src < s1; src++)
-    {
-      int src_count = num[src];
-      while(src_count > 0)
-      {
-        // if there are enough to send all needed to dstCommPattern_LoadBalance_Spread
-        if(src_count >= dst_rcv)
-        {
-          (*cp)(src,dst) += dst_rcv;
-          (*dlta)(src,dst) = dst_rcv;
-          src_count -= dst_rcv;
-          dst++;
-          dst_rcv = even_distro + (dst < extra_distro);
-        }
-        else if(src_count < dst_rcv)
-        {
-          (*cp)(src,dst) += src_count;
-          (*dlta)(src,dst) = src_count;
-          dst_rcv -= src_count;
-          src_count = 0;
-        }
-      }
-    }
-  }
-  // Load balancing with added data (no migration)
-  // Updates pattern (current comm pattern) and init (initialization comm pattern)
-  //   by distributing the data counts specified in num
-  // This load balancing algorithm finds the lightest loaded processes and
-  //   adds data to them first. This is the opposite end of the spectrum from
[…]

### 2016-01-25 amsi_meta/CommunicationManager.h

@@ -1,53 +1,0 @@
-#ifndef COMMUNICATION_MANAGER_H_
-#define COMMUNICATION_MANAGER_H_
-#include "unordered_map.h"
-#include "amsiCreateCoupling.h"
-#include "AddData_LoadBalancing.h"
-#include "PlanMigration.h"
-#include <map>
-#include <string>
-#include <utility>
-#include <vector>
-namespace amsi
-{
-    class TaskManager;
-    class CommPattern;
-    enum Role
-    {
-      SENDER = 0,
-      RECVER = 1
-    };
-    class CommunicationManager
-    {
-    public:
-      CommunicationManager();
-      size_t defineRelation(size_t id1, size_t id2);
-      bool CommRelation_Exists(size_t id1, size_t id2);
-      // will return null until the pattern for a relation is actually initialized
-      CommPattern * getCommPattern(size_t rdd_id);
-      std::pair<size_t,size_t> Relation_GetTasks(size_t r_id);
-    protected:
-      friend class ControlService;
-      size_t CommRelation_GetID(size_t id1, size_t id2);
-      MPI_Comm CommRelation_GetInterComm(size_t r_id);
-      size_t RelationDistribution_GetID(size_t r_id, DataDistribution * dd);
-      /// called prior to execution of tasks
-      void InitInterComms(TaskManager * tm); 
-      void CommPattern_Create(size_t rdd_id, DataDistribution * dd, int t1s, int t2s,CommPatternAlgo func);
-      void CommPattern_CreateRecv(size_t rdd_id, int t1s, int t2s);
-      void CommPattern_Invert(size_t n_rdd_id, size_t o_rdd_id);
-    private:
-      static int inter_comm_tag;
-      typedef std::map<size_t,MPI_Comm> rcommsmap_type;
-      rcommsmap_type relation_comms; // sender: comm
-      typedef std::multimap<size_t,size_t> relationmap_type;
-      relationmap_type relations;     // (sender_task, recver_task)
-      typedef std::map<size_t, std::pair<size_t,size_t> > rtmap_t;
-      rtmap_t rtmap;
-      // actual communication patterns created during execution
-      commpatternmap_type comm_patterns;
-      std::map<size_t,CommPatternAlgo> comm_pattern_build;
-    };
-} // namespace amsi
-
-#endif

### 2016-01-25 amsi_meta/PlanMigration.h

@@ -1,32 +1,0 @@
-#ifndef PLANMIGRATION_H_
-#define PLANMIFRATION_H_
-#include "AddData_LoadBalancing.h"
-#include "amsiMetaConfig.h"
-#include <amsiMPI.h>
-#include <vector>
-namespace amsi
-{
-  class DataDistribution;
-  class CommPattern;
-  void CommPattern_PlanMigration_Test(std::vector<int> & m_index,
-                                      std::vector<int> & m_send_to,
-                                      CommPattern * pattern,
-                                      const DataDistribution * dd,
-                                      int rank,
-                                      MPI_Comm task_comm);
-  void CommPattern_PlanMigration_Full(std::vector<int> & m_index,
-                                      std::vector<int> & m_send_to,
-                                      CommPattern * pattern,
-                                      const DataDistribution * dd,
-                                      int rank,
-                                      MPI_Comm task_comm);
-# ifdef ZOLTAN
-  void CommPattern_PlanMigration_Zoltan(std::vector<int> & m_index,
-                                        std::vector<int> & m_send_to,
-                                        CommPattern * pattern,
-                                        const DataDistribution * dd,
-                                        int rank,
-                                        MPI_Comm task_comm);
-# endif
-}
-#endif

### 2016-01-25 amsi_meta/Task.h

@@ -1,75 +1,0 @@
-#ifndef TASK_H_
-#define TASK_H_
-#include "amsiMetaConfig.h"
-#include "amsiHost.h"
-#ifdef CORE
-#include <PCU.h>
-#endif
-#include <amsiMPI.h>
-#include <map>
-#include <string>
-namespace amsi
-{
-  class ProcessSet;
-  class DataDistribution;
-  class Task;
-  typedef int(*ExecuteFunc)(int&,char**&,MPI_Comm);
-  int taskSize(Task * t);
-  class Task{
-  public:
-    Task(ProcessSet* p);
-    ~Task();
-    int localRank() {return local_rank;}
-    int localToGlobalRank(int);
-    int size();
-    const MPI_Comm comm() {return task_comm;}
-    const MPI_Group group() {return task_group;}
-    void setExecutionFunction(ExecuteFunc e) {exec = e;}
-    int execute(int& argc,char**& argv)
-    {
-      int result = 0;
-      // don't execute of the global rank is not valid
-      if(localToGlobalRank(localRank()) >= 0)
-      {
-#       ifdef CORE
-        PCU_Switch_Comm(task_comm);
-#       endif
-
-        result +=  (*exec)(argc,argv,task_comm);
-
-#       ifdef CORE
-        PCU_Switch_Comm(AMSI_COMM_WORLD);
-#       endif
-      }
-      return result;
-    };
-    size_t createDD(const std::string & nm);
-    bool verifyDD(const std::string & nm);
-    void setLocalDDValue(const std::string & nm, int ct);
-    int getLocalDDValue(size_t dd_id);
-    void assembleDD(const std::string & nm);
-    size_t getDD_ID(const std::string & nm);
-    DataDistribution * getDD(const std::string & dd_nm)
-    {
-      size_t dd_id = getDD_ID(dd_nm);
-      return getDD(dd_id);
-    }
-    DataDistribution * getDD(size_t dd_id)
-    {
-      return data[dd_id]; // HACKY check it actually exists first...
-    }
-    const ProcessSet * getProcessSet() { return proc; }
-  protected:
-    MPI_Comm task_comm;
-    MPI_Group task_group;
-    ProcessSet * proc;
-    typedef std::map<size_t,DataDistribution*> datamap_t;
-    datamap_t data;
-  private:
-    int rank;
-    int local_rank;
-    ExecuteFunc exec;
-    hash<std::string> id_gen;
-  };
-}
-#endif

### 2016-01-25 meta/CommunicationManager.h

@@ -1,0 +1,53 @@
+#ifndef COMMUNICATION_MANAGER_H_
+#define COMMUNICATION_MANAGER_H_
+#include "unordered_map.h"
+#include "amsiCreateCoupling.h"
+#include "amsiAddData.h"
+#include "PlanMigration.h"
+#include <map>
+#include <string>
+#include <utility>
+#include <vector>
+namespace amsi
+{
+    class TaskManager;
+    class CommPattern;
+    enum Role
+    {
+      SENDER = 0,
+      RECVER = 1
+    };
+    class CommunicationManager
+    {
+    public:
+      CommunicationManager();
+      size_t defineRelation(size_t id1, size_t id2);
+      bool CommRelation_Exists(size_t id1, size_t id2);
+      // will return null until the pattern for a relation is actually initialized
+      CommPattern * getCommPattern(size_t rdd_id);
+      std::pair<size_t,size_t> Relation_GetTasks(size_t r_id);
+    protected:
+      friend class ControlService;
+      size_t CommRelation_GetID(size_t id1, size_t id2);
+      MPI_Comm CommRelation_GetInterComm(size_t r_id);
+      size_t RelationDistribution_GetID(size_t r_id, DataDistribution * dd);
+      /// called prior to execution of tasks
+      void InitInterComms(TaskManager * tm); 
+      void CommPattern_Create(size_t rdd_id, DataDistribution * dd, int t1s, int t2s,CommPatternAlgo func);
+      void CommPattern_CreateRecv(size_t rdd_id, int t1s, int t2s);
+      void CommPattern_Invert(size_t n_rdd_id, size_t o_rdd_id);
+    private:
+      static int inter_comm_tag;
+      typedef std::map<size_t,MPI_Comm> rcommsmap_type;
+      rcommsmap_type relation_comms; // sender: comm
+      typedef std::multimap<size_t,size_t> relationmap_type;
+      relationmap_type relations;     // (sender_task, recver_task)
+      typedef std::map<size_t, std::pair<size_t,size_t> > rtmap_t;
+      rtmap_t rtmap;
+      // actual communication patterns created during execution
+      commpatternmap_type comm_patterns;
+      std::map<size_t,CommPatternAlgo> comm_pattern_build;
+    };
+} // namespace amsi
+
+#endif

### 2016-01-25 meta/amsiScale.cc
pulling in some files from the redevelopment branch, going to try to switch over to their version of functionality for some low-level amsi stuff and get rid of the various *Manager classes which are basically pointless and horrible

@@ -1,0 +1,44 @@
+#include "amsiScale.h"
+namespace amsi
+{
+  void Scale::assignRanks(RankSet * rs, MPI_Comm cm)
+  {
+    RankSet * ns = new DefaultRankSet;
+    merge(rnks,rs,ns);
+    delete rnks;
+    rnks = ns;
+    setValid(false);
+  }
+  bool Scale::validResign(RankSet * rs, MPI_Comm cm)
+  {
+    bool result = false;
+    DefaultRankSet intrsctn;
+    intersection(rnks,rs,&intrsctn);
+    if(intrsctn.size() == rs->size())
+      result = true;
+    return true;
+  }
+  void Scale::resignRanks(RankSet * rs, MPI_Comm cm)
+  {
+    if(validResign(rs,cm))
+    {
+      RankSet * ns = new DefaultRankSet;
+      diff(rnks,rs,ns);
+      delete rnks;
+      rnks = ns;
+      setValid(false);
+    }
+    else
+      std::cerr << "ERROR: invalid resignRanks attempt!" << std::endl;
+  }
+  bool assignedTo(Scale * scl)
+  {
+    bool result = false;
+    int rnk = -1;
+    MPI_Comm_rank(AMSI_COMM_WORLD,&rnk);
+    const RankSet * rs = scl->getRankSet();
+    if(rs->contains(rnk))
+      result = true;
+    return result;
+  }
+}

### 2016-01-25 meta/amsiScale.h

@@ -1,0 +1,76 @@
+#ifndef AMSI_SCALE_H_
+#define AMSI_SCALE_H_
+#include "amsiMetaConfig.h"
+#include <amsiDistributed.h>
+#include <amsiHost.h>
+#include <amsiMPI.h>
+#include <amsiRankSet.h>
+#include <amsiObserver.h>
+#include <map>
+#include <string>
+namespace amsi
+{
+  // responsible for maintaining the MPI_Comm used for scale-collective operations... thats it, other scale-related functionality should go elsewhere
+  class Scale : public Subject, public Distributed
+  {
+  protected:
+    RankSet * rnks;
+    MPI_Comm comm;
+    void reinitComm()
+    {
+      if(!isValid())
+      {
+        if(comm && comm != MPI_COMM_NULL)
+          MPI_Comm_free(&comm);
+        comm = makeComm(rnks);
+        setValid(true);
+        notify(SYNCHRONIZE);
+      }
+    }
+  public:
+    Scale()
+      : Subject()
+      , Distributed(false)
+      , rnks(new DefaultRankSet)
+      , comm()
+    { }
+    void assignRanks(RankSet * ps, MPI_Comm cm = AMSI_COMM_WORLD);
+    // possibly make validResign a protected member, and change resignRanks
+    //  to return bool if the resign cannot/did not succeed
+    bool validResign(RankSet * ps, MPI_Comm cm = AMSI_COMM_WORLD);
+    void resignRanks(RankSet * ps, MPI_Comm cm = AMSI_COMM_WORLD);
+    // collective on current rank set in amsi comm world
+    void synchronize() { reinitComm(); }
+    const RankSet * getRankSet() const {return rnks;}
+    const MPI_Comm getComm()
+    {
+      AMSI_DEBUG
+      (
+        if(!isValid())
+          std::cerr << "WARNING: retrieving communicator "
+                    << "from invalid Scale!" << std::endl;
+      )
+      return comm;
+    }
+  };
+  /**
+   * Determine whether the local process is assigned to the supplied Scale.
+   * @param sc The Scale to check assignment to
+   * @return Whether the local process is assigned to the Scale
+   * @note If the Scale is invalid the local process may be assigned to the Scale
+   *        but not belong to the MPI_Comm managed by the Scale. Prior to retrieving
+   *        the managed MPI_Comm from the Scale a call to .isValid() should be made
+   *        to determine the validity of the Scale, followed by a call to
[…]

### 2016-01-25 util/amsiFail.h

@@ -1,0 +1,15 @@
+#ifndef AMSI_FAIL_H_
+#define AMSI_FAIL_H_
+
+#include <limits>
+
+namespace amsi
+{
+# if __cpp_alias_templates >= 200704 // check to see if template aliasing is available in gcc
+  template <typename T> using fail = std::numeric_limits<T>::min();
+# else
+  template <typename T> T fail() {return std::numeric_limits<T>::min();}
+# endif
+}
+
+#endif

### 2016-01-25 util/amsiLocality.cc

@@ -1,0 +1,72 @@
+#include "amsiLocality.h"
+#include "amsiRankSet.h"
+#include <algorithm>
+namespace amsi
+{
+  hwloc_topology_t machine_topology;
+  uuid getNodeID()
+  {
+    uuid lid = 0;
+    char node_nm[MPI_MAX_PROCESSOR_NAME] = {};
+    int nm_len = 0;
+    MPI_Get_processor_name(node_nm, &nm_len);
+    std::string nm_str(node_nm);
+    lid = generateUUID(nm_str);
+    return lid;
+  }
+  void getRanksOnNode(MPI_Comm cm, uuid n_id, RankSet & rnks)
+  {
+    // init
+    int sz = 0;
+    MPI_Comm_size(cm,&sz);
+    int rnk = -1;
+    MPI_Comm_rank(cm,&rnk);
+    // retrieve
+    uuid pids[sz];
+    pids[rnk] = n_id;
+    MPI_Allgather(&n_id,1,mpi_type<uuid>(),
+                  &pids[0],1,mpi_type<uuid>(),
+                  cm);
+    // process
+    uuid * it = &pids[0];
+    while((it = std::find(it,pids+sz,n_id)) != pids+sz)
+    {
+      rnks.insert(std::distance(&pids[0],it));
+      it++;
+    }
+  }
+  void getRanksOnSocket(MPI_Comm cm, uuid n_id, RankSet & rnks)
+  {
+  }
+  void getNodeSet(MPI_Comm cm, std::vector<uuid> & nds)
+  {
+    int sz = 0;
+    MPI_Comm_size(cm,&sz);
+    int rnk = -1;
+    MPI_Comm_rank(cm,&rnk);
+    nds.resize(sz);
+    uuid n_id = getNodeID();
+    nds[rnk] = n_id;
+    MPI_Allgather(&n_id,1,mpi_type<uuid>(),
+                  &nds[0],1,mpi_type<uuid>(),
+                  cm);
+    std::sort(nds.begin(), nds.end());
+    std::vector<uuid>::iterator ne = std::unique(nds.begin(), nds.end());
+    nds.erase(ne,nds.end());
+  }
+  rank_t getNthRankOnNode(uuid nid, rank_t nth, MPI_Comm cm)
+  {
+    DefaultRankSet rnks;
+    getRanksOnNode(cm,nid,rnks);
+    std::vector<rank_t> on_nd;
[…]

### 2016-01-25 util/amsiRankSet.cc

@@ -1,0 +1,98 @@
+#include "amsiRankSet.h"
+#include "amsiPArray.h"
+#include "amsiUUID.h"
+#include <cassert>
+#include <algorithm>
+namespace amsi
+{
+  void SimpleRankSet::toArray(rank_t * rs) const
+  {
+    assert(rs);
+    std::copy(s.begin(),s.end(),rs);
+  }
+  RankSet * SimpleRankSet::copy()
+  {
+    RankSet * result = new SimpleRankSet(*this);
+    return result;
+  }
+  void intersection(const RankSet * a, const RankSet * b, RankSet * rslt)
+  {
+    int asz = a->size();
+    int bsz = b->size();
+    rank_t ar[asz];
+    rank_t br[bsz];
+    a->toArray(ar);
+    b->toArray(br);
+    int rsz = asz > bsz ? asz : bsz;
+    rank_t rs[rsz];
+    rank_t * nsz = std::set_intersection(ar,ar+asz,
+                                         br,br+bsz,
+                                         rs);
+    insertFrom(rslt,&rs[0],nsz);
+  }
+  void merge(const RankSet * a, const RankSet * b, RankSet * rslt)
+  {
+    rank_t asz = a->size();
+    rank_t bsz = b->size();
+    rank_t as[asz];
+    rank_t bs[bsz];
+    a->toArray(as);
+    b->toArray(bs);
+    insertFrom(rslt,&as[0],as+asz);
+    insertFrom(rslt,&bs[0],bs+bsz);
+  }
+  void diff(const RankSet * a, const RankSet * b, RankSet * rslt)
+  {
+    rank_t asz = a->size();
+    rank_t bsz = b->size();
+    rank_t as[asz];
+    rank_t bs[bsz];
+    a->toArray(as);
+    b->toArray(bs);
+    rank_t ns[asz];
+    rank_t * nsz = std::set_difference(as,as+asz,
+                                       bs,bs+bsz,
+                                       ns);
+    insertFrom(rslt,ns,nsz);
+  }
+  void getGlobalRanks(RankSet * rslt, MPI_Comm cm)
+  {
+    int sz = 0;
+    int rnk = -1;
+    int grnk = -1;
+    MPI_Comm_size(cm,&sz);
+    MPI_Comm_rank(cm,&rnk);
+    MPI_Comm_rank(AMSI_COMM_WORLD,&grnk);
+    int * rnks = new int[sz]();
[…]

### 2016-01-25 util/amsiRankSet.h

@@ -1,0 +1,42 @@
+#ifndef AMSI_RANK_SET_H_
+#define AMSI_RANK_SET_H_
+#include "amsiMPI.h"
+#include <set>
+namespace amsi
+{
+  class RankSet
+  {
+  public:
+    virtual rank_t size() const = 0;
+    virtual void insert(rank_t) = 0;
+    virtual bool contains(rank_t) const = 0;
+    virtual void toArray(rank_t *) const = 0;
+    virtual RankSet * copy() = 0;
+  };
+  class SimpleRankSet : public RankSet
+  {
+  protected:
+    std::set<rank_t> s;
+  public:
+    virtual rank_t size() const {return s.size();}
+    virtual void insert(rank_t r) {s.insert(r);}
+    virtual bool contains(rank_t r) const {return s.count(r) > 0;}
+    virtual void toArray(rank_t * rs) const;
+    virtual RankSet * copy();
+  };
+  typedef SimpleRankSet DefaultRankSet;
+  /// note: container must have operator[] member function
+  template <typename T>
+    void insertFromContainer(RankSet * rs, T & rnks);
+  template <typename T>
+    void pushToContainer(RankSet * rs, T begin, T end);
+  /// note: collective on the supplied communicator
+  void getGlobalRanks(RankSet*,MPI_Comm = AMSI_COMM_WORLD);
+  /// note: collective on the RankSet if using MPI3 standard
+  MPI_Comm makeComm(RankSet*);
+  void intersection(const RankSet * a, const RankSet * b, RankSet * rslt);
+  void merge(const RankSet * a, const RankSet * b, RankSet * rslt);
+  void diff(const RankSet * a, const RankSet * b, RankSet * rslt);
+}
+#endif
+#include "amsiRankSet_impl.h"

### 2016-01-25 util/amsiUUID.h

@@ -1,0 +1,30 @@
+#ifndef AMSI_UUID_H_
+#define AMSI_UUID_H_
+#include "amsiHost.h"
+#include <cstdlib>
+namespace amsi
+{
+  typedef size_t uuid;
+  template <typename T>
+    inline uuid generateUUID(const T & src)
+  {
+    static hash<T> hasher;
+    return hasher(src);
+  }
+  template <class T>
+    inline void combineUUID(uuid & a, T & b)
+  {
+    static hash<T> hasher;
+    a ^= hasher(b) + 0x9e3779b9 + (a << 6) + (a >> 2);
+  }
+  template <class T>
+    inline uuid generateRangeUUID(const T & begin,
+				  const T & end)
+  {
+    uuid result = 0;
+    for(T it = begin; it != end; it++)
+      combineUUID(result,*it);
+    return result;
+  }
+}
+#endif

### 2016-01-25 meta/ProcessAllocator.h
work on incorporating scales and reducing the role of the taskmanager

@@ -77,7 +77,6 @@ namespace amsi {
       return result;
     }
   };
-
 }
 
 #endif

### 2016-01-25 meta/TaskManager.cc

@@ -1,61 +1,25 @@
 #include "Task.h"
 #include "TaskManager.h"
-
 #include "amsiMeta.h"
 #include <amsiMPI.h>
-
 #include <cassert>
 #include <iostream>
-
 namespace amsi {
-
   Task * getLocal()
   {
     return tm->getLocalTask();
   }
-
-  TaskManager::TaskManager(MPI_Comm gc) :
-    id_gen(),
-    tasks(),
-    local_task(),
-    process_allocator(),
-    config_locked(false)
+  TaskManager::TaskManager(MPI_Comm gc)
+    : id_gen()
+    , tasks()
+    , local_task()
+    , process_allocator()
+    , config_locked(false)
   {
     int size = 0;
     MPI_Comm_size(gc,&size);
-
     process_allocator = static_cast<ProcessAllocator*>(new ExclusiveProcessAllocator(size));
   }
-  
-  TaskGroup * TaskManager::TaskGroup_Create(const std::string & nm)
-  {
-    TaskGroup * result = NULL;
-    size_t id = id_gen(nm);
-    if(taskgroups.find(id) != taskgroups.end())
-      result = taskgroups[id] = new TaskGroup();
-    return result;
-  }
-
-  TaskGroup * TaskManager::TaskGroup_Create(const std::string & nm,
-					    const std::vector<Task*>& t)
-  {
-    TaskGroup * result = NULL;
-    size_t id = id_gen(nm);
-    if(taskgroups.find(id) != taskgroups.end())
-      result = taskgroups[id] = new TaskGroup(t);
-    return result;
-  }
-
-  size_t TaskManager::TaskGroup_GetID(const std::string & nm) const
-  {
-    size_t result = 0;
-    result = id_gen(nm);
-    assert(taskgroups.find(result) != taskgroups.end());
-    return result;
-  }
-
-  
-  
   Task * TaskManager::createTask(const std::string & nm, int s)
   {
     Task * result = NULL;
@@ -65,32 +29,16 @@ namespace amsi {
     result = tasks[id] = new Task(assigned);
     return result;
   }
-  
-  // experimental
-  void TaskManager::Task_AddOverride(const std::string & nm, Task * t)
-  {
-    size_t id = id_gen(nm);
-    tasks[id] = t;
-  }
-  
-  bool TaskManager::Task_Exists(const std::string & nm)
-  {
-    std::map<size_t,Task*>::iterator it = tasks.find(id_gen(nm));
[…]

### 2016-01-25 meta/TaskManager.h

@@ -1,69 +1,46 @@
 #ifndef TASKMANAGER_H_
 #define TASKMANAGER_H_
-
 #include "Task.h"
-#include "TaskGroup.h"
 #include "ProcessSet.h"
 #include "ProcessAllocator.h"
 #include "amsiMetaConfig.h"
 #include "amsiHost.h"
-
+#include "amsiScale.h"
+#include "amsiLocality.h"
+#include <amsiUUID.h>
+#include <amsiRankSet.h>
 #include <mpi.h>
 #include <string>
-
-namespace amsi {
+namespace amsi
+{
   Task * getLocal();
   /// The definition of the TaskManager class, which is used primarily to
-  ///  create and manage tasks and generate unique ids for tasks and 
+  ///  create and manage tasks and generate unique ids for tasks and
   ///  data distributions defined on tasks, and to manage task states for
   ///  any local tasks.
   class TaskManager
-  { 
+  {
   public:
     TaskManager(MPI_Comm);
-    
-    TaskGroup * TaskGroup_Create(const std::string & nm);
-    TaskGroup * TaskGroup_Create(const std::string & nm, const std::vector<Task*>& t);
-    size_t TaskGroup_GetID(const std::string & nm) const;
-    TaskGroup * TaskGroup_Get(const std::string & nm);
-    TaskGroup * TaskGroup_Get(size_t id);
-    
     Task * createTask(const std::string & nm, int s);
-    
-    // experimental
-    void Task_AddOverride(const std::string & nm, Task * t);
-    
-    bool Task_Exists(const std::string & nm);
-    
     Task * getTask(const std::string & nm);
     Task * Task_Get(size_t id);
-    
     size_t getTaskID(const std::string & nm);
     Task * getLocalTask();
-    
-    size_t DataDist_GetID(const std::string & t_nm, 
-			  const std::string & dd_nm);
-    
+    size_t DataDist_GetID(const std::string & t_nm,
+                          const std::string & dd_nm);
     bool lockConfiguration();
     int Execute(int&,char**&);
   protected:
   private:
     TaskManager();
-    
     hash<std::string> id_gen;
     typedef std::map<size_t,Task*> taskmap_t;
     taskmap_t tasks;
-    
-    typedef std::map<size_t,TaskGroup*> taskgroupmap_t;
-    taskgroupmap_t taskgroups;
-    
     Task * local_task;
-
     ProcessAllocator * process_allocator;
[…]

### 2016-01-25 meta/amsiMeta.cc

@@ -7,11 +7,9 @@
 #include <pystring.h>
 #include <fstream>
 namespace amsi {
-
   TaskManager * tm = NULL;
   CommunicationManager * cm = NULL;
   FileSystemInfo * fs = NULL;
-
   std::string options_filename;
   bool parse_options(int argc,char ** argv)
   {
@@ -24,35 +22,30 @@ namespace amsi {
     {
       if(option == 'a')
       {
-	options_filename = optarg;
-	have_options = true;
+        options_filename = optarg;
+        have_options = true;
       }
     }
     optind = 0;
     opterr = 1;
     return have_options;
   }
-  
   void amsiMetaInit::amsiInit(int argc, char ** argv)
   {
     amsiUtilInit::amsiInit(argc,argv);
     if(parse_options(argc,argv))
       configureFromFile(options_filename);
   }
-
   void amsiMetaInit::amsiFree()
   {
     amsiUtilInit::amsiFree();
   }
-
   void configureFromFile(const std::string & filename)
   {
     std::fstream file(filename.c_str(),std::fstream::in);
     assert(file.is_open());
-
     tm = new TaskManager(AMSI_COMM_WORLD);
     cm = new CommunicationManager();
-    
     std::vector<std::string> tokens;
     std::string line;
     bool parsing_scales = true;
@@ -63,42 +56,41 @@ namespace amsi {
       line = pystring::strip(line);
       if(parsing_scales)
       {
-	if(pystring::startswith(line,std::string("#")))
-	{
-	  parsing_scales = false;
-	  parsing_relations = true;
-	  continue;
-	}
-	pystring::partition(line," ",tokens);
-	assert(tokens.size() == 3);
-	tm->createTask(tokens[0],atoi(tokens[2].c_str()));
+        if(pystring::startswith(line,std::string("#")))
+        {
+          parsing_scales = false;
+          parsing_relations = true;
+          continue;
+        }
+        pystring::partition(line," ",tokens);
+        assert(tokens.size() == 3);
+        tm->createTask(tokens[0],atoi(tokens[2].c_str()));
       }
       else if(parsing_relations)
       {
-	if(pystring::startswith(line,std::string("#")))
-	{
-	  parsing_relations=false;
-	  parsing_configure=true;
-	  continue;
-	}
-	pystring::partition(line," ",tokens);
-	assert(tokens.size() == 3);
-	size_t t1 = tm->getTaskID(tokens[0]);
[…]

### 2016-01-25 meta/amsiScale.cc

@@ -1,4 +1,5 @@
 #include "amsiScale.h"
+#include "amsiRankSet.h"
 namespace amsi
 {
   void Scale::assignRanks(RankSet * rs, MPI_Comm cm)
@@ -40,5 +41,11 @@ namespace amsi
     if(rs->contains(rnk))
       result = true;
     return result;
+  }
+  void assignFirstN(const RankSet * rs, rank_t n, Scale * sc)
+  {
+    DefaultRankSet a, b;
+    takeN(rs,n,&a,&b);
+    sc->assignRanks(&a);
   }
 }

### 2016-01-25 meta/amsiScale.h

@@ -64,6 +64,7 @@ namespace amsi
    *        .synchronize() in order to make the Scale valid (if necessary).
    */
   bool assignedTo(Scale * sc);
+  void assignFirstN(const RankSet * rs, rank_t n, Scale * sc);
   /**
    * Convenience function to get the size of the current rank set assigned to
    *  the provided scale.

### 2016-01-25 scripts/config.sh

@@ -3,27 +3,22 @@
 #
 # usage: ./config.sh [build_type]
 #
-
 if [ -z $1 ]; then
   BUILD_TYPE=Debug
 else
   BUILD_TYPE=$1
 fi
-
 if [ "$BUILD_TYPE" == "Debug" ]; then
   BUILD_DIR=../build_debug
 elif [ "$BUILD_TYPE" == "Release" ] ; then
   BUILD_DIR=../build_release
 fi
-
 if [ ! -d $BUILD_DIR ]; then
   mkdir $BUILD_DIR
 fi
 cd $BUILD_DIR
 rm -rf ./* #stupid and dangerous
-
 module load cmake
-
 HOSTNAME=`hostname`
 if [ "$HOSTNAME" == "q.ccni.rpi.edu" ]; then
   module load xl
@@ -33,9 +28,7 @@ if [ "$HOSTNAME" == "q.ccni.rpi.edu" ]; then
   module load /gpfs/u/home/PASC/PASCtbnw/barn-shared/petsc-3.5.2/arch-linux2-c-opt/lib/modules/3.5.2-arch-linux2-c-opt
   module load proprietary/simmetrix/simModSuite/10.0-150716
   module load boost
-
   export CMAKE_PREFIX_PATH=$CMAKE_PREFIX_PATH:/gpfs/u/home/PASC/PASCtbnw/barn-shared/install
-
    cmake \
     -DCMAKE_TOOLCHAIN_FILE=$CMAKE_XL_TOOLCHAIN \
     -DCMAKE_BUILD_TYPE=$BUILD_TYPE \
@@ -47,33 +40,28 @@ if [ "$HOSTNAME" == "q.ccni.rpi.edu" ]; then
     -DSIM_MPI=bgmpi \
     -DBOOST_INCLUDE_DIR=$BOOST_INCLUDE_DIR \
     ..
-
 else
     module load openmpi/1.3.3
     module load simmetrix/simModSuite
     module load $DEVROOT/petsc/petsc-3.6.3/ompi133-debug/lib/petsc/conf/modules/petsc/3.6.3
-
     cmake \
-	-DCMAKE_BUILD_TYPE=$BUILD_TYPE \
-	-DBUILD_TESTS=ON \
-	-DCMAKE_INSTALL_PREFIX=$DEVROOT/install/amsi/sim/openmpi-1.3.3/ \
[…]

### 2016-01-25 test/meta/CommunicationManager/test.cc

@@ -28,15 +28,15 @@ int main(int argc, char * argv[])
   MPI_Comm_size(MPI_COMM_WORLD,&size);
 
   std::cout << "Initializing test object(s):" << std::endl;
-
+/*
   // Create a task manager which uses the MPI_COMM_WORLD comm to create the global ProcessSet
   TaskManager * tm = new TaskManager(MPI_COMM_WORLD);
-
+  
   // Create a task using id "macro" consisting of 8 processes
-  tm->createTask("macro",8);
+  Task * macro = tm->createTask("macro",8);
   
   // Create a task using id "micro" consisting of 24 processes
-  tm->createTask("micro",24);
+  Task * micro = tm->createTask("micro",24);
 
   // Get the unique id's for the "macro" and "micro" tasks
   size_t t_id1 = tm->getTaskID("macro");
@@ -47,10 +47,10 @@ int main(int argc, char * argv[])
 
   // Make sure the CommunicationManager constructed
   failed += test_neq("CommunicationManager()",static_cast<CommunicationManager*>(NULL),cm);
-  
-  // Define a CommRelation between macro and micro and save the id 
+
+  // Define a CommRelation between macro and micro and save the id
   size_t r_id1 = cm->defineRelation(t_id1,t_id2);
-  
+
   // Test that the CommRelation id is valid
   failed += test_neq(".CommRelation_Define()",static_cast<size_t>(0),r_id1);
 
@@ -62,7 +62,7 @@ int main(int argc, char * argv[])
 
   failed += test(".Relation_GetTasks()",t_id1,t_ids.first);
   failed += test(".Relation_GetTasks()",t_id2,t_ids.second);
-
+*/
   test("Total Failed",0,failed);
 
   MPI_Finalize();

### 2016-01-25 test/meta/TaskManager/test.cc

@@ -1,96 +1,67 @@
 // in-build
 #include "../../test.h"
-
 // in-project
 #include "TaskManager.h"
-
 // standard
 #include <iostream>
 #include <utility>
-
 using namespace amsi;
-
 /// @brief Simple task-executing placeholed function for testing
-/// @param The number of command-line arguments 
+/// @param The number of command-line arguments
 /// @param The command-line arguments
 /// @param The comm on which this task is operating
 /// @return An integer, 0 if success, non-zero if fail
 int task1_run(int &, char **&, MPI_Comm)
 {
   int failed = 0;
-  
   // if this is even run the test succeeds
   failed += test(".Execute()",true,true);
   return failed;
 }
-
 /// test entry point
 int main(int argc, char * argv[])
 {
   int failed = 0;
   MPI_Init(&argc,&argv);
-
   int rank = -1;
   MPI_Comm_rank(MPI_COMM_WORLD,&rank);
-
   int size;
   MPI_Comm_size(MPI_COMM_WORLD,&size);
-
   std::cout << "Initializing test object(s):" << std::endl;
-
   // Create a taskmanger test object using the world comm as the primary parallel execution space
   TaskManager * tm = new TaskManager(MPI_COMM_WORLD);
-
   // Make sure the task manager was successfully constructed
   failed += test_neq("TaskManager()",
-		     static_cast<void*>(NULL),
-		     static_cast<void*>(tm));
-
+                     static_cast<void*>(NULL),
+                     static_cast<void*>(tm));
   // Create a task using the task manager using 8 processes
   Task * t1 = NULL;
   t1 = tm->createTask("test_task",8);
- 
   // Test whether the task manager successfully created the task
   failed += test_neq(".Task_Create()",static_cast<void*>(NULL),static_cast<void*>(t1));
-
   // Test whether the task manager successfully recongizes the newly-created task exists
-  failed += test(".Task_Exists()",true,tm->Task_Exists("test_task"));
-
+//  failed += test(".Task_Exists()",true,tm->Task_Exists("test_task"));
   // Test whether the task manager recongizes a non-existent task
-  failed += test(".Task_Exists()",false,tm->Task_Exists("test_task_fail"));
-
+  //failed += test(".Task_Exists()",false,tm->Task_Exists("test_task_fail"));
   // Retrieve a pointer to the newly-created task
   Task * t2 = NULL;
[…]

### 2016-01-25 util/amsiHost.h

@@ -1,8 +1,6 @@
 #ifndef AMSIHOST_H_
 #define AMSIHOST_H_
-
 #include "amsiUtilConfig.h"
-
 #ifdef BGQ
   #include <boost/unordered/unordered_map.hpp>
   #include <boost/functional/hash.hpp>

### 2016-01-25 util/amsiMPI.h

@@ -5,6 +5,7 @@
 #include <iostream>
 extern MPI_Comm AMSI_COMM_WORLD;
 extern MPI_Comm AMSI_COMM_LOCAL;
+typedef int rank_t;
 namespace amsi
 {
   template <typename T>
@@ -35,19 +36,19 @@ namespace amsi
     return mx;
   }
   template <typename T>
-    void send(T * msg, int to, int cnt = 1, MPI_Comm cm = AMSI_COMM_LOCAL)
+    void send(T * msg, rank_t to, int cnt = 1, MPI_Comm cm = AMSI_COMM_LOCAL)
   {
     MPI_Send(msg,cnt,mpi_type<T>(),to,0,cm);
   }
   template <typename T>
-    MPI_Request asend(T * msg, int to, int cnt = 1, MPI_Comm cm = AMSI_COMM_LOCAL)
+    MPI_Request asend(T * msg, rank_t to, int cnt = 1, MPI_Comm cm = AMSI_COMM_LOCAL)
   {
     MPI_Request rqst;
     MPI_Isend(msg,cnt,mpi_type<T>(),to,0,cm,&rqst);
     return rqst;
   }
   template <typename T>
-    MPI_Status recv(T * msg, int frm = MPI_ANY_SOURCE, int cnt = 1, MPI_Comm cm = AMSI_COMM_LOCAL)
+    MPI_Status recv(T * msg, rank_t frm = MPI_ANY_SOURCE, int cnt = 1, MPI_Comm cm = AMSI_COMM_LOCAL)
   {
     MPI_Status s;
     MPI_Recv(msg,cnt,mpi_type<T>(),frm,MPI_ANY_TAG,cm,&s);
@@ -55,7 +56,7 @@ namespace amsi
   }
   // test if there is a message from rank frm waiting to be recv'd return a sz > 0 if so
   template <typename T>
-    int arecv_sz(int & frm = MPI_ANY_SOURCE, MPI_Comm cm = AMSI_COMM_LOCAL)
+    int arecv_sz(rank_t & frm = MPI_ANY_SOURCE, MPI_Comm cm = AMSI_COMM_LOCAL)
   {
     int sz = 0;
     MPI_Status s;
@@ -68,7 +69,7 @@ namespace amsi
   }
   // asynchronously recv a message from frm with cnt elements
   template <typename T>
[…]

### 2016-01-25 util/amsiRankSet.cc

@@ -55,6 +55,16 @@ namespace amsi
                                        ns);
     insertFrom(rslt,ns,nsz);
   }
+  void takeN(const RankSet * src, rank_t nm, RankSet * a, RankSet * b)
+  {
+    int sz = src->size();
+    assert(nm < sz);
+    int rnks[sz];
+    src->toArray(rnks);
+    for(int ii = 0; ii < nm; ii++)
+      a->insert(rnks[ii]);
+    diff(src,a,b);
+  }
   void getGlobalRanks(RankSet * rslt, MPI_Comm cm)
   {
     int sz = 0;

### 2016-01-25 util/amsiRankSet.h

@@ -34,9 +34,11 @@ namespace amsi
   void getGlobalRanks(RankSet*,MPI_Comm = AMSI_COMM_WORLD);
   /// note: collective on the RankSet if using MPI3 standard
   MPI_Comm makeComm(RankSet*);
+  /// todo: these need to be rewritten to avoid use of toArray, which might mean moving them into the class or something similar, but serializing the rank set will scale very poorly
   void intersection(const RankSet * a, const RankSet * b, RankSet * rslt);
   void merge(const RankSet * a, const RankSet * b, RankSet * rslt);
   void diff(const RankSet * a, const RankSet * b, RankSet * rslt);
+  void takeN(const RankSet * src, rank_t nm, RankSet * a, RankSet * b);
 }
-#endif
 #include "amsiRankSet_impl.h"
+#endif

### 2016-01-26 scripts/config.sh
config.sh file changed to configure amsi in my directory.

@@ -51,7 +51,7 @@ if [ "$HOSTNAME" == "q.ccni.rpi.edu" ]; then
 else
     module load openmpi/1.3.3
     module load simmetrix/simModSuite
-    module load $DEVROOT/petsc/petsc-3.6.3/ompi133-debug/lib/petsc/conf/modules/petsc/3.6.3
+    module load $DEVROOT/petsc-3.6.3/ompi133-debug/lib/petsc/conf/modules/petsc/3.6.3
 
     cmake \
 	-DCMAKE_BUILD_TYPE=$BUILD_TYPE \
@@ -59,7 +59,7 @@ else
 	-DCMAKE_INSTALL_PREFIX=$DEVROOT/install/amsi/sim/openmpi-1.3.3/ \
 	-DCMAKE_C_COMPILER="mpicc" \
 	-DCMAKE_CXX_COMPILER="mpicxx" \
-	-DCORE_DIR=$DEVROOT/install/core-sim/openmpi-1.3.3 \
+	-DCORE_DIR=$DEVROOT/install/core/sim/openmpi-1.3.3 \
 	-DSIM_MPI=openmpi14 \
 	..
 

### 2016-01-26 doc/articles/amsi_biotissue/paper.tex

@@
 
-In order to avoid the introduction of unnecessary parallel barriers into the code, AMSI control decisions are only  during operations which are already collective over the set of processes effected by a given control decision.
-
 % need to expand this substantially
-Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{parker2006component} \cite{chopard2011framework} \cite{} and these frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah}.
+Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{davison2000uintah} \cite{parker2002component} \cite{chopard2011framework}.
+These frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah} \cite{parker20026component}.
+Efforts have also been made to provide standard commponent interfaces to encourage more widespread component adoption in such systems \cite{armstrong1999toward}.
 
@@
 Constructing a new multi-scale simulation from a set of legacy single-scale simulation codes involves several steps.
-Each Scale in the simulation must have a 'main-like' function with the signature \verb|int (int *, char *[], MPI_Comm)| which initializes the various supporting libraries used by that single-scale code on the provided MPI\_Comm and then begins conducting the simulation at that Scale.
+Each Scale in the simulation must have a 'main-like' function with the signature
+\begin{verbatim}
+int (*xxx_main)(int *, char *[], MPI_Comm)
+\end{verbatim}
+which initializes the various supporting libraries used by that single-scale code on the provided MPI\_Comm and then begins conducting the simulation at that Scale.
 Typically this function can simply be adapted from a main function already used by the single-scale simulation code, but initializing on the provided MPI\_Comm rather than MPI\_Comm\_World.
-Other MPI calls in the code operating on MPI\_COMM\_WORLD should be replaced with AMSI\_COMM\_LOCAL which is the current MPI\_Comm associated with the AMSI Scale associated with the single-scale code.
+Other MPI calls in the code operating on MPI\_COMM\_WORLD should be replaced with AMSI\_COMM\_LOCAL which is the current MPI\_Comm currently associated with the AMSI Scale.
 Multi-scale interaction points are determined and as wells as calls to create the appropriate DataDistributions and CouplingPatterns, which may be created during initialization for simplicity and modified later during execution.
@@
 When a Scale has many scale-tasks operating it is necessary to write new code to manage the instantiation and management of the various scale-tasks, and update the relevant DataDistributions as required.
+
+As noted, AMSI has a flexible definition of what constitutes a scale-task.
+This allows application developers flexibility in their implementation, but it also prevents certain optimizations -- such as task scheduling -- that a stronger definition of a scale-task can provide.
+For instance, the UINTAH framework \cite{davison200uintah} has a more well-defined task definition which allows the runtime to make scheduling optimizations to reduce process idle time and increase overall parallel performance \cite{meng2010dynamic}.
 
@@
 
-[major revisions end here, though there are sporadic edits herafter]
+\textbf{[major revisions end here, though there are sporadic edits herafter]}
 %% revised up to here, with sporadic edits after
@@
 This results in each micro-scale node being displaced.
-Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation, resulting in a high coordination number at each node and consequently a numerically stable fiber network.
+Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation, resulting in a high coordination number at each node and consequently a numerically stable global Jacobian.
 The micro-scale boundary condition necessitates that the force exerted by each fiber network node located on the boundary of the RVE be zero.
@@
 
-The simulation two nested Newton-Raphson operations occurring; during each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
+Thus the multi-scale simulation consists of two nested Newton-Raphson operations occurring; during each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
+Thus the micro-scale iterative processes will tend to dominate the total computational effort required to fully converge to a solution.
 
@@
 %
-\begin{lstlisting}[language=Bash,caption={process assignment calculation}]
+%\begin{lstlisting}[language=Bash,caption={process assignment calculation}]
+\begin{verbatim}
 NP=$(( $NUM_MACRO * 128))
 NUM_MICRO=$(( $NP - $NUM_MACRO))
-\end{lstlisting}
+\end{verbatim}
+%\end{lstlisting}
 \label{macro micro ratio}
@@
 As this fiber-matrix RVE represents a more significant source of computational complexity, it is independently parallelized -- in contrast to the fiber-only RVE -- and will represent a compelling use case for the parallel execution space discretization and scale-coupling communication management capabilities of the AMSI libraries.
+
+\section{Acknowledgements}
+
 \bibliography{citations}

### 2016-01-26 util/amsiLocality.cc
reworked the machine locality interface and documented

@@ -10,11 +10,30 @@ namespace amsi
     char node_nm[MPI_MAX_PROCESSOR_NAME] = {};
     int nm_len = 0;
     MPI_Get_processor_name(node_nm, &nm_len);
-    std::string nm_str(node_nm);
-    lid = generateUUID(nm_str);
+    //std::string nm_str(node_nm);
+    lid = generateUUID(std::string(node_nm));
     return lid;
   }
-  void getRanksOnNode(MPI_Comm cm, uuid n_id, RankSet & rnks)
+  void getNodeSet(std::vector<uuid> & nds, MPI_Comm cm)
+  {
+    int sz = 0;
+    MPI_Comm_size(cm,&sz);
+    int rnk = -1;
+    MPI_Comm_rank(cm,&rnk);
+    nds.resize(sz);
+    uuid n_id = getNodeID();
+    nds[rnk] = n_id;
+    MPI_Allgather(&n_id,1,mpi_type<uuid>(),
+                  &nds[0],1,mpi_type<uuid>(),
+                  cm);
+    std::sort(nds.begin(), nds.end());
+    std::vector<uuid>::iterator ne = std::unique(nds.begin(), nds.end());
+    nds.erase(ne,nds.end());
+  }
+  void getRanksOnSocket(uuid n_id, RankSet & rnks, MPI_Comm cm)
+  {
+  }
+  void getRanksOnNode(uuid n_id, RankSet & rnks, MPI_Comm cm)
   {
     // init
     int sz = 0;
@@ -35,29 +54,10 @@ namespace amsi
       it++;
     }
   }
-  void getRanksOnSocket(MPI_Comm cm, uuid n_id, RankSet & rnks)
-  {
-  }
-  void getNodeSet(MPI_Comm cm, std::vector<uuid> & nds)
-  {
-    int sz = 0;
-    MPI_Comm_size(cm,&sz);
-    int rnk = -1;
-    MPI_Comm_rank(cm,&rnk);
-    nds.resize(sz);
-    uuid n_id = getNodeID();
-    nds[rnk] = n_id;
-    MPI_Allgather(&n_id,1,mpi_type<uuid>(),
-                  &nds[0],1,mpi_type<uuid>(),
-                  cm);
-    std::sort(nds.begin(), nds.end());
-    std::vector<uuid>::iterator ne = std::unique(nds.begin(), nds.end());
-    nds.erase(ne,nds.end());
-  }
[…]

### 2016-01-26 util/amsiLocality.h

@@ -8,15 +8,60 @@
 namespace amsi
 {
   extern hwloc_topology_t machine_topology;
-  /// brief : get a uuid associated with the hardware processor associated with the local mpi rank
-  uuid getNodeID();
-  // brief : get all mpi ranks in comm associated with the hardware processor p_id
   class RankSet;
-  // todo (m) : consider moving MPI_Comm to the last parameter and having a default value of AMSI_COMM_WORLD
-  void getRanksOnSocket(MPI_Comm c, uuid p_id, RankSet & rnks);
-  void getRanksOnNode(MPI_Comm c, uuid n_id, RankSet & ranks);
-  void getNodeSet(MPI_Comm c, std::vector<uuid> & nodes);
-  void getNthRankOnNode(uuid n_id, rank_t nth, rank_t rnk, MPI_Comm = AMSI_COMM_WORLD);
-  void getNthRankOnNodes(const std::vector<uuid> & nids, rank_t nth, RankSet * rnks, MPI_Comm = AMSI_COMM_WORLD);
+  /**
+   * Get a uuid (misnomer) which identifies the hardware processor
+   *  the local mpi rank is executing on. This is based on hashing
+   *  MPI_Get_processor_name which by the standard gives a unique
+   *  identifying string, so barring a hash conflict this should be
+   *  fine.
+   * @todo Should this be getSocketID?
+   */
+  uuid getNodeID();
+  /**
+   * Get a vector containing the uuid of every hardware processor
+   *  in the specified comm in no particular order.
+   * @param nodes A vector containing the set of unique hardware processors
+   *              belonging to the specified comm. Should be empty when passed.
+   * @param c The mpi comm over which to perform the operation
+   * @note This call is collective over the specified comm.
+   * @todo Should this be getSocketSet?
+   */
+  void getNodeSet(std::vector<uuid> & nodes, MPI_Comm c = AMSI_COMM_WORLD);
+  // brief : get all mpi ranks in comm associated with the hardware processor p_id
+  /**
+   * Get the mpi ranks of the processes executing on the specific
+   * hardware processor specified.
+   * @note Currently unimplemented
+   * @todo implement!
+   */
+  void getRanksOnSocket(uuid p_id, RankSet & rnks, MPI_Comm c = AMSI_COMM_WORLD);
+  /**
+   * Get the mpi ranks (in the specified comm) of the processes executing
+   *  on the specific machine node specified.
+   * @param n_id The id of the node to retrieve the ranks executing on.
[…]

### 2016-01-26 meta/amsiFlow.h
mostly a bit a speculative work towards future developments, nothing too concrete

@@ -1,0 +1,34 @@
+#ifndef AMSI_FLOW_H_
+#define AMSI_FLOW_H_
+namespace amsi
+{
+  // fsm of scale status
+  enum ScaleState
+  {
+    READY_IDLE,
+    READY_ACTIVE,
+    BLOCK_IDLE,
+    BLOCK_ACTIVE,
+    RUNNING,
+    STT_CNT
+  };
+  enum ScaleTransition
+  {
+    ACTIVATE,
+    IDLE,
+    RELEASE,
+    TRNS_CNT
+  };
+  int scale_status[STT_CNT][TRNS_CNT] = {
+    {READY_ACTIVE,READY_IDLE,READY_IDLE},    // ready idle
+    {RUNNING,READY_IDLE,READY_ACTIVE},       // ready active
+    {BLOCK_ACTIVE,BLOCK_IDLE,READY_IDLE},    // block idle
+    {BLOCK_ACTIVE,BLOCK_IDLE,READY_ACTIVE},  // block active
+    {RUNNING,BLOCK_ACTIVE,READY_ACTIVE}      // running
+  }
+  ScaleState flowTransition(ScaleState stt, ScaleTransition trnstn)
+  {
+    return scale_status[stt][trsntn];
+  }
+}
+#endif

### 2016-01-26 meta/CommunicationManager.cc
mostly a bit a speculative work towards future developments, nothing too concrete

@@ -5,180 +5,156 @@
 #include "TaskManager.h"
 namespace amsi
 {
-    // assumes static initialization of communication relations
-    int CommunicationManager::inter_comm_tag = 0;
+  // assumes static initialization of communication relations
+  int CommunicationManager::inter_comm_tag = 0;
 
-    /// @brief Standard constructor.
-    CommunicationManager::CommunicationManager() :
-      relation_comms(),
-      relations(),
-      rtmap(),
-      comm_patterns(),
-      comm_pattern_build()
-    {}
-
-    /// @brief Defines a communication relationship between two tasks, should
-    ///        only be called during initialization and adaptation phases of
-    ///        AMSI operation.
-    /// @param id1 Identifier for the sending task in the CommRelation
-    /// @param is2 Identifier for the recving task in the CommRelation
-    /// @return size_t Identifier for the CommRelation
-    size_t CommunicationManager::defineRelation(size_t id1, size_t id2)
+  /// @brief Standard constructor.
+  CommunicationManager::CommunicationManager()
+    : relation_comms()
+    , relations()
+    , rtmap()
+    , comm_patterns()
+    , comm_pattern_build()
+  {}
+  /// @brief Defines a communication relationship between two tasks, should
+  ///        only be called during initialization and adaptation phases of
+  ///        AMSI operation.
+  /// @param id1 Identifier for the sending task in the CommRelation
+  /// @param is2 Identifier for the recving task in the CommRelation
+  /// @return size_t Identifier for the CommRelation
+  size_t CommunicationManager::defineRelation(size_t id1, size_t id2)
+  {
+    size_t result = 0;
+    // check if relation already exists
+    if(!CommRelation_Exists(id1,id2))
     {
-      size_t result = 0;
-      // check if relation already exists
-      if(!CommRelation_Exists(id1,id2))
+      relations.insert(std::make_pair(id1,id2));
+      result = CommRelation_GetID(id1,id2);
+      rtmap[result] = std::make_pair(id1,id2);
+    }
+    return result;
+  }
+  /// @brief Retrieve a CommPattern by id
+  /// @param rdd_id The id of the CommPattern
+  /// @return CommPattern* Pointer to the identified CommPattern or NULL if
+  ///                      no such pattern exists
+  CommPattern * CommunicationManager::getCommPattern(size_t rdd_id)
+  {
+    CommPattern * result = NULL;
+    if(comm_patterns.find(rdd_id) != comm_patterns.end())
[…]

### 2016-01-26 meta/CommunicationManager.h

@@ -10,44 +10,44 @@
 #include <vector>
 namespace amsi
 {
-    class TaskManager;
-    class CommPattern;
-    enum Role
-    {
-      SENDER = 0,
-      RECVER = 1
-    };
-    class CommunicationManager
-    {
-    public:
-      CommunicationManager();
-      size_t defineRelation(size_t id1, size_t id2);
-      bool CommRelation_Exists(size_t id1, size_t id2);
-      // will return null until the pattern for a relation is actually initialized
-      CommPattern * getCommPattern(size_t rdd_id);
-      std::pair<size_t,size_t> Relation_GetTasks(size_t r_id);
-    protected:
-      friend class ControlService;
-      size_t CommRelation_GetID(size_t id1, size_t id2);
-      MPI_Comm CommRelation_GetInterComm(size_t r_id);
-      size_t RelationDistribution_GetID(size_t r_id, DataDistribution * dd);
-      /// called prior to execution of tasks
-      void InitInterComms(TaskManager * tm); 
-      void CommPattern_Create(size_t rdd_id, DataDistribution * dd, int t1s, int t2s,CommPatternAlgo func);
-      void CommPattern_CreateRecv(size_t rdd_id, int t1s, int t2s);
-      void CommPattern_Invert(size_t n_rdd_id, size_t o_rdd_id);
-    private:
-      static int inter_comm_tag;
-      typedef std::map<size_t,MPI_Comm> rcommsmap_type;
-      rcommsmap_type relation_comms; // sender: comm
-      typedef std::multimap<size_t,size_t> relationmap_type;
-      relationmap_type relations;     // (sender_task, recver_task)
-      typedef std::map<size_t, std::pair<size_t,size_t> > rtmap_t;
-      rtmap_t rtmap;
-      // actual communication patterns created during execution
-      commpatternmap_type comm_patterns;
-      std::map<size_t,CommPatternAlgo> comm_pattern_build;
-    };
+  class TaskManager;
+  class CommPattern;
+  enum Role
+  {
+    SENDER = 0,
+    RECVER = 1
+  };
+  class CommunicationManager
+  {
+  public:
+    CommunicationManager();
+    size_t defineRelation(size_t id1, size_t id2);
+    bool CommRelation_Exists(size_t id1, size_t id2);
+    // will return null until the pattern for a relation is actually initialized
+    CommPattern * getCommPattern(size_t rdd_id);
[…]

### 2016-01-26 meta/Task.cc

@@ -15,7 +15,7 @@ namespace amsi
   }
   /// @brief Standard constructor, collective on the ProcessSet passed in
   /// @param p The ProcessSet describing the parallel ranks on which the Task will execute
-  Task::Task(ProcessSet * p) 
+  Task::Task(ProcessSet * p)
     : task_comm()
     , task_group()
     , proc(p)
@@ -23,34 +23,28 @@ namespace amsi
     , rank()
     , local_rank()
     , id_gen()
-    {
-      // Create a comm for the task
-      MPI_Group world_group;
-      MPI_Comm_group(MPI_COMM_WORLD,&world_group);
-
-      int n = p->size();
-      int pr[n];
-      for(int ii = 0; ii < n; ii++)
-	pr[ii] = (*p)[ii];
-
-      MPI_Group_incl(world_group,n,pr,&task_group);
-      MPI_Comm_create(MPI_COMM_WORLD,task_group,&task_comm);
-
-      MPI_Comm_rank(MPI_COMM_WORLD,&rank);
-      if(task_comm != MPI_COMM_NULL)
-        MPI_Comm_rank(task_comm,&local_rank);
-      else
-	local_rank = proc->indexOf(rank);
-    }
-
-    Task::~Task()
-    {
-      delete proc;
-
-      for(datamap_t::iterator it = data.begin(), itend = data.end(); it != itend; it++)
-	delete (it->second);
-    }
-  
+  {
+    // Create a comm for the task
+    MPI_Group world_group;
+    MPI_Comm_group(MPI_COMM_WORLD,&world_group);
+    int n = p->size();
+    int pr[n];
+    for(int ii = 0; ii < n; ii++)
+      pr[ii] = (*p)[ii];
+    MPI_Group_incl(world_group,n,pr,&task_group);
+    MPI_Comm_create(MPI_COMM_WORLD,task_group,&task_comm);
+    MPI_Comm_rank(MPI_COMM_WORLD,&rank);
+    if(task_comm != MPI_COMM_NULL)
+      MPI_Comm_rank(task_comm,&local_rank);
+    else
+      local_rank = proc->indexOf(rank);
+  }
+  Task::~Task()
+  {
+    delete proc;
+    for(datamap_t::iterator it = data.begin(), itend = data.end(); it != itend; it++)
+      delete (it->second);
+  }
 # ifdef ZOLTAN
   // hacky hard-coded bullshit
[…]

### 2016-01-26 meta/amsiScale.cc

@@ -1,5 +1,7 @@
 #include "amsiScale.h"
 #include "amsiRankSet.h"
+#include <cassert>
+#include <numeric>
 namespace amsi
 {
   void Scale::assignRanks(RankSet * rs, MPI_Comm cm)
@@ -47,5 +49,26 @@ namespace amsi
     DefaultRankSet a, b;
     takeN(rs,n,&a,&b);
     sc->assignRanks(&a);
+  }
+  void assignCounts(const RankSet * rs, rank_t nm, rank_t * cnts, Scale * scls[])
+  {
+    DefaultRankSet rmndr[nm];
+    DefaultRankSet sbsts[nm];
+    takeN(rs,cnts[0],&sbsts[0],&rmndr[0]);
+    scls[0]->assignRanks(&sbsts[0]);
+    for(int ii = 1; ii < nm; ii++)
+    {
+      takeN(&rmndr[ii-1],cnts[ii],&sbsts[ii],&rmndr[ii]);
+      scls[ii]->assignRanks(&sbsts[ii]);
+    }
+  }
+  void assignRatios(const RankSet * rs, rank_t cnt, double * rts, Scale * scls[])
+  {
+    rank_t sz = rs->size();
+    rank_t cnts[cnt];
+    for(int ii = 0; ii < cnt; ii++)
+      cnts[ii] = (rank_t)(((double)sz)*rts[ii]);
+    assert(std::accumulate(&cnts[0],&cnts[cnt-1],0) == sz);
+    assignCounts(rs,cnt,&cnts[0],scls);
   }
 }

### 2016-01-27 meta/ProcessSet.cc
cleanup of process set code, and examination of responsibilities of the process set for eventual removal in lieu of the new RankSet designed for dynamic process reallocation

@@ -1,73 +1,61 @@
 #include "ProcessSet.h"
-
 #include <iostream>
-
-namespace amsi {
-
-    /// @brief Default constructor for a range ProcessSet, takes low and high ranks
-    /// @param l The first process rank in the set
-    /// @param h One over the last process rank in the set
-    ProcessSet_T<std::pair<int,int> >::ProcessSet_T(int l, int h) :
-      val(std::make_pair(l,h))
-    {}
-
-
-    ProcessSet_T<std::pair<int,int> >::ProcessSet_T(std::pair<int,int> r) :
-      val(r)
-    {}
-
-    /// @brief Constructor taking a std::pair<int,int> range instead of high and low values
-    /// @param r The pair representing the range for the process set
-    ProcessSet_T<std::pair<int,int> >::ProcessSet_T(std::pair<int,int> & r) :
-      val(r)
-    {}
-    
-    /// @brief Retrieve process ranks from the set
-    /// @param i The index of the rank the retrieve
-    /// @return int The i-th rank in the set
-    int ProcessSet_T<std::pair<int,int> >::operator[](int i) const
+namespace amsi
+{
+  /// @brief Default constructor for a range ProcessSet, takes low and high ranks
+  /// @param l The first process rank in the set
+  /// @param h One over the last process rank in the set
+  ProcessSet_T<std::pair<int,int> >::ProcessSet_T(int l, int h)
+    : val(std::make_pair(l,h))
+  {}
+  ProcessSet_T<std::pair<int,int> >::ProcessSet_T(std::pair<int,int> r)
+    : val(r)
+  {}
+  /// @brief Constructor taking a std::pair<int,int> range instead of high and low values
+  /// @param r The pair representing the range for the process set
+  ProcessSet_T<std::pair<int,int> >::ProcessSet_T(std::pair<int,int> & r)
+    : val(r)
+  {}
+  /// @brief Retrieve process ranks from the set
+  /// @param i The index of the rank the retrieve
+  /// @return int The i-th rank in the set
+  int ProcessSet_T<std::pair<int,int> >::operator[](int i) const
+  {
+    return i < 0 ? -1 : i <= (val.second - val.first) ? val.first + i : -1;
+  }
+  /// @brief How many ranks are in the set
+  /// @return int The set size
+  int ProcessSet_T<std::pair<int,int> >::size() const
+  {
+    //std::cout << val.first << " " << val.second << std::endl;
[…]

### 2016-01-27 meta/ProcessSet.h

@@ -1,58 +1,52 @@
 #ifndef PROCESSSET_H_
 #define PROCESSSET_H_
-
 #include <utility>
-
-namespace amsi {
-
-    /// The interface specification to which all ProcessSet classes must adhere.
-    /// A ProcessSet simply represents a mathematical set of process ranks, and allows
-    ///  several operations on them.
-    class ProcessSet{
-    public: 
-      virtual int operator[](int) const = 0;
-      virtual int size() const = 0;
-      virtual bool isIn(int) const = 0;
-      virtual int indexOf(int) const= 0;
-      virtual ProcessSet * extract(int) = 0;
-    };
-
-    /// A templated subclass of the ProcessSet abstract interface, used primarily 
-    ///  to limit inheritance polymorphism 
-    template <typename T>
+namespace amsi
+{
+  /// The interface specification to which all ProcessSet classes must adhere.
+  /// A ProcessSet simply represents a mathematical set of process ranks, and allows
+  ///  several operations on them.
+  class ProcessSet
+  {
+  public:
+    virtual int operator[](int) const = 0;
+    virtual int size() const = 0;
+    virtual bool isIn(int) const = 0;
+    virtual int indexOf(int) const= 0;
+    virtual ProcessSet * extract(int) = 0;
+  };
+  /// A templated subclass of the ProcessSet abstract interface, used primarily
+  ///  to limit inheritance polymorphism
+  template <typename T>
     class ProcessSet_T : public ProcessSet
-    {
-    public:
-      ProcessSet_T(T&);
-      virtual int operator[](int) const;
-      virtual int size() const;
-      virtual bool isIn(int) const;
-      virtual int indexOf(int) const;
-      virtual ProcessSet * extract(int);
-    private:
-      T val;
-    };
-
-    /// A partial template specialization of the templated subclass which uses
-    ///  a pair<T,T> as the primary set specifier to specify sets representing
-    ///  ranges of process ranks.
-    template <>
+  {
+  public:
+    ProcessSet_T(T&);
+    virtual int operator[](int) const;
+    virtual int size() const;
+    virtual bool isIn(int) const;
+    virtual int indexOf(int) const;
+    virtual ProcessSet * extract(int);
+  private:
+    T val;
+  };
+  /// A partial template specialization of the templated subclass which uses
+  ///  a pair<T,T> as the primary set specifier to specify sets representing
[…]

### 2016-01-27 doc/articles/amsi_biotissue/paper.tex
cleanup of process set code, and examination of responsibilities of the process set for eventual removal in lieu of the new RankSet designed for dynamic process reallocation

@@
 
-Typically simulation flow control is determined solely by a runtime system.
-A framework may also limit specific features of the numerical implementation to those provided by the framework.
-This includes domain discretization models and associated adaptive processes -- structured and unstructured meshing and associated refinement algorithms, particle-in-cell methods, etc.
+%Typically simulation flow control is determined solely by a runtime system.
+%A framework may also limit specific features of the numerical implementation to those provided by the framework.
+%This includes domain discretization models and associated adaptive processes -- structured and unstructured meshing and associated refinement algorithms, particle-in-cell methods, etc.
 -- convergence algorithms for both the minimization of the relevant residuals (Newton-Raphson, etc.) as well as iterative solution of the resulting linear system (KSP, GMRES, ILU, etc.).

### 2016-01-27 interface/SimFEA.cc

@@ -1,15 +1,11 @@
 #include "SimFEA.h"
 #include <cassert>
-
 namespace amsi {
   namespace Analysis {
-    
     template <>
     void Model_PrintInfo<pGModel>(pGModel model, std::ostream & out)
     {
-
     }
-
     template <>
     void Mesh_PrintInfo<pMesh>(pMesh mesh, std::ostream & out)
     {
@@ -19,49 +15,47 @@ namespace amsi {
       out << "\t" << M_numEdges(mesh) << " edges." << std::endl;
       out << "\t" << M_numVertices(mesh) << " vertices." << std::endl;
     }
-
-   void Model_GetClassifiedEntities(pMesh mesh,
-				    pGEntity entity,
-				    int dim,
-				    std::list<pEntity> & classified)
+    void Model_GetClassifiedEntities(pMesh mesh,
+                                     pGEntity entity,
+                                     int dim,
+                                     std::list<pEntity> & classified)
     {
       switch(dim)
       {
       case 3:
       {
-	RIter it = M_classifiedRegionIter(mesh,entity);
-	pRegion pr;
-	while ((pr = RIter_next(it))) classified.push_back((pEntity)pr);
-	RIter_delete(it);
-	break;
+        RIter it = M_classifiedRegionIter(mesh,entity);
+        pRegion pr;
+        while ((pr = RIter_next(it))) classified.push_back((pEntity)pr);
+        RIter_delete(it);
+        break;
       }
       case 2:
       {
-	FIter it = M_classifiedFaceIter(mesh,entity,1);
-	pFace pf;
-	while ((pf = FIter_next(it))) classified.push_back((pEntity)pf);
-	FIter_delete(it);
-	break;
+        FIter it = M_classifiedFaceIter(mesh,entity,1);
+        pFace pf;
+        while ((pf = FIter_next(it))) classified.push_back((pEntity)pf);
+        FIter_delete(it);
+        break;
       }
       case 1:
       {
-	EIter it = M_classifiedEdgeIter(mesh,entity,1);
-	pEdge pe;
-	while ((pe = EIter_next(it))) classified.push_back((pEntity)pe);
-	EIter_delete(it);
-	break;
+        EIter it = M_classifiedEdgeIter(mesh,entity,1);
+        pEdge pe;
+        while ((pe = EIter_next(it))) classified.push_back((pEntity)pe);
+        EIter_delete(it);
+        break;
       }
       case 0:
[…]

### 2016-01-27 doc/articles/amsi_biotissue/paper.tex
bit of work on the paper, also adding some classes taken from the redevelopment branch for implementing a simple nonlinear analysis

@@
 Some approaches begin with the specification of multi-scale contributions during the specification of mathematical model.
-However, multi-scale influences can be incorporated in a strictly numerical or computational sense, by combining implementations of established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
+However, multi-scale influences can be incorporated in a strictly numerical or computational sense, by combining implementations of established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97,weinan2003heterogenous}.
 
@@
 Numerical multi-scale simulations fall broadly into two parallel paradigms, distinct from the choice of subdomain assigned to fine scale simulation.
-In the sequential multi-scale paradigm, fine-scale computation is used to precompute quantities of interest for establishing an engineering-scale physical model, possibly using a lookup table or some other method to incorporate the multi-scale influences into the engineering scale \cite{weinan2011principles} \cite{garcia2008sequential}.
-In the concurrent multi-scale paradigm, all scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{weinan2011principles} \cite{zeng2010concurrent}.
+In the sequential multi-scale paradigm, fine-scale computation is used to precompute quantities of interest for establishing an engineering-scale physical model, possibly using a lookup table or some other method to incorporate the multi-scale influences into the engineering scale \cite{weinan2011principles, garcia2008sequential}.
+In the concurrent multi-scale paradigm, all scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{weinan2011principles,zeng2010concurrent}.
 
@@
 For example a micro-scale simulation will typically be implemented with units of $\mu$m while a macro-scale simulation might be implemented using units of cm.
-Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales \cite{weinan2003heterogenous}, such as dimensionalizing terms calculated in a dimensionless domain \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.
+Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales \cite{weinan2003heterogenous}, such as dimensionalizing terms calculated in a dimensionless domain \cite{stylianopoulos2007volume,chandran2007deterministic}.
 
@@
 \section{Adaptive Multi-scale Simulation Infrastructure}\label{AMSI}
-The Adaptive Multi-scale Simulation Infrastructure (AMSI) is a set of libraries and tools developed by the Scientific Computation Research Center (SCOREC) at Rensselaer Polytechnic Institute.
-AMSI is designed to support the implementation and execution of dynamic concurrent multi-scale simulations on massively-parallel HPC machines.
-AMSI provides support for several key operations needed for concurrent multi-scale simulations: modelling and managing scales, scale-coupling parallel communication, adaptively adding/removing scale-tasks, and load-balancing coupled scales.
-A key goal in the design and implementation of these operations is a focus on general applicability and avoiding adherence to specific interfaces.
+The Adaptive Multi-scale Simulation Infrastructure (\verb|AMSI|) is a set of libraries and tools developed by the Scientific Computation Research Center (SCOREC) at Rensselaer Polytechnic Institute.
+\verb|AMSI| is designed to support the implementation and execution of dynamic concurrent multi-scale simulations on massively-parallel HPC machines.
+Services supported by \verb|AMSI| are based on constructing an internal model of the parallel structure of the multi-scale simulation.
+The most critical of these services is support for parallel scale-coupling communication between unrelated single-scale simulation codes.
+Additionally, \verb|AMSI| facilitates the use of adaptive processes on coupled scales while maintaining the scale-coupling relationships associated with the adapted scale.
 
-The interfaces provided by the AMSI libraries are intended to facilitate the coupling of functional single-scale simulation codes.
+A key goal in the design and implementation of these operations is a focus on general applicability.
+As such \verb|AMSI| places no requirements on the interfaces exposed by a single-scale simulation code being incorporated into a multi-scale simulation, outside of a single 'main-like' function which initializes the single-scale simulation using a supplied \verb|MPI_COMM|.
+The interfaces provided by the \verb|AMSI| libraries are intended to facilitate the coupling of functional single-scale simulation codes.
 This approach is used in order to mitigate development overhead associated with casting a simulation into a set of framework-specific components, or developing an entire ad-hoc set of multi-scale operations each time a new multi-scale code is being developed.
-Several key multi-scale operations are planned and enacted by the AMSI libraries, allowing the application programmer (a programmer using the AMSI libraries to implement a new multi-scale code or interaction) to determine the most effective use of the tools provided.
+Several key multi-scale operations are planned and enacted by the \verb|AMSI| libraries, allowing the application programmer (a programmer using the \verb|AMSI| libraries to implement a new multi-scale code or interaction) to determine the most effective use of the tools provided.
 All interfaces are designed to be physics-agnostic, operating only on computational quantities with no meta-model of the relation between a coupling term and the mathematical derivation of the multi-scale interaction.
 
-AMSI operates by maintaining a set of minimal simulation metadata in order to model various quantities of interest during simulation execution.
-AMSI provides mechanisms to assist in the implementation of run-time decision making and control decisions related to multi-scale execution, with a focus on allowing such adaptive procedures to be planned and implemented in a decentralized manner.
+\verb|AMSI| operates by maintaining a set of minimal simulation metadata in order to model various quantities of interest during simulation execution.
+\verb|AMSI| provides mechanisms to assist in the implementation of run-time decision making and control decisions related to multi-scale execution, with a focus on allowing such adaptive procedures to be planned and implemented in a decentralized manner.
 This design approach allows application developers to avoid centralized control decision bottlenecks.
@@
 % need to expand this substantially
-Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{davison2000uintah} \cite{parker2002component} \cite{chopard2011framework}.
-These frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah} \cite{parker20026component}.
+Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{davison2000uintah,parker2002component,chopard2011framework}.
+These frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah,parker2002component}.
 Efforts have also been made to provide standard commponent interfaces to encourage more widespread component adoption in such systems \cite{armstrong1999toward}.
@@
 This approach also introduces new complexities involved in defining the set of components and how they interoperate in the framework, shifting some complexity up the abstraction hierarchy.
-The core features of AMSI used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.
+The core features of \verb|AMSI| used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.
 
@@
 
-AMSI supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations of these interfaces built on top of other software supplied by SCOREC \cite{core} for finite element analysis operations, Simmetrix mesh database software \cite{simmetrix}, and the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}.
-Usage of the supplied numerical components is not required to make use of the multi-scale capabilities of the AMSI libraries, but provide a jumping-off point for application developers needing to implement a new single-scale code to incorporate into a multi-scale simulation.
+\verb|AMSI| supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations of these interfaces built on top of other software supplied by SCOREC \cite{core} for finite element analysis operations, Simmetrix mesh database software \cite{simmetrix}, and the PETSc linear algebraic solver \cite{petsc-web-page,petsc-user-ref,petsc-efficient}.
+Usage of the supplied numerical components is not required to make use of the multi-scale capabilities of the \verb|AMSI| libraries, but provide a jumping-off point for application developers needing to implement a new single-scale code to incorporate into a multi-scale simulation.
 
@@
 \subsection{Simulation Construction with AMSI}
-In order to construct a multi-scale simulation using the AMSI libraries the following must be specified: the Scales in the problem, the ScaleCouplings between individual scales, the DataDistribution of coupling data produced by each Scale in a coupling, and the CouplingPattern used to communicate coupling data during coupling operations.
+In order to construct a multi-scale simulation using the \verb|AMSI| libraries the following must be specified: the Scales in the problem, the ScaleCouplings between individual scales, the DataDistribution of coupling data produced by each Scale in a coupling, and the CouplingPattern used to communicate coupling data during coupling operations.
 
@@
 In the current implementation the RankSet of all Scales in a simulation must be non-overlapping.
-A Scale is required for the specification of all other data structures in the AMSI libraries.
+A Scale is required for the specification of all other data structures in the \verb|AMSI| libraries.
 
@@
 Symmetrical coupling relations therefore require two ScaleCouplings, one denoting the production of data by A for consumption by B and one denoting the reverse.
-Scales not related through a ScaleCoupling cannot create CouplingPatterns which are required for multi-scale communication and therefore cannot pass scale-coupling data through the AMSI libraries.
+Scales not related through a ScaleCoupling cannot create CouplingPatterns which are required for multi-scale communication and therefore cannot pass scale-coupling data through the \verb|AMSI| libraries.
 
@@
 
-ScaleCoupling transformation functions are not currently supported by the AMSI libraries and must be implemented by the application programmer for the each specific multi-scale interaction.
+ScaleCoupling transformation functions are not currently supported by the \verb|AMSI| libraries and must be implemented by the application programmer for the each specific multi-scale interaction.
 
 Constructing a new multi-scale simulation from a set of legacy single-scale simulation codes involves several steps.
-Each Scale in the simulation must have a 'main-like' function with the signature
+Each Scale in the simulation provides a 'main-like' function with the signature
 \begin{verbatim}
@@
 Typically this function can simply be adapted from a main function already used by the single-scale simulation code, but initializing on the provided MPI\_Comm rather than MPI\_Comm\_World.
-Other MPI calls in the code operating on MPI\_COMM\_WORLD should be replaced with AMSI\_COMM\_LOCAL which is the current MPI\_Comm currently associated with the AMSI Scale.
-Multi-scale interaction points are determined and as wells as calls to create the appropriate DataDistributions and CouplingPatterns, which may be created during initialization for simplicity and modified later during execution.
+Other MPI calls in the code operating on MPI\_COMM\_WORLD should be replaced with AMSI\_COMM\_LOCAL which is the MPI\_Comm currently associated with the \verb|AMSI| Scale.
+Multi-scale interaction points are determined and as well as calls to create the appropriate DataDistributions and CouplingPatterns, which may be created during initialization for simplicity and modified later during execution.
 Calls to appropriately set DataDistribution values and modify CouplingPatterns should be added after adaptive procedures are conducted at the desired scale.
-AMSI does not treat any Scale as controlling any other Scale, thus care must be taken to avoid two Scales conducting adaptive procedures on the same CouplingPattern at the same time, though both Scales may conduct such procedures as long as all operations from adaptive procedures at one Scale are completed prior to starting adaptive procedures at the second Scale.
+\verb|AMSI| does not treat any Scale as controlling any other Scale, thus care must be taken to avoid two Scales conducting adaptive procedures on the same CouplingPattern at the same time, though both Scales may conduct such procedures as long as all operations from adaptive procedures at one Scale are completed prior to starting adaptive procedures at the second Scale.
 This allows the CouplingPattern to reach a globally valid state prior to undergoing additional modification.
@@
 
-As noted, AMSI has a flexible definition of what constitutes a scale-task.
+As noted, \verb|AMSI| has a flexible definition of what constitutes a scale-task.
 This allows application developers flexibility in their implementation, but it also prevents certain optimizations -- such as task scheduling -- that a stronger definition of a scale-task can provide.
-For instance, the UINTAH framework \cite{davison200uintah} has a more well-defined task definition which allows the runtime to make scheduling optimizations to reduce process idle time and increase overall parallel performance \cite{meng2010dynamic}.
+For instance, the UINTAH framework \cite{davison2000uintah} has a more well-defined task definition which allows the runtime to make scheduling optimizations to reduce process idle time and increase overall parallel performance \cite{meng2010dynamic}.
 
@@
 
-Each Scale in AMSI represents some number of individual scale-tasks.
-Scale-tasks are not an AMSI construct, merely an abstraction to allow grouping of multiple copies of the same single-scale simulation into a single Scale.
+Each Scale in \verb|AMSI| represents some number of individual scale-tasks.
+Scale-tasks are not an \verb|AMSI| construct, merely an abstraction to allow grouping of multiple copies of the same single-scale simulation into a single Scale.
 Minimally a Scale possess a single scale-task.
@@
 This procedure is called Assembly, where each rank contributes the local (and hence valid) portion of the control data to arrive at a valid Scale-global value.
-In practice this reduces to an MPI\_Allgather operation for the DataDistribution, but for other types of AMSI control data (most specifically the CouplingPattern) the implementation of the Assembly operation requires additional operations.
+In practice this reduces to an MPI\_Allgather operation for the DataDistribution, but for other types of \verb|AMSI| control data (most specifically the CouplingPattern) the implementation of the Assembly operation requires additional operations.
 
@@
 
-All communication and planning operations in the AMSI libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer.
+All communication and planning operations in the \verb|AMSI| libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer.
 Usage only of implicit ordering was decided upon in order to reduce the memory impact as the simulation scales, as any meta-data which scales with the simulation will necessarily pose a memory issue once scaling hits some threshold.
@@
 
-Since scale-tasks are an abstraction the AMSI systems cannot provide direct operations on scale-tasks, but provide mechanisms for the application to inform AMSI of such operations when they occur.
+Since scale-tasks are an abstraction the \verb|AMSI| systems cannot provide direct operations on scale-tasks, but provide mechanisms for the application to inform \verb|AMSI| of such operations when they occur.
 
@@
 Migrating a scale task in whole constitutes using load balancing operations on the process set associated with the entire scale to move all pertinent data between subset of the managed process set.
-AMSI provides operations to enact load-balancing operations from plans such as those generated by Zoltan algorithms.
+\verb|AMSI| provides operations to enact load-balancing operations from plans such as those generated by Zoltan algorithms.
 In the current implementation these operations are collective on the process set associated with the managed scale.
 
-While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the AMSI implementation.
+While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the \verb|AMSI| implementation.
 Any scale is capable of informing the process set associated with a coupled scale of the addition or deletion of individual scale-tasks.
@@
 
-Scale-linking quantities are dynamically registered with the AMSI system using a data distribution defined on a scale by the scale-tasks operating at that scale.
+Scale-linking quantities are dynamically registered with the \verb|AMSI| system using a data distribution defined on a scale by the scale-tasks operating at that scale.
 Production of a communication pattern is required prior to executing a scale-coupling communication, and the pattern must be reconciled (synchronized across the coupling relationship) prior to being used to guide a coupling communication operation (and when the communication pattern is altered by the scale providing the coupling data).
-Several simple communication pattern production algorithms are provided by AMSI, but these are relatively trivial due to the nature of scale-coupling communications and the general nature of the AMSI libraries.
+Several simple communication pattern production algorithms are provided by \verb|AMSI|, but these are relatively trivial due to the nature of scale-coupling communications and the general nature of the \verb|AMSI| libraries.
 Hooks are provided for users to implement algorithms for the construction of communication patterns tailored to the specifics of their problem.
@@
 The current implementation includes blocking versions of the scale-coupling communication operations.
-Asynchronous versions have been designed but are not currently implemented due to lack of immediate need in the primary AMSI use case.
+Asynchronous versions have been designed but are not currently implemented due to lack of immediate need in the primary \verb|AMSI| use case.
 
@@
 After a micro-scale simulation converges to a solution, the Jacobian and force vector are then evaluated again at the solution position, the force exerted along the boundary of the RVE is summed for each of the principal and shear components of a stress tensor, which is then dimensionalized and sent back to the engineering-scale for use in the elemental tangent stiffness matrix formulation.
-More detail into the derivation of this multi-scale system can be found in \cite{stylianopoulos2008thesis} \cite{agoram2001coupled} \cite{stylianopoulos2007multiscale} .
-Discussion of the dimensionalization of the dimensionless force terms produced by each fiber network can be found in \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.
+More detail into the derivation of this multi-scale system can be found in \cite{stylianopoulos2008thesis,agoram2001coupled,stylianopoulos2007multiscale} .
+Discussion of the dimensionalization of the dimensionless force terms produced by each fiber network can be found in \cite{stylianopoulos2007volume,chandran2007deterministic}.
 
@@
 Separate timings were taken for various sections of the simulation code, isolating the micro-scale computation, macro-scale linear system assembly and calculation of elemental linear systems, and macro-scale linear system solution.
-In addition, the time for spent in the AMSI subroutines for communication was also measured.
+In addition, the time for spent in the \verb|AMSI| subroutines for communication was also measured.
 Thus the behavior of each section of the code is independently evaluated in our scaling study, as well as the overall execution time.
@@
 As indicated in the weak scaling results this is related primarily to the macro-scale linear solve communication time increasing.
-At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU\_dist solver \cite{li05superlu} \cite{lishao10superluilu} \cite{lidemmel03superludist} ).
+At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU\_dist solver \cite{li05superlu,lishao10superluilu,lidemmel03superludist} ).
 Using instead an iterative solution method to reach a desired level of accuracy may result in stronger scaling at macro-scale, but as yet this approach has been unneeded due to the preponderance of time spent computing micro-scale results in use cases.
@@
 Many problems of interest will possess finite elements numbered in the tens to hundreds of millions of degrees of freedom, such that a suitable partitioning of the mesh to cores on the machine in terms of elements per partition (and directly the macro-scale time-to-solution) may require a substantial fraction of the total cores on the machine for macro-scale to execute in reasonable time.
-AMSI provides services for dynamically adding and removing individual scale-tasks to a specific scale, which can be utilized to limit RVE usage to only those locations in the multi-scale mesh most influential on the overall evolution of the simulation.
+\verb|AMSI| provides services for dynamically adding and removing individual scale-tasks to a specific scale, which can be utilized to limit RVE usage to only those locations in the multi-scale mesh most influential on the overall evolution of the simulation.
 
-AMSI-specific results of some sort...?
+\verb|AMSI|-specific results of some sort...?
 
@@
 The Adaptive Multi-scale Simulation Infrastructure provides a valuable set of tools to facilitate the implementation and execution of multi-scale simulations without imposing restrictions on the application developer.
-Usage of the core operations of the AMSI libraries does not require adherence to a specific set of interfaces nor usage of an abstract modeling language, which complicate inclusion of legacy components into multi-scale simulations.
+Usage of the core operations of the \verb|AMSI| libraries does not require adherence to a specific set of interfaces nor usage of an abstract modeling language, which complicate inclusion of legacy components into multi-scale simulations.
 
@@
 Development of libraries concerned with higher levels of abstraction of the multi-scale problem will also be valuable in easing the development of new multi-scale simulations from the perspective of the application programmer.
-High level abstractions such as the Scale Separation Map in \cite{chopard2011framework} could provide mechanisms to provide initial configuration information to the AMSI system during the initialization phase, as well as assist in the implementation of (or usage of existing library-provided) up-/down-scaling operations.
+High level abstractions such as the Scale Separation Map in \cite{chopard2011framework} could provide mechanisms to provide initial configuration information to the \verb|AMSI| system during the initialization phase, as well as assist in the implementation of (or usage of existing library-provided) up-/down-scaling operations.
 
-An RVE simulation taking into account inter-fibril matrix material has been developed \cite{lake2012mechanics} \cite{zhang2013cross} \cite{zhang2013coupled}.
+An RVE simulation taking into account inter-fibril matrix material has been developed \cite{lake2012mechanics,zhang2013cross,zhang2013coupled}.
 This alternative RVE will be incorporated into the multi-scale model as an alternative to the fiber-only RVE, to be used in areas of the macro-scale domain determined to be most critical to the overall evolution of the global system, or where error accumulation through model inaccuracy is to be minimized.
-As this fiber-matrix RVE represents a more significant source of computational complexity, it is independently parallelized -- in contrast to the fiber-only RVE -- and will represent a compelling use case for the parallel execution space discretization and scale-coupling communication management capabilities of the AMSI libraries.
+As this fiber-matrix RVE represents a more significant source of computational complexity, it is independently parallelized -- in contrast to the fiber-only RVE -- and will represent a compelling use case for the parallel execution space discretization and scale-coupling communication management capabilities of the \verb|AMSI| libraries.
 
@@
 
-\bibliography{citations}
+\bibliography{multiscale_theory,multiscale_frameworks,load_balancing,parallel,software,biotissue}
 

### 2016-01-27 interface/amsiNonlinearAnalysis.cc

@@ -1,0 +1,18 @@
+#include "amsiNonlinearAnalysis.h"
+namespace amsi
+{
+  bool numericalSolve(Iteration * it, Convergence * cn)
+  {
+    bool failure = false;
+    while(!cn->converged())
+    {
+      it->iterate();
+      if(cn->failed())
+      {
+        failure = true;
+        break;
+      }
+    }
+    return failure;
+  }
+}

### 2016-01-27 interface/amsiNonlinearAnalysis.h

@@ -1,0 +1,18 @@
+#ifndef AMSI_NONLINEAR_ANALYSIS_H_
+#define AMSI_NONLINEAR_ANALYSIS_H_
+namespace amsi
+{
+  class Iteration
+  {
+  public:
+    virtual void iterate() = 0;
+  };
+  class Convergence
+  {
+  public:
+    virtual bool converged() = 0;
+    virtual bool failed() {return false;}
+  };
+  bool numericalSolve(Iteration * it, Convergence * cn);
+}
+#endif

### 2016-05-25 doc/articles/amsi_biotissue/paper.tex
working on stuff for volume convergence

@@
-\title{The Adaptive Multi-scale Simulation Infrastructure - Design and Application}
+\title{The Adaptive Multi-scale Simulation Infrastructure}
 \author{W.R. Tobin, V.W.L. Chan, M.S. Shephard}
@@
 \bibliographystyle{siam}
-
 \usepackage{amsfonts}
@@
 \usepackage{listings}
-%\usepackage{mathtools}
 \usepackage{url}
-%\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
-%\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
-
 \lstset{basicstyle=\small,
@@
         captionpos=t}
-
 \begin{document}
 \maketitle
-
 \begin{abstract}
 \end{abstract}
-
 \section{Introduction}\label{introduction}
@@
 
-Another Newton-Raphson iterative process occurs at micro-scale to converge to the set of displacements resulting in force equilibrium for all the fibers within the fiber network constituting the RVE.
+'Another Newton-Raphson iterative process occurs at micro-scale to converge to the set of displacements resulting in force equilibrium for all the fibers within the fiber network constituting the RVE.
 After a micro-scale simulation converges to a solution, the Jacobian and force vector are then evaluated again at the solution position, the force exerted along the boundary of the RVE is summed for each of the principal and shear components of a stress tensor, which is then dimensionalized and sent back to the engineering-scale for use in the elemental tangent stiffness matrix formulation.

### 2016-05-25 interface/SimFEA.h

@@ -27,6 +27,7 @@ namespace amsi
            pACase pd,
            MPI_Comm cm = AMSI_COMM_SCALE);
     pGModel getGeometricDomain() { return model; }
+    pMesh getPart() { return part; }
     virtual void Adapt();
     virtual void addFieldToMap(pField);
   };

### 2016-05-25 interface/amsiDummyLAS.h

@@ -79,6 +79,9 @@ namespace amsi
     {
       vec = vc;
     }
+    // stubb
+    virtual void GetAccumVector(double *&)
+    {    }
     virtual void SetVector(const double * vec)
     {
       memcpy(&vc[0],&vec[0],dof*sizeof(double));

### 2016-05-25 interface/amsiLAS.h

@@ -35,12 +35,16 @@ namespace amsi
     virtual void GetVector(double *&) = 0;
     virtual void SetVector(const double *) = 0;
     virtual void GetVectorNorm(double &) = 0;
+    virtual void GetAccumVector(double *&) = 0;
     virtual void GetAccumVectorNorm(double &) = 0;
     virtual void GetDotNorm(double &) = 0;
     virtual void PrintMatrix(std::ostream &) {};
     virtual void PrintVector(std::ostream &) {};
     virtual void PrintSolution(std::ostream &) {};
     virtual double MatrixMax() { return 0.0; }
+    virtual int GlobalDOFs() { return 0; }
+    virtual int LocalDOFs() { return 0; }
+    virtual int LocalOffset() { return 0; }
     virtual ~LAS() {};
   };
   /*

### 2016-05-25 interface/amsiPETScLAS.cc

@@ -25,6 +25,7 @@ namespace amsi
       int n = num_local_unknowns;
       x_arr = new double[n];
       b_arr = new double[n];
+      b_i_arr = new double[n];
       VecCreateMPI(PETSC_COMM_WORLD,n,N,&b_i);
       VecSetOption(b_i, VEC_IGNORE_NEGATIVE_INDICES, PETSC_TRUE);
       MatCreateAIJ(PETSC_COMM_WORLD,n,n,N,N,300,PETSC_NULL,300,PETSC_NULL,&A);
@@ -198,19 +199,23 @@ namespace amsi
    *
    *@todo Construct a test case.
    */
-  void PetscLAS::GetVector(double * & vec)
+  void PetscLAS::GetVector(double * & vc)
   {
-    PetscScalar * temp;
-    VecScatter ctx;
-    VecScatterCreateToAll(b_i,&ctx,&w);
-    VecScatterBegin(ctx,b_i,w,INSERT_VALUES,SCATTER_FORWARD);
-    VecScatterEnd(ctx,b_i,w,INSERT_VALUES,SCATTER_FORWARD);
-    VecGetArray(w,&temp);
-    memcpy(b_arr,(double*)temp,globalNumEqs*sizeof(PetscScalar));
-    VecRestoreArray(w,&temp);
-    VecScatterDestroy(&ctx);
-    VecZeroEntries(w);
-    vec = b_arr;
+    PetscScalar * Bi;
+    VecGetArray(b_i,&Bi);
+    int n = vec_high-vec_low;
+    memcpy(b_i_arr,(double*)Bi,n*sizeof(PetscScalar));
+    VecRestoreArray(b_i,&Bi);
+    vc = &b_i_arr[0];
+  }
+  void PetscLAS::GetAccumVector(double * & vc)
+  {
+    PetscScalar * B;
+    VecGetArray(b,&B);
+    int n = vec_high-vec_low;
+    memcpy(b_arr,(double*)B,n*sizeof(PetscScalar));
+    VecRestoreArray(b,&B);
+    vc = &b_arr[0];
   }
   /**
    *@brief Overwrite the values in the associated vector with those from the provided array.
@@ -286,6 +291,7 @@ namespace amsi
     VecDestroy(&w);
     delete[] x_arr;
     delete[] b_arr;
+    delete[] b_i_arr;
   }
   /**
    *@brief Constructor.
[…]

### 2016-05-25 interface/amsiPETScLAS.h

@@ -25,6 +25,7 @@ namespace amsi
     void GetVector(double *&);
     void SetVector(const double *);
     void GetVectorNorm(double &);
+    void GetAccumVector(double *&);
     void GetAccumVectorNorm(double &);
     void GetDotNorm(double &);
     void GetSolution(double *&);
@@ -34,6 +35,9 @@ namespace amsi
     virtual void PrintVector(std::ostream &);
     virtual void PrintSolution(std::ostream &);
     double MatrixMax();
+    virtual int GlobalDOFs() { return globalNumEqs; }
+    virtual int LocalDOFs() { return vec_high - vec_low; }
+    virtual int LocalOffset() { return vec_low; }
     ~PetscLAS();
   private:
     void freeMem();
@@ -45,7 +49,7 @@ namespace amsi
     Vec b_i;     // current vector
     Vec b;
     Vec w;     // work vector
-    double * x_arr, * b_arr;
+    double * x_arr, * b_arr, * b_i_arr;
     int globalNumEqs;
     int vec_low;
     int vec_high;

### 2016-05-25 interface/apfFunctions.cc

@@ -1,4 +1,6 @@
 #include "apfFunctions.h"
+#include "ElementalSystem.h"
+#include <apfDynamicMatrix.h>
 #include <apfField.h>
 #include <apfGeometry.h>
 #include <apfMesh.h>
@@ -126,5 +128,71 @@ namespace amsi
     mat[0][1] = mat[1][0] = vec[3];
     mat[1][2] = mat[2][1] = vec[4];
     mat[0][2] = mat[2][0] = vec[5];
+  }
+  class MeasureDisplaced : public amsi::ElementalSystem
+  {
+  public:
+    MeasureDisplaced(apf::Field * field, int o)
+      : ElementalSystem(field,o)
+      , dim(0)
+      , vol(0)
+    {}
+    void inElement(apf::MeshElement * me)
+    {
+      ElementalSystem::inElement(me);
+      fs = apf::getShape(f);
+      es = fs->getEntityShape(apf::getMesh(f)->getType(apf::getMeshEntity(me)));
+      dim = apf::getDimension(me);
+    }
+    void atPoint(apf::Vector3 const &p, double w, double dV)
+    {
+      int & nen = nenodes; // = 4 (tets)
+      // 1. Get coordinates on underlying mesh
+      apf::Mesh * mesh = apf::getMesh(f);
+      apf::Field * apf_coord_field = mesh->getCoordinateField();
+      apf::Element * mesh_coord_elem = apf::createElement(apf_coord_field,me);
+      apf::NewArray<apf::Vector3> mesh_xyz;
+      apf::getVectorNodes(mesh_coord_elem,mesh_xyz);
+      // 2. Get coordinates from apf_primary_field (passed in), which contains the accumulated displacement
+      apf::NewArray<apf::Vector3> primary_field_xyz;
+      apf::getVectorNodes(e,primary_field_xyz);
+      // 3. Calculate current coordinates
+      apf::DynamicMatrix xyz(nen,dim); xyz.zero();
+      for (int ii = 0; ii < nen; ii++)
+        for (int jj = 0; jj < dim; jj++)
+          xyz(ii,jj) = mesh_xyz[ii][jj] + primary_field_xyz[ii][jj];
+      // For Updated Lagrangian, the Jacobian of the updated coordinates are used
+      // Note: that entires of Jacobian is hard coded for Linear tetrahedra elements.
[…]

### 2016-05-25 interface/apfFunctions.h

@@ -283,6 +283,7 @@ namespace amsi
   void SymmMatrixToVoigtVector(const apf::Matrix3x3 & mat, apf::Vector<6> & vec);
   void VoigtVectorToSymmMatrix(const apf::Vector<6> & vec, apf::Matrix3x3 & mat);
   void VoigtVectorToSymmMatrix(const apf::DynamicVector & vec, apf::Matrix3x3 & mat);
+  double measureDisplaced(apf::MeshEntity * ment, apf::Field * u);
 }
 #include "apfFunctions_impl.h"
 #endif

### 2016-05-25 interface/apfsimWrapper.cc

@@ -1,6 +1,31 @@
 #include "apfsimWrapper.h"
+#include "apfFunctions.h"
+#include "simAnalysis.h"
 namespace amsi
 {
+  double measureEntity(pGEntity ent, pMesh msh, apf::Mesh * apf_msh)
+  {
+    // todo : determine dimension of entity
+    double vol = 0.0;
+    std::list<pEntity> ents;
+    getClassifiedEnts(msh,ent,3,std::back_inserter(ents));
+    for(std::list<pEntity>::iterator ent = ents.begin(); ent != ents.end(); ++ent)
+    {
+      apf::MeshElement * mnt = apf::createMeshElement(apf_msh,apf::castEntity(*ent));
+      vol += apf::measure(mnt);
+      apf::destroyMeshElement(mnt);
+    }
+    return vol;
+  }
+  double measureDisplacedEntity(pGEntity ent, pMesh msh, apf::Field * u)
+  {
+    double vol = 0.0;
+    std::list<pEntity> ents;
+    getClassifiedEnts(msh,ent,3,std::back_inserter(ents));
+    for(std::list<pEntity>::iterator ent = ents.begin(); ent != ents.end(); ++ent)
+      vol += measureDisplaced(apf::castEntity(*ent),u);
+    return vol;
+  }
   void applyUniqueRegionTags(pGModel mdl, pMesh msh, apf::Mesh * apfmsh)
   {
     pGEntity rgn = NULL;

### 2016-05-25 interface/apfsimWrapper.h

@@ -8,10 +8,10 @@
 #include <list>
 namespace amsi
 {
+  double measureEntity(pGEntity ent, pMesh msh, apf::Mesh * apf_mesh);
+  double measureDisplacedEntity(pGEntity ent, pMesh msh, apf::Field * u);
   class Sim2ApfDirichletBCApplier : public SimDirichletBCApplier
-  {
-    
-  };
+  {  };
   /*
   template <int dim>
     class SimIterator

### 2016-05-25 interface/apfsimWrapper.cc
Added comm_sum to volume calculations in apfsimWrapper.cc to fix issue of hanging. Issue arose from processors calculating different values of volume resulting in volume convergence in one processor and not the other.

@@ -15,6 +15,7 @@ namespace amsi
       vol += apf::measure(mnt);
       apf::destroyMeshElement(mnt);
     }
+    vol = amsi::comm_sum(vol);
     return vol;
   }
   double measureDisplacedEntity(pGEntity ent, pMesh msh, apf::Field * u)
@@ -24,6 +25,7 @@ namespace amsi
     getClassifiedEnts(msh,ent,3,std::back_inserter(ents));
     for(std::list<pEntity>::iterator ent = ents.begin(); ent != ents.end(); ++ent)
       vol += measureDisplaced(apf::castEntity(*ent),u);
+    vol = amsi::comm_sum(vol);
     return vol;
   }
   void applyUniqueRegionTags(pGModel mdl, pMesh msh, apf::Mesh * apfmsh)

### 2016-09-06 interface/apfsimWrapper.cc
work on standalone model region volume calculation test case

@@ -34,12 +34,12 @@ namespace amsi
      * pGFace ent:     face entity of geometric model.
      * pMesh msh:      partitioned mesh.
      * apf::Field * u: displacement field. */
-     double vol = 0.0;
+    double vol = 0.0;
     /** Extract face entities of mesh that lie on face entities of geometric model */
-     std::list<pEntity> ents;
-     getClassifiedEnts(msh,pGF,2,std::back_inserter(ents));
+    std::list<pEntity> ents;
+    getClassifiedEnts(msh,pGF,2,std::back_inserter(ents));
     /** for debugging purposes */
-     AMSI_DEBUG (
+    AMSI_DEBUG (
       std::vector<int> tags;
       int ent_size = 0;
       for (int ii=ent_size; ii<ents.size(); ii++)
@@ -47,11 +47,11 @@ namespace amsi
         tags.push_back(GEN_tag(pGF));
         ent_size = ents.size();
       }
-      )
-     int surf_elem = 0;
-     apf::Mesh * apfmsh = apf::getMesh(u);
-     for(std::list<pEntity>::iterator ent = ents.begin(); ent != ents.end(); ++ent)
-     {
+    )
+    int surf_elem = 0;
+    apf::Mesh * apfmsh = apf::getMesh(u);
+    for(std::list<pEntity>::iterator ent = ents.begin(); ent != ents.end(); ++ent)
+    {
       if(apfmsh->isOwned(apf::castEntity(*ent)))
       {
         pRegion mshRgn0 = F_region((pFace)(*ent),0);
@@ -69,14 +69,12 @@ namespace amsi
           normal_dir = 1;
         else
           normal_dir = -1;
-/*
         AMSI_DEBUG (
           std::cout<<"surf_elem:"<<surf_elem<<" belongs to face: "<<tags[surf_elem]
           <<" Rgn "<<GRgn0_tag<<" inside,"
           <<" Rgn "<<GRgn1_tag<<" outside,"
           <<" normal direction "<<normal_dir<<std::endl;
-          )
-*/
+        )
         vol += measureDisplacedFromSurf(apf::castEntity(*ent),u,normal_dir);
         surf_elem++;
       }

### 2016-09-06 interface/simWrapper.h

@@ -32,7 +32,7 @@ namespace amsi
           PList_append(lst,NULL);
           sz++;
         }
-	tmp.v[0] = sz;
+        tmp.v[0] = sz;
         ent = PList_item(lst,n);
       }
     }

### 2016-09-06 scripts/autobuild.sh

@@ -1,4 +1,4 @@
-#!/bin/bash
+#!/bin/bash -x
 
 if [ -z $1 ]; then
   BUILD_TYPE=Debug

### 2016-09-06 test/interface/analysis/volume.cc

@@ -1,0 +1,26 @@
+#include "apfsimWrapper.h"
+#include "simWrapper.h"
+#include <cstdlib>
+int main(int agrc, char * argv[])
+{
+  int result = 0;
+  amsi::use_simmetrix = true;
+  amsi::interfaceInit(argc,argv);
+  pGModel mdl = GM_load(argv[1],0,NULL);
+  pParMesh msh = PM_load(argv[2],sthreadnone,mdl,NULL);
+  int tg = atoi(argv[3]);
+  pMesh prt = PM_mesh(msh,0);
+  apf::Field * u =  apf::createLagrangeField(msh,"u",apf::VECTOR,1);
+  pGRegion rgn = (pGRegion)GM_entityByTag(mdl,3,tg);
+  std::vector<pGEntity> fcs;
+  amsi::cutPaste<pGEntity>(GEN_faces((pGEntity)rgn),fcs);
+  double vol = 0.0;
+  for(std::vector<pGEntity>::iterator fc = fcs.begin(); fc != fcs.end(); ++fc)
+    vol += measureVol_pGFace((pGFace)*fc,tg,prt,u);
+  int rnk = -1;
+  MPI_Comm_rank(MPI_COMM_WORLD,&rnk);
+  if(rnk == 0)
+    std::cout << vol << std::endl;
+  amsi::interfaceFree();
+  return result;
+}

### 2016-09-06 util/amsiFail.h

@@ -1,8 +1,6 @@
 #ifndef AMSI_FAIL_H_
 #define AMSI_FAIL_H_
-
 #include <limits>
-
 namespace amsi
 {
 # if __cpp_alias_templates >= 200704 // check to see if template aliasing is available in gcc
@@ -11,5 +9,4 @@ namespace amsi
   template <typename T> T fail() {return std::numeric_limits<T>::min();}
 # endif
 }
-
 #endif

### 2016-09-06 util/amsiUUID.h

@@ -19,7 +19,7 @@ namespace amsi
   }
   template <class T>
     inline uuid generateRangeUUID(const T & begin,
-				  const T & end)
+                                  const T & end)
   {
     uuid result = 0;
     for(T it = begin; it != end; it++)

### 2016-10-20 doc/articles/amsi_biotissue/draft2.tex
missing file

--- 2016-01-27 doc/articles/amsi_biotissue/paper.tex
@@
         captionpos=t}
-
 \begin{document}
 \maketitle
-
 \begin{abstract}
 \end{abstract}
+\section{Introduction}\label{introduction}
 
-\section{Introduction}\label{introduction}
-[Need a couple sentences to talk about how multi-scale is important]
-
-Implementation of multi-scale numerical simulations typically rely on either an ad-hoc set of algorithms designed for a particular problem, or on large frameworks and runtimes.
-Ad-hoc approaches, by their nature, are targeted to a particular multi-scale problem for a specific set of single-scale numerical codes, and thus offer no wider utility beyond their immediate usage as a simulation code.
-Framework and runtime approaches are more flexible and admit a broader range of reuse, but typically necessitate that single-scale operations be adapted to work in the ecosystem they create.
-
-Development of multi-scale simulations and methods to support the implementation of such simulations has been ongoing for several decades.
-Some approaches begin with the specification of multi-scale contributions during the specification of mathematical model.
-However, multi-scale influences can be incorporated in a strictly numerical or computational sense, by combining implementations of established physical models operating at different scales referencing and effecting shared physical fields and values \cite{shenoy97} \cite{weinan2003heterogenous}.
-
-% classes of multi-scale problems
-In order for a numerically multi-scale approach to be effective the domain associated with the scale of interest (the engineering scale) must be of sufficient size that conducting the entire simulation with terms and granularity orders of magnitude removed would make the simulation computationally intractable.
-Thus some computationally reasonable subset of the engineering scale is assigned significantly finer scale(s) to simulate.
-The choice of location of supporting fine-scale simulations is largely dependent on the specific mathematical models and numerical methods used to implement the engineering scale simulation.
-Two general classes of fine-scale simulation locality have been noted \cite{weinan2011principles}, 'A' type problems where local granularity requirements such as boundary layers or crack propagation necessitate use of localized fine-scale simulation, and 'B' type problems where fine-scale simulation is used to supplement or replace a continuum model at the engineering scale.
-For either case the subset of the engineering-scale domain associated with finer scales are areas of the domain where the fine-scale contributions have the greatest impact on the evolution of the engineering scale simulation, in order to leverage the most utility from the multi-scale approach.
-
-% types of numerical paradigms related to multi-scale, maybe mention that a problem using the concurrent multi-scale paradigm is discussed later in the paper?
-Numerical multi-scale simulations fall broadly into two parallel paradigms, distinct from the choice of subdomain assigned to fine scale simulation.
-In the sequential multi-scale paradigm, fine-scale computation is used to precompute quantities of interest for establishing an engineering-scale physical model, possibly using a lookup table or some other method to incorporate the multi-scale influences into the engineering scale \cite{weinan2011principles} \cite{garcia2008sequential}.
-In the concurrent multi-scale paradigm, all scales are computed simultaneously and the required inter-scale coupling values are provided just-in-time \cite{weinan2011principles} \cite{zeng2010concurrent}.
-
-% introduce scale-coupling topic to be discussed further in AMSI and methods sections
-Numerically multi-scale simulations associate the computational implementations of single-scale physical models and cause them to interact by passing key physical terms between scales.
-This scale-coupling operation and may require the application of transformer functions to account for key physical differences between disparate scales.
-The primary purpose of such transformer functions is to convert physical coupling values produced by simulations operations at one scale to be meaningfully applied to simulation operations at the coupled, consuming scale.
-These transformer functions are typified by conversions between the unit metrics used by coupled scales.
-For example a micro-scale simulation will typically be implemented with units of $\mu$m while a macro-scale simulation might be implemented using units of cm.
-Scale-coupling transformer functions may also be used for any additional coupling-term operations required to account for differences between coupled scales \cite{weinan2003heterogenous}, such as dimensionalizing terms calculated in a dimensionless domain \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.
-
-Implementation of multi-scale simulations has fallen widely into two categories: purely ad-hoc implementations, and implementations supported by a structured framework or runtime.
-In the case of an ad-hoc implementation all support for multi-scale interaction is hard-coded into the simulation code itself.
-Frameworks and runtimes provide a robust set of scale-coupling operations, scale management,  and overall simulation management.
-The application programmer must tailor the numerical operations occurring at a discrete scale to operate using a set of interfaces or high-level modelling language required by the framework/runtime.
-Ad-hoc implementations allow for overall optimizations at the cost of development time (and potentially complexity).
-Framework/runtime implementations often have a component-based approach that breaks the implementation of a simulation (single- or multi-scale) into discrete interoperable components.
-This approach requires that operations are implemented to work specifically within the set of interfaces understood by the framework/runtime.
-
-% introduce legacy component concept which should be discussed in AMSI and methods sections
-Decades of numerical simulation development have resulted in a large and diverse array of widely used codes for single-scale simulations.
-This set of legacy codes represents a massive investment of development efforts, and thus leveraging the capabilities of such legacy codes for use in multi-scale simulations is desirable.
-Incorporating such codes into a functional numerical implementation of a multi-scale model has been a largely ad-hoc, per-code affair.
-Providing tools to ease the incorporation of legacy components into multi-scale simulations using general procedures while requiring minimal intervention in the legacy codebase results in less fracturing of the legacy codebases when incorporated into varied multi-scale simulations, and avoids the option of complete reimplementation of legacy code functionality in specialized multi-scale ad-hoc codes, or structured frameworks or runtimes.
-
-\section{Adaptive Multi-scale Simulation Infrastructure}\label{AMSI}
-The Adaptive Multi-scale Simulation Infrastructure (AMSI) is a set of libraries and tools developed by the Scientific Computation Research Center (SCOREC) at Rensselaer Polytechnic Institute.
-AMSI is designed to support the implementation and execution of dynamic concurrent multi-scale simulations on massively-parallel HPC machines.
-AMSI provides support for several key operations needed for concurrent multi-scale simulations: modelling and managing scales, scale-coupling parallel communication, adaptively adding/removing scale-tasks, and load-balancing coupled scales.
-A key goal in the design and implementation of these operations is a focus on general applicability and avoiding adherence to specific interfaces.
-
-The interfaces provided by the AMSI libraries are intended to facilitate the coupling of functional single-scale simulation codes.
-This approach is used in order to mitigate development overhead associated with casting a simulation into a set of framework-specific components, or developing an entire ad-hoc set of multi-scale operations each time a new multi-scale code is being developed.
-Several key multi-scale operations are planned and enacted by the AMSI libraries, allowing the application programmer (a programmer using the AMSI libraries to implement a new multi-scale code or interaction) to determine the most effective use of the tools provided.
-All interfaces are designed to be physics-agnostic, operating only on computational quantities with no meta-model of the relation between a coupling term and the mathematical derivation of the multi-scale interaction.
-
-AMSI operates by maintaining a set of minimal simulation metadata in order to model various quantities of interest during simulation execution.
-AMSI provides mechanisms to assist in the implementation of run-time decision making and control decisions related to multi-scale execution, with a focus on allowing such adaptive procedures to be planned and implemented in a decentralized manner.
-This design approach allows application developers to avoid centralized control decision bottlenecks.
-Control decisions and associated parallel communication (as with most adaptive procedures) are best implemented at locations in code already requiring parallel barriers for synchronization.
-
-% need to expand this substantially
-Frameworks and runtime systems have been developed which are intended for the development of specific multi-scale problems \cite{davison2000uintah} \cite{parker2002component} \cite{chopard2011framework}.
-These frameworks have in some cases been expanded to facilitate the implementation of additional multi-scale problems \cite{berzins2010uintah} \cite{parker20026component}.
-Efforts have also been made to provide standard commponent interfaces to encourage more widespread component adoption in such systems \cite{armstrong1999toward}.
-
-Such systems require that the application developer cast the problem of interest into a set of predefined components (interfaces, tasks, etc.) or specify the problem through a high-level modelling language in order to take full advantage the services provided by the framework.
-Leveraging a component-based design is valuable from a software engineering perspective, as it allows composition of disparate components into a cohesive application.
-This approach also introduces new complexities involved in defining the set of components and how they interoperate in the framework, shifting some complexity up the abstraction hierarchy.
-The core features of AMSI used for multi-scale simulation management and communication are provided with no requirements of adherence to a specific component definition.
-
-%Typically simulation flow control is determined solely by a runtime system.
-%A framework may also limit specific features of the numerical implementation to those provided by the framework.
-%This includes domain discretization models and associated adaptive processes -- structured and unstructured meshing and associated refinement algorithms, particle-in-cell methods, etc.
--- convergence algorithms for both the minimization of the relevant residuals (Newton-Raphson, etc.) as well as iterative solution of the resulting linear system (KSP, GMRES, ILU, etc.).
-
-AMSI supplies several numerical components consisting of an adaptive finite element system, linear algebraic solver, and nonlinear convergence operators available as both simple C++ class interface definitions, as well as implementations of these interfaces built on top of other software supplied by SCOREC \cite{core} for finite element analysis operations, Simmetrix mesh database software \cite{simmetrix}, and the PETSc linear algebraic solver \cite{petsc-web-page} \cite{petsc-user-ref} \cite{petsc-efficient}.
-Usage of the supplied numerical components is not required to make use of the multi-scale capabilities of the AMSI libraries, but provide a jumping-off point for application developers needing to implement a new single-scale code to incorporate into a multi-scale simulation.
-
-Implementing required numerical operations inside an existing framework is of course possible, though it may prove an additional -- undesirable overhead for an application programmer interested only in utilizing the multi-scale framework to implement a simulation, rather than implementing the various numerical operations often optimized in widely available packages and libraries.
+\section{Adaptive Multi-scale Simulation Infrastructure}\label{amsi}
 
@@
 
-\subsection{Simulation Construction with AMSI}
-In order to construct a multi-scale simulation using the AMSI libraries the following must be specified: the Scales in the problem, the ScaleCouplings between individual scales, the DataDistribution of coupling data produced by each Scale in a coupling, and the CouplingPattern used to communicate coupling data during coupling operations.
-
-A Scale is a data structure modelling the parallel state of a discrete simulation scale and all operations (or tasks) occurring at that scale.
-All Scales have a set of MPI\_Ranks associated with them through an intermediary structure called the RankSet.
-In the current implementation the RankSet of all Scales in a simulation must be non-overlapping.
-A Scale is required for the specification of all other data structures in the AMSI libraries.
-
-A ScaleCoupling is an object denoting that scale-coupling communication will take place between two scales at some point during the simulation.
-All interacting Scales in a simulation must declare a ScaleCoupling between them.
-For generality the ScaleCoupling relationship is non-commutative, thus Scale A is related to Scale B if and only if A produces coupling data which will be sent to B.
-Symmetrical coupling relations therefore require two ScaleCouplings, one denoting the production of data by A for consumption by B and one denoting the reverse.
-Scales not related through a ScaleCoupling cannot create CouplingPatterns which are required for multi-scale communication and therefore cannot pass scale-coupling data through the AMSI libraries.
-
-A DataDistribution is a structure denoting the parallel distribution of terms produced by a scale for use in a ScaleCoupling relationship for a specific CouplingPattern.
-The DataDistribution is minimally described as a parallel array of integers on the RankSet associated with some Scale A, thus the length of the array is $|A|$ (the cardinality of the RankSet currently assigned to Scale A).
-The DataDistribution may be dynamically modified during the simulation to reflect single-scale adaptive procedures or specifically multi-scale adaptive features (such as selectively adding/removing specific fine-scale simulations).
-Each Scale may have many DataDistributions, and a DataDistribution is not tied to a particular ScaleCoupling, in the event that the same distribution characterizes the scale-coupling data produced by a Scale needed by multiple coupled scales.
-
-A CouplingPattern describes the parallel communication required to send scale-coupling data produced by the the sending Scale in a ScaleCoupling and described by a DataDistribution.
-A CouplingPattern based on the ScaleCoupling of scale $A$ coupled to scale $B$ is described most simply as a matrix $C \in \mathbb{N}^{|A|\times|B|}$ (where $|X|$ is the cardinality of the RankSet currently assigned to scale X).
-The quantity $C_{ij}$ represents the quantity of scale-coupling terms (of arbitrary type) to be sent from process $i$ in the process set assigned to $A$ to process $j$ in the process set assigned to $B$.
-As the DataDistribution upon which the CouplingPattern is based is allowed to change due to adaptive procedures, all CouplingPatterns related to a DataDistribution are updated when adding/removing terms from a DataDistribution.
-Finalization of the creation of a CouplingPattern causes the creation of a DataDistribution on the receiving Scale, as modifications to a DataDistribution may happen on either side of the ScaleCoupling, but the initial CouplingPattern must be created by the producing Scale.
-Determination of how a CouplingPattern should change based on changes to the underlying DataDistribution may differ for potentially each multi-scale use-case, so hooks are provided so the application program may correctly update CouplingPatterns for their specific use case.
-As both Scales in the ScaleCoupling must know about the modification in order to appropriately handle the modified scale-coupling data being communicated, all modifications to the CouplingPattern and DataDistribution are communicated from the producing Scale to the consuming Scale.
-Due to this requirement it is best to update DataDistributions completely in a single operation following any adaptive procedures, since these are often parallel barriers already.
-Similar operations are provided for scale-sensitive load-balancing operations taking place on a sending or receiving scale in a ScaleCoupling, as the associated scale must be informed of the changes to the relevant DataDistribution.
-
-ScaleCoupling transformation functions are not currently supported by the AMSI libraries and must be implemented by the application programmer for the each specific multi-scale interaction.
-
-Constructing a new multi-scale simulation from a set of legacy single-scale simulation codes involves several steps.
-Each Scale in the simulation must have a 'main-like' function with the signature
-\begin{verbatim}
-int (*xxx_main)(int *, char *[], MPI_Comm)
-\end{verbatim}
-which initializes the various supporting libraries used by that single-scale code on the provided MPI\_Comm and then begins conducting the simulation at that Scale.
-Typically this function can simply be adapted from a main function already used by the single-scale simulation code, but initializing on the provided MPI\_Comm rather than MPI\_Comm\_World.
-Other MPI calls in the code operating on MPI\_COMM\_WORLD should be replaced with AMSI\_COMM\_LOCAL which is the current MPI\_Comm currently associated with the AMSI Scale.
-Multi-scale interaction points are determined and as wells as calls to create the appropriate DataDistributions and CouplingPatterns, which may be created during initialization for simplicity and modified later during execution.
-Calls to appropriately set DataDistribution values and modify CouplingPatterns should be added after adaptive procedures are conducted at the desired scale.
-AMSI does not treat any Scale as controlling any other Scale, thus care must be taken to avoid two Scales conducting adaptive procedures on the same CouplingPattern at the same time, though both Scales may conduct such procedures as long as all operations from adaptive procedures at one Scale are completed prior to starting adaptive procedures at the second Scale.
-This allows the CouplingPattern to reach a globally valid state prior to undergoing additional modification.
-When a Scale has many scale-tasks operating it is necessary to write new code to manage the instantiation and management of the various scale-tasks, and update the relevant DataDistributions as required.
-
-As noted, AMSI has a flexible definition of what constitutes a scale-task.
-This allows application developers flexibility in their implementation, but it also prevents certain optimizations -- such as task scheduling -- that a stronger definition of a scale-task can provide.
-For instance, the UINTAH framework \cite{davison200uintah} has a more well-defined task definition which allows the runtime to make scheduling optimizations to reduce process idle time and increase overall parallel performance \cite{meng2010dynamic}.
-
-\subsection{Parallel Operations}\label{paralle_ops}
-
-Each Scale in AMSI represents some number of individual scale-tasks.
-Scale-tasks are not an AMSI construct, merely an abstraction to allow grouping of multiple copies of the same single-scale simulation into a single Scale.
-Minimally a Scale possess a single scale-task.
-For example the biotissue simulation (discussed in Section \ref{biotissue}) the Macro Scale has a single scale-task, a parallelized finite element analysis simulation.
-However a Scale may also possess many scale-tasks, each a duplicate of a single-scale simulation.
-Again in the biotissue example case, the Micro\_FO Scale has a number of scale-tasks (individual Representative Volume Element (RVE) simulations) related to the number of Macro mesh elements.
-The management of these scale-tasks the responsibility of the application programmer, though changes made that redistributed terms modelled in any DataDistributions should be noted and used to update appropriately once dynamic/adaptive operations have ceased.
-Adding/removing terms from a DataDistribution may result in the requirement that scale-tasks be instantiated/deleted to accommodate the modification.
-
-When DataDistributions are modified the modifications must be reflected across the RankSet associated with the Scale on which the DataDistribution is defined.
-This procedure is called Assembly, where each rank contributes the local (and hence valid) portion of the control data to arrive at a valid Scale-global value.
-In practice this reduces to an MPI\_Allgather operation for the DataDistribution, but for other types of AMSI control data (most specifically the CouplingPattern) the implementation of the Assembly operation requires additional operations.
-
-When CouplingPatterns are modified (as a result of a changing DataDistribution) the Scale informing the modification must also ensure that the coupled Scale receives the corrected CouplingPattern.
-This process is called Reconciliation, and has two phases the inter-scale phase and the local-scale phase.
-In the inter-scale phase a sparse representation of the relevant CouplingPattern data (relevant meaning related to the rank receiving the data, informing that rank fully of only the role it plays in the communication) is communicated from the modifying scale to the modified scale.
-In the scale-local phase a specialized Assembly operation is performed on the received partial updates to update the full CouplingPattern representation on all processes in the modified Scale.
-
-All communication and planning operations in the AMSI libraries use an implicit local numbering for data distribution terms, the index associated with the term in a local serialized buffer.
-Usage only of implicit ordering was decided upon in order to reduce the memory impact as the simulation scales, as any meta-data which scales with the simulation will necessarily pose a memory issue once scaling hits some threshold.
-
-\textbf{[major revisions end here, though there are sporadic edits herafter]}
-%% revised up to here, with sporadic edits after
-
-Since scale-tasks are an abstraction the AMSI systems cannot provide direct operations on scale-tasks, but provide mechanisms for the application to inform AMSI of such operations when they occur.
-
-Additionally, scale-tasks associated with specific subsets of a process set (either a single process or several processes) may be migrated in part or in whole.
-Migrating a scale-task in part amounts to using load balancing facilities on the associated processes to redistribute all data scale-task from a scale-task to a subset of the previously associated processes, allowing the freed processes to be assigned data from other scale-tasks.
-Migrating a scale task in whole constitutes using load balancing operations on the process set associated with the entire scale to move all pertinent data between subset of the managed process set.
-AMSI provides operations to enact load-balancing operations from plans such as those generated by Zoltan algorithms.
-In the current implementation these operations are collective on the process set associated with the managed scale.
-
-While the primary scale (the scale associated with producing the solution for the overall simulation) is often thought of as controlling the tertiary scales in terms of dynamic instantiation, this is not assumed in the AMSI implementation.
-Any scale is capable of informing the process set associated with a coupled scale of the addition or deletion of individual scale-tasks.
-At present it is dependent on the developer to implement the capability of a scale to accept additional scale-tasks.
-
-Scale-coupling communication is guided by data distributions and communication patterns.
-A data distribution is a representation of individual units of scale-coupling data distributed across a single scale-task.
-The format of the coupling data is arbitrary (so long as it can be serialized and provided for use by MPI communication primitives).
-A data distribution represents the smallest unit of data important to the scale-coupling communication, which is necessarily distinct for various multi-scale use cases.
-
-Scale-linking quantities are dynamically registered with the AMSI system using a data distribution defined on a scale by the scale-tasks operating at that scale.
-Production of a communication pattern is required prior to executing a scale-coupling communication, and the pattern must be reconciled (synchronized across the coupling relationship) prior to being used to guide a coupling communication operation (and when the communication pattern is altered by the scale providing the coupling data).
-Several simple communication pattern production algorithms are provided by AMSI, but these are relatively trivial due to the nature of scale-coupling communications and the general nature of the AMSI libraries.
-Hooks are provided for users to implement algorithms for the construction of communication patterns tailored to the specifics of their problem.
-
-At present an application developer must implement up-/down-scaling operations required for their particular multi-scale use case, hooks will be provided in future releases to automate application of such operations during scale-coupling.
-The current implementation includes blocking versions of the scale-coupling communication operations.
-Asynchronous versions have been designed but are not currently implemented due to lack of immediate need in the primary AMSI use case.
-
-\section{Soft Tissue Simulation}\label{biotissue}
-The Biotissue code is a multi-scale simulation of soft organic tissues, consisting of an engineering-scale simulation using the finite element method, and a micro-scale quasistatics force-balance simulation.
-The Cauchy Momentum Balance Equation for a body in static equilibrium is used as the governing equation for the macro-scale simulation.
-%
-\begin{align}
-\nabla \cdot \pmb{\sigma} =& \; 0 \text{ in } \Omega \label{momentum_balance} \nonumber \\
-\pmb{\sigma} \cdot \pmb{n} =& \; \pmb{t} \text{ on } \partial \Omega \nonumber
-\end{align}
-%
-Specific quantities needed to compute the elemental tangent stiffness matrices and force vectors are instead supplied by the micro-scale force-balance simulations, which occur at each numerical integration point in the engineering-scale mesh \ref{biotissue_hierarchy}.
-%
-\begin{figure}
-  \begin{center}
-    \includegraphics[height=2in]{biotissue_visual_hierarchy.eps}
-  \end{center}
-  \caption{\small biotissue multi-scale problem}
-  \label{biotissue_hierarchy}
-\end{figure}
-%
-
-The incremental displacements (deltas generated by the previous Newton-Raphson iteration) on each engineering-scale finite element node are used to distort the a dimensionless representative volume element (RVE) through a dimensionalized abstraction of the RVE embedded in the engineering domain and an associated displacement mapping term $\frac{d\mathbf{RVE}}{d\mathbf{FE}}$, the first order relation between the incremental displacement of an RVE with respect to the displacement of the associated mesh element.
-This results in each micro-scale node being displaced.
-Each RVE is a dimensionless unit cube containing a connected collagen fiber network, for the present results the fibers are implemented as truss elements and their distribution is generated using Delauney triangulation, resulting in a high coordination number at each node and consequently a numerically stable global Jacobian.
-The micro-scale boundary condition necessitates that the force exerted by each fiber network node located on the boundary of the RVE be zero.
-The micro-scale global Jacobian is formulated along with the force vector, wherein the boundary condition is enforced by setting specific rows of the force vector associated with boundary nodes to zero.
-
-Another Newton-Raphson iterative process occurs at micro-scale to converge to the set of displacements resulting in force equilibrium for all the fibers within the fiber network constituting the RVE.
-After a micro-scale simulation converges to a solution, the Jacobian and force vector are then evaluated again at the solution position, the force exerted along the boundary of the RVE is summed for each of the principal and shear components of a stress tensor, which is then dimensionalized and sent back to the engineering-scale for use in the elemental tangent stiffness matrix formulation.
-More detail into the derivation of this multi-scale system can be found in \cite{stylianopoulos2008thesis} \cite{agoram2001coupled} \cite{stylianopoulos2007multiscale} .
-Discussion of the dimensionalization of the dimensionless force terms produced by each fiber network can be found in \cite{stylianopoulos2007volume} \cite{chandran2007deterministic}.
-
-Thus the multi-scale simulation consists of two nested Newton-Raphson operations occurring; during each macro-scale Newton iteration every micro-scale RVE must undergo a full Newton-Raphson convergence process.
-Thus the micro-scale iterative processes will tend to dominate the total computational effort required to fully converge to a solution.
+\subsection{Parallel Operations}\label{parallel}
 
 \section{Results and Discussion}\label{results}
-The biotissue simulation was run on the AMOS BG/Q machine housed at the CCI facilities.
-Both strong and weak scaling studies were conducted up to an eighth of the total machine core count, or 16384 cores.
-The ratio of macro-scale to micro-scale assigned processes was held at 1:127, and the total core count was kept as an integer power of 2 in order for the SLURM job scheduler to submit the jobs to the compute nodes in a timely manner.
-The precise calculation used to determine macro and micro process counts is given by
-%
-%\begin{lstlisting}[language=Bash,caption={process assignment calculation}]
-\begin{verbatim}
-NP=$(( $NUM_MACRO * 128))
-NUM_MICRO=$(( $NP - $NUM_MACRO))
-\end{verbatim}
-%\end{lstlisting}
-\label{macro micro ratio}
-%
-
-Separate timings were taken for various sections of the simulation code, isolating the micro-scale computation, macro-scale linear system assembly and calculation of elemental linear systems, and macro-scale linear system solution.
-In addition, the time for spent in the AMSI subroutines for communication was also measured.
-Thus the behavior of each section of the code is independently evaluated in our scaling study, as well as the overall execution time.
-
-%
-\begin{figure}
-  \begin{center}
-    \includegraphics[height=2in]{ws_phil.eps}
-  \end{center}
-  \caption{\small weak scaling}
-  \label{weak scaling}
-\end{figure}
-%
-Keeping the work per process approximately constant (1k elements per macro-scale process, with the 1:127 macro-micro ratio as specified in listing \ref{macro micro ratio}) it can be seen that the time to compute every RVE in the system stays relatively consistent, while the time to solve the macro-scale linear system increases, causing a modest increase in the overall time per macro-scale Newton iteration.
-
-%
-\begin{figure}
-  \begin{center}
-    \includegraphics[height=2in]{ss_phil.eps}
-  \end{center}
-  \caption{\small strong scaling}
-  \label{strong scaling}
-\end{figure}
-%
-As the problem size is held constant with respect to the number of individual elements in the engineering scale domain discretization the simulation displays near optimal convergence to the macro-scale solve time alone, which is also reducing in time cost.
-The limiting factor in the speedup of the overall simulation is the potential speedup of the macro-scale simulation rather than the micro-scale simulation with fiber-only RVEs.
-As indicated in the weak scaling results this is related primarily to the macro-scale linear solve communication time increasing.
-At present an incomplete LU factorization is used to solve the linear system formulated at macro-scale each Newton iteraion (using PETSc configured with the superLU\_dist solver \cite{li05superlu} \cite{lishao10superluilu} \cite{lidemmel03superludist} ).
-Using instead an iterative solution method to reach a desired level of accuracy may result in stronger scaling at macro-scale, but as yet this approach has been unneeded due to the preponderance of time spent computing micro-scale results in use cases.
-
-Many problems of interest will possess finite elements numbered in the tens to hundreds of millions of degrees of freedom, such that a suitable partitioning of the mesh to cores on the machine in terms of elements per partition (and directly the macro-scale time-to-solution) may require a substantial fraction of the total cores on the machine for macro-scale to execute in reasonable time.
-AMSI provides services for dynamically adding and removing individual scale-tasks to a specific scale, which can be utilized to limit RVE usage to only those locations in the multi-scale mesh most influential on the overall evolution of the simulation.
-
-AMSI-specific results of some sort...?
-
-Could include utilization/idle time metrics, but those are far more relevant to the load balancing paper...
 
@@
 
-The Adaptive Multi-scale Simulation Infrastructure provides a valuable set of tools to facilitate the implementation and execution of multi-scale simulations without imposing restrictions on the application developer.
-Usage of the core operations of the AMSI libraries does not require adherence to a specific set of interfaces nor usage of an abstract modeling language, which complicate inclusion of legacy components into multi-scale simulations.
-
-Scale-sensitive load-balancing operations have been developed for the biotissue multi-scale test problem.
-Unfortunately the problem does not exhibit load persistence in the current parallel execution space task discretization, so an alternative parallel discretization such as overdecomposition with work-stealing to enforce more local balancing operations and occasional global balancing operations is being explored.
-
-Development of libraries concerned with higher levels of abstraction of the multi-scale problem will also be valuable in easing the development of new multi-scale simulations from the perspective of the application programmer.
-High level abstractions such as the Scale Separation Map in \cite{chopard2011framework} could provide mechanisms to provide initial configuration information to the AMSI system during the initialization phase, as well as assist in the implementation of (or usage of existing library-provided) up-/down-scaling operations.
-
-An RVE simulation taking into account inter-fibril matrix material has been developed \cite{lake2012mechanics} \cite{zhang2013cross} \cite{zhang2013coupled}.
-This alternative RVE will be incorporated into the multi-scale model as an alternative to the fiber-only RVE, to be used in areas of the macro-scale domain determined to be most critical to the overall evolution of the global system, or where error accumulation through model inaccuracy is to be minimized.
-As this fiber-matrix RVE represents a more significant source of computational complexity, it is independently parallelized -- in contrast to the fiber-only RVE -- and will represent a compelling use case for the parallel execution space discretization and scale-coupling communication management capabilities of the AMSI libraries.
-
 \section{Acknowledgements}
 
-\bibliography{citations}
+\bibliography{multiscale_theory,multiscale_frameworks,load_balancing,parallel,software,biotissue}
 

### 2016-10-20 interface/FiberOnlyIntegrator.h
don't know where this came from, or how long it has been here, and it is definitely not in the build and probably should go in bio... but I'm adding it to be safe before it delete it

@@ -1,0 +1,91 @@
+#ifndef BIO_FO_INTEGRATOR_H_
+#define BIO_FO_INTEGRATOR_H_
+#include <apf.h>
+#include <amsiFEA.h>
+namespace bio
+{
+  void makeS(apf::DynamicVector & s,
+             apf::DynamicMatrix & S)
+  {
+    S.resize(9,9);
+    S.zero();
+    S(0,0) = S(3,3) = S(6,6) = s(0);
+    S(0,1) = S(3,4) = S(6,7) = s(1);
+    S(0,2) = S(3,5) = S(6,8) = s(2);
+    S(1,0) = S(4,3) = S(7,6) = s(1);
+    S(1,1) = S(4,4) = S(7,7) = s(3);
+    S(1,2) = S(4,5) = S(7,8) = s(4);
+    S(2,0) = S(5,3) = S(8,6) = s(2);
+    S(2,1) = S(5,4) = S(8,7) = s(4);
+    S(2,2) = S(5,5) = S(8,8) = s(5);
+  }
+  class FiberOnlyIntegrator : public apf::Integrator
+  {
+  protected:
+    amsi::ElementalSystem es;
+    apf::Field * du;
+    apf::Field * rve;
+    apf::Element * e;
+    int dim;
+    int nedofs;
+    int nenodes;
+    int gid;
+  public:
+    FiberOnlyIntegrator(int o,
+                        apf::Field * u,
+                        apf::Field * r)
+      : apf::Integrator(o)
+      , du(u)
+      , rve(r)
+      , es()
+      , e(NULL)
+      , dim(0)
+      , nedof(0)
+      , nenode(0)
+      , gid(0)
+    {
+      assert(f);
+    }
+    void inElement(apf::MeshElement * melmnt)
+    {
+      e = apf::createElement(fld,melmnt);
+      nenode = apf::countNodes(e);
+      dim = apf::getDimension(melmnt);
+      nedof = nenode * dim; //eh....
+      es.resize(nedof);
+      gid = 0;
+    }
+    void outElement();
+    void atPoint(apf::Vector3 const & p, double w, double dV)
+    {
+      //rveinfo * rve_info = ;
+      apf::NewArray<apf::Vector3> dN_dxi;
+      apf::getShapeGrads(e,p,dN_dxi);
+      apf::DynamicMatrix BL(6,nenode);
[…]

### 2016-10-20 interface/FiberOnlyIntegrator.h
removing unknown file

@@ -1,91 +1,0 @@
-#ifndef BIO_FO_INTEGRATOR_H_
-#define BIO_FO_INTEGRATOR_H_
-#include <apf.h>
-#include <amsiFEA.h>
-namespace bio
-{
-  void makeS(apf::DynamicVector & s,
-             apf::DynamicMatrix & S)
-  {
-    S.resize(9,9);
-    S.zero();
-    S(0,0) = S(3,3) = S(6,6) = s(0);
-    S(0,1) = S(3,4) = S(6,7) = s(1);
-    S(0,2) = S(3,5) = S(6,8) = s(2);
-    S(1,0) = S(4,3) = S(7,6) = s(1);
-    S(1,1) = S(4,4) = S(7,7) = s(3);
-    S(1,2) = S(4,5) = S(7,8) = s(4);
-    S(2,0) = S(5,3) = S(8,6) = s(2);
-    S(2,1) = S(5,4) = S(8,7) = s(4);
-    S(2,2) = S(5,5) = S(8,8) = s(5);
-  }
-  class FiberOnlyIntegrator : public apf::Integrator
-  {
-  protected:
-    amsi::ElementalSystem es;
-    apf::Field * du;
-    apf::Field * rve;
-    apf::Element * e;
-    int dim;
-    int nedofs;
-    int nenodes;
-    int gid;
-  public:
-    FiberOnlyIntegrator(int o,
-                        apf::Field * u,
-                        apf::Field * r)
-      : apf::Integrator(o)
-      , du(u)
-      , rve(r)
-      , es()
-      , e(NULL)
-      , dim(0)
-      , nedof(0)
-      , nenode(0)
-      , gid(0)
-    {
-      assert(f);
-    }
-    void inElement(apf::MeshElement * melmnt)
-    {
-      e = apf::createElement(fld,melmnt);
-      nenode = apf::countNodes(e);
-      dim = apf::getDimension(melmnt);
-      nedof = nenode * dim; //eh....
-      es.resize(nedof);
-      gid = 0;
-    }
-    void outElement();
-    void atPoint(apf::Vector3 const & p, double w, double dV)
-    {
-      //rveinfo * rve_info = ;
-      apf::NewArray<apf::Vector3> dN_dxi;
-      apf::getShapeGrads(e,p,dN_dxi);
-      apf::DynamicMatrix BL(6,nenode);
[…]

### 2017-09-07 doc/articles/amsi/amsi.tex
general paper work

% SISC 3 : Software and High-Performance Computing. Papers in this category should concern the novel design and development of computational methods and high-quality software, parallel algorithms, high-performance computing issues, new architectures, data analysis, or visualization. The primary focus should be on computational methods that have potentially large impact for an important class of scientific or engineering problems.

\title{The Adaptive Multiscale Simulation Infrastructure - Methods and Soft-tissue Example Simulation}
\author{W.R. Tobin, V.W.L. Chan, M.S. Shephard}
\date{\today}
\documentclass[review]{siamart1116}

\begin{document}
\maketitle

\begin{abstract}
Discuss multi-scale and frameworks, show how AMSI does multi-scale, and how it is different than others. Introduce the soft-tissue problem. Discuss hierarchical multi-scale and relate it to well-known multi-scale models, and what requirements it imposes. Discuss how AMSI works, and how to use AMSI to address the requirements of a hierarchical multi-scale problem. Show how AMSI was used to construct the soft-tissue problem. Show results somehow...

\end{abstract}

Targeting the SIAM Journal On Scientific Computing

\section{Introduction}\label{sec:intro}

\subsection{Multi-scale Simulation}\label{sec:multiscale}

Boilerplate about why multi-scale simulation is required, what issues it attempts to confront.

Introduce concept that multi-scale is difficult from an implemenetation and computational efficiency standpoint, while there has been a proliferation in the development of actual physical coupling models through largely ad-hoc implementations.

Briefly discuss hierarchical multi-scale vs concurrent multi-scale, mention that AMSI is currently focused on providing support for hierarchical multi-scale problems, though leaving development paths open for concurrent.

Discuss 2-3 other multi-scale frameworks (UINTAH, MPCCI, one more..?)

\subsection{AMSI}\label{sec:amsi}

Introduce AMSI, show how it intends to adress the issues of difficult implementation as a top priority while attempting to maintain computational efficiency as well.

Show how the AMSI approach is different from the other multi-scale frameworks discussed above.

\subsection{Soft-Tissue Problem}\label{sec:bio}

Brief overview of the multi-scale structure of AMSI with reference back to papers covering the physics and some results particular to the physics in greater detail (the neuron paper).

\section{Methods}\label{sec:methods}

Maybe add a section before \ref{sec:hierarchic} taking the most generic parts of \ref{sec:amsi-construction} and discussing them first, before moving into hierarchical in particular?

\subsection{General Hierarchical Multi-scale}\label{sec:hierarchic}

Dig a little deeper into hierarchical multi-scale, discuss the computational aspects of this type of multi-scale, and reference some well-known multi-scale models that fall into this category.

\subsection{AMSI Simulation Construction}\label{sec:amsi-construction}

Discuss how AMSI operates in general, then go in to how to use AMSI to address the needs of a hierarchical multi-scale problem.

\subsection{Soft-Tissue Implementation}\label{sec:bio-construction}

Show how AMSI was used with the existing single-scale simulations to construct the biotissue problem.

\section{Results}\label{sec:results}

What constitutes a good set of results? It likely mostly depends on how we describe our difference from the existing frameworks and what our key goals in AMSI are

  - We mostly focus on the ease of implementing a multi-scale simulation from existing simulations - but this is hard to develop a metric for

  - We also want computational efficiency, and while we can calculate our overall compute efficiency for any run we don't really have anything to compare against

\section{Future Work}\label{sec:future-work}

Brief overview of obstacles being worked on towards concurrent multi-scale.

\bibliographystyle{siamplain}
\bibliography{amsi}
\end{document}

### 2017-09-07 multiscale/amsiControlService_impl.h
add amSender and amRecver to simplify communication algorithms and make things clearer overall

@@ -205,14 +205,14 @@ namespace amsi
     Task * tl = task_man->getLocalTask();
     int task_rank = tl->localRank();
     Task * t1 = task_man->Task_Get(t_ids.first);
-    Task * t2 = task_man->Task_Get(t_ids.second);
+    //Task * t2 = task_man->Task_Get(t_ids.second);
     int t1s = taskSize(t1);
     int tp_sz = 0;
     MPI_Type_size(tp,&tp_sz);
     PCU_Switch_Comm(comm_man->CommRelation_GetInterComm(r_dd_id.first));
     PCU_Comm_Begin();
     CommPattern * ptrn = NULL; //t1->getLocalDDValue(r_dd_id.second);
-    if(tl == t1)
+    if(amSender(rdd_id))
     {
       unsigned cnt_snt_frm = countRanksSentToFrom(ptrn,task_rank);
       std::vector<int> snt_rnks(cnt_snt_frm);
@@ -248,7 +248,7 @@ namespace amsi
       size_t rcv_sz = 0;
       while(PCU_Comm_Read(&frm,&rcv,&rcv_sz)) { }
     }
-    else
+    if(amRecver(rdd_id))
     {
       PCU_Comm_Send();
       size_t bfr_sz = 0;
@@ -323,8 +323,7 @@ namespace amsi
       // If user indicated a user specified load balancing function but did not provide one
       if(option == -1 && userFunc == NULL)
         option = 0;
-      // If sending task
-      if(tl == t1)
+      if(amSender(rdd_id))
       {
         DataDistribution dd_delta(t1s,task_rank);
         dd_delta = data.size();
@@ -358,8 +357,6 @@ namespace amsi
         (*algo)(pattern,delta_pattern,&dd_delta);
         CommPattern_Assemble(rdd_id);
         CommPattern_Assemble(delta_id);
-        
-        
         // Reorder objects vector based on comm patterns (Need to fill data indices? Probably not)
         // Trying to make something that will work for both vectors and lists...
         int numOldData = objects.size() - data.size();
@@ -393,13 +390,13 @@ namespace amsi
         CommPattern_Reconcile(rdd_id);
         CommPattern_Reconcile(delta_id);
       }
-      else
+      if(amRecver(rdd_id))
       {
         // Get the new orig and init comm patterns
         CommPattern_Reconcile(rdd_id);
         CommPattern_Reconcile(delta_id);
         // Fill data vector with appropriate indices
-        CommPattern * pattern = comm_man->getCommPattern(rdd_id);
[…]

### 2017-09-07 multiscale/amsiAddData.cc
use vectors instead of variably-sized arrays with are not part of the standard but a common compiler extension

@@ -19,7 +19,7 @@ namespace amsi
   {
     int s1 = cp->getNumSenders();
     int s2 = cp->getNumRecvers();
-    int dd[s1]{};
+    std::vector<int> dd(s1);
     for(int ii = 0; ii < s1; ++ii)
       dd[ii] = (*dd_dlta)[ii];
     int total_data = 0;
@@ -70,7 +70,7 @@ namespace amsi
   {
     int s1 = cp->getNumSenders();
     int s2 = cp->getNumRecvers();
-    int dd[s1]{};
+    std::vector<int> dd(s1);
     for(int ii = 0; ii < s1; ++ii)
       dd[ii] = (*dd_dlta)[ii];
     int cnts[s2];

### 2017-09-07 multiscale/amsiDataDistribution.cc
don't attempt to resize the weighting vectors in a data distribution which is unweighted, this fixes a crash that would occur when assembling an unweighted data distribution

@@ -68,8 +68,9 @@ namespace amsi
     MPI_Allgather(&lcl,1,mpi_type<int>(),
                   &dd[0],1,mpi_type<int>(),
                   comm);
-    for(unsigned ii = 0; ii < dd.size(); ii++)
-      wgts[ii].resize(dd[ii]);
+    if(wgtd)
+      for(unsigned ii = 0; ii < dd.size(); ii++)
+        wgts[ii].resize(dd[ii]);
     assembled = true;
   }
   int DataDistribution::operator[](unsigned idx) const

### 2017-09-07 multiscale/amsiPlanMigration.cc
suppress warnings due to unused parameters

@@ -34,7 +34,7 @@ namespace amsi
     void CommPattern_PlanMigration_Test(std::vector<int> & m_index,
                                         std::vector<int> & m_send_to,
                                         CommPattern * pattern,
-                                        const DataDistribution * dd,
+                                        const DataDistribution *,
                                         int rank,
                                         MPI_Comm task_comm)
     {
@@ -99,7 +99,7 @@ namespace amsi
     void CommPattern_PlanMigration_Full(std::vector<int> & m_index,
                                         std::vector<int> & m_send_to,
                                         CommPattern * pattern,
-                                        const DataDistribution * dd,
+                                        const DataDistribution *,
                                         int rank,
                                         MPI_Comm task_comm)
     {
@@ -172,10 +172,10 @@ namespace amsi
 # ifdef ZOLTAN
   void CommPattern_PlanMigration_Zoltan(std::vector<int> & m_index,
                                         std::vector<int> & m_send_to,
-                                        CommPattern * pattern,
+                                        CommPattern *,
                                         const DataDistribution * dd,
-                                        int rank,
-                                        MPI_Comm task_comm)
+                                        int,
+                                        MPI_Comm)
   {
     Zoltan_Struct * zs = dd->GetZS();
 

### 2017-09-07 multiscale/amsiTask.h
remove nonsensical consts

@@ -20,8 +20,8 @@ namespace amsi
     bool assignedTo() { return local_rank != -1; }
     int localToGlobalRank(int);
     int size();
-    const MPI_Comm comm() {return task_comm;}
-    const MPI_Group group() {return task_group;}
+    MPI_Comm comm() {return task_comm;}
+    MPI_Group group() {return task_group;}
     void setExecutionFunction(ExecuteFunc e) {exec = e;}
     int execute(int& argc,char**& argv)
     {

### 2017-09-07 analysis/FullLULAS.h
suppress unused parameter warnings

@@ -26,11 +26,11 @@ namespace amsi
     { memset(&matrix[0][0],0,num_rows*num_rows*sizeof(double)); return true; }
     bool ZeroVector()
     { memset(&vector[0],0,num_rows*sizeof(double)); return true; }
-    void GetSolution(double * sol)
+    void GetSolution(double *& sol)
     { sol = &solution[0]; }
-    void GetVector(double * vec)
+    void GetVector(double *& vec)
     { vec = &vector[0]; }
-    void SetVector(double * vec)
+    void SetVector(const double * vec)
     { std::copy(vec, vec + vector.size(), vector.begin()); }
     bool SeperateSolve() {return false;}
   };

### 2017-09-07 analysis/NonFiniteElement.cc
suppress unused parameter warnings

@@ -16,7 +16,7 @@ namespace amsi
           T(ii,jj) += B0T(ii,kk) * hookes[kk][jj];
     apf::multiply(T,B0,K0);
   }
-  void nonLinearStiffness(int num_nodes,
+  void nonLinearStiffness(int,
                           apf::DynamicMatrix & BNL,
                           apf::Matrix3x3 & pk2,
                           apf::DynamicMatrix & KNL)
@@ -57,9 +57,9 @@ namespace amsi
       B(5,3*ii + 2) = disp[0];
     }
   }
-  void computeGreenStrain(apf::DynamicMatrix & H,
-                          apf::DynamicMatrix & y,
-                          apf::Matrix3x3 & E)
+  void computeGreenStrain(apf::DynamicMatrix & ,
+                          apf::DynamicMatrix & ,
+                          apf::Matrix3x3 & )
   {
     /*
       double L[3][3];

### 2017-09-07 analysis/apfSimFEA.cc
remove redundant inherited function override and just use the SimFEA version

@@ -53,13 +53,5 @@ namespace amsi
       SimFEA::Adapt();
     }
   }
-  void apfSimFEA::addFieldToMap(apf::Field * fd)
-  {
-    /*
-    apf::SIMDataOf<double> * field_data = dynamic_cast<apf::SIMDataOf<double>*>(fd->getData());
-    if(field_data != NULL)
-      SimFEA::addFieldToMap(field_data->getSimField());
-    */
-  }
 }
 

### 2017-09-07 analysis/amsiAnalysis.cc
remove nonsense consts and suppress unused parameter warnings

@@ -9,7 +9,7 @@ namespace amsi
 {
   #define ANALYSIS_OPTIONS(OP) OP(petsc), OP(petsc_options), OP(simmetrix), OP(simmetrix_license), OP(num_analysis_options)
   enum AnalysisOptions{ANALYSIS_OPTIONS(MAKE_ENUM_OP)};
-  const char * const getAnalysisOptionString(int ii)
+  const char * getAnalysisOptionString(int ii)
   {
     static const char * const AnalysisOptionStrings[] = {ANALYSIS_OPTIONS(MAKE_STRING_OP)};
     assert(ii < num_analysis_options);
@@ -24,7 +24,7 @@ namespace amsi
     }
     return -1;
   }
-  const char * const getAnalysisSectionString(int ii)
+  const char * getAnalysisSectionString(int ii)
   {
     static const char * const AnalysisConfigSectionStrings[] = {ANALYSIS_CONFIG_SECTIONS(MAKE_STRING_OP)};
     assert(ii < num_analysis_config_sections);
@@ -64,7 +64,7 @@ namespace amsi
     use_simmetrix = true;
     simmetrix_license_file = sim_lic;
   }
-  void simmetrixInit(int argc, char ** argv, MPI_Comm cm)
+  void simmetrixInit(int, char **, MPI_Comm cm)
   {
     PMU_setCommunicator(cm);
     Sim_readLicenseFile(simmetrix_license_file.c_str());

### 2017-09-07 analysis/amsiDummyLAS.h
suppress unused parameter warnings

@@ -62,7 +62,7 @@ namespace amsi
       nrm = sqrt(nrm);
     }
     // stubb
-    virtual void GetAccumSolutionNorm(double & nrm)
+    virtual void GetAccumSolutionNorm(double &)
     {   }
     virtual bool ZeroMatrix()
     {
@@ -94,7 +94,7 @@ namespace amsi
       nrm = sqrt(nrm);
     }
     // stubb
-    virtual void GetAccumVectorNorm(double & nrm)
+    virtual void GetAccumVectorNorm(double & )
     {   }
     // stubb
     virtual void GetDotNorm(double &)

### 2017-09-07 analysis/amsiLinearStressStrainIntegrator.cc
suppress unused parameter warnings

@@ -20,7 +20,7 @@ namespace amsi
       for(int jj = 0; jj < num_field_components; jj++)
         (*U)[ii*num_field_components + jj] = u[ii][jj];
   }
-  void LinearStressStrainIntegrator::atPoint(apf::Vector3 const & p, double w, double dV)
+  void LinearStressStrainIntegrator::atPoint(apf::Vector3 const & p, double, double)
   {
     apf::NewArray<apf::Vector3> grads;
     apf::getShapeGrads(e,p,grads);

### 2017-09-07 analysis/amsiPETScLAS.cc
suppress unused parameter warnings

@@ -8,7 +8,7 @@ namespace amsi
   void PetscLAS::Reinitialize(int num_local_unknowns,
                               int num_global_unknowns,
                               int global_offset,
-                              int * nonzero_per_row)
+                              int * )
   {
     Reinitialize(num_local_unknowns,num_global_unknowns,global_offset);
   }

### 2017-09-07 analysis/amsiPETScLAS2.h

@@ -97,7 +97,7 @@ namespace las
       ops->zero(f);
       ops->zero(K);
     }
-    void Reinitialize(int lcl, int glb, int low, int * nnz)
+    void Reinitialize(int lcl, int glb, int low, int *)
     {
       Reinitialize(lcl,glb,low);
     }

### 2017-09-07 analysis/apfBoundaryConditions_impl.h
suppress unused parameter warnings

@@ -50,7 +50,7 @@ namespace amsi
                         I bgn,
                         I nd,
                         NeumannIntegrator * i,
-                        double t)
+                        double)
   {
     apf::Field * fld = apf::getField(nm);
     for(I it = bgn; it != nd; ++it)

### 2017-09-07 analysis/apfMeasure.cc

@@ -75,7 +75,7 @@ namespace amsi
       dim = apf::getDimension(me);
       mesh_coord_elem = apf::createElement(apf::getMesh(f)->getCoordinateField(),me);
     }
-    void atPoint(apf::Vector3 const &p, double w, double dV)
+    void atPoint(apf::Vector3 const &, double w, double)
     {
       int & nen = nenodes; // = 4 (tets)
       // 1. Get coordinates on underlying mesh
@@ -130,7 +130,7 @@ namespace amsi
       /** We want to consider 3D space, however MeshElement me is for 2D entity. Therefore we need to hardcode dim for now.. */
       dim = 3;
     }
-    void atPoint(apf::Vector3 const &p, double w, double dV)
+    void atPoint(apf::Vector3 const &, double, double)
     {
       int & nen = nenodes; // = 3 (triangle)
       // 1. Get coordinates on underlying mesh

### 2017-09-07 analysis/simBoundaryConditions_impl.h

@@ -53,7 +53,7 @@ namespace amsi
   }
   */
   template <typename I>
-    int applySimDirichletBCs(apf::Numbering * nm, pMesh msh, I bgn, I nd, double t)
+    int applySimDirichletBCs(apf::Numbering * nm, pMesh, I bgn, I nd, double t)
   {
     int fxd = 0;
     for(auto it = bgn; it != nd; it++)
@@ -72,7 +72,7 @@ namespace amsi
     return fxd;
   }
   template <typename I>
-    void applySimNeumannBCs(LAS * las, apf::Numbering * nm, pMesh msh, I bgn, I nd, double t)
+    void applySimNeumannBCs(LAS * las, apf::Numbering * nm, pMesh, I bgn, I nd, double t)
   {
     apf::Field * fld = apf::getField(nm);
     for(auto it = bgn; it != nd; it++)

### 2017-09-07 analysis/simBoundaryConditions.cc
use delete[] instead of delete

@@ -141,7 +141,7 @@ namespace amsi
     else if(isSpaceExpr(ii))
       rslt = AttributeTensor0_evalDS((pAttributeTensor0)att,&args[0]);
     va_end(prms);
-    delete args;
+    delete [] args;
     return rslt;
   }
   double SimTensor1Query::getValue(int ii, ...)

### 2017-09-07 util/amsiDistributed.h
should use the passed in value, not assume false, though I think this class is currently unused

@@ -17,7 +17,7 @@ namespace amsi
   public:
      Distributed(MPI_Comm c = MPI_COMM_NULL, bool v = false)
       : cm(c)
-      , vld(false)
+      , vld(v)
     { }
     bool isLocallyValid() {return vld;}
     bool isValid()

### 2017-09-07 util/amsiTee.h
just adding some comments

@@ -4,6 +4,7 @@
 #include <iostream>
 namespace amsi
 {
+  // a class to facilitate simultaneous printing to two stream buffers
   template <typename char_type, typename traits = std::char_traits<char_type> >
     class basic_teebuf : public std::basic_streambuf<char_type, traits>
   {
@@ -37,7 +38,14 @@ namespace amsi
     std::basic_streambuf<char_type, traits> * sb1;
     std::basic_streambuf<char_type, traits> * sb2;
   };
+  // realization of the template combined buffer class on a basic char
   typedef basic_teebuf<char> teebuf;
+  /**
+   *  @brief a class used to pring to two ostreams, since
+   *         this class is itself an ostream, it can be
+   *         composed to write to an arbitrary number
+   *         of ostreams.
+   */
   class teestream : public std::ostream
   {
   public:

### 2017-09-07 test/analysis/boundary_conditions/neumann/test.cc
suppress some unused variable warnings

@@ -21,16 +21,18 @@ int main(int argc, char ** argv)
   pGModel mdl = GM_load(argv[1],0,NULL);
   pParMesh sm_msh = PM_load(argv[2],mdl,NULL);
   pMesh prt = PM_mesh(sm_msh,0);
+  (void)prt;
   // get all analysis cases
   std::vector<pACase> css;
   amsi::getTypeCases(SModel_attManager(mdl),"analysis",std::back_inserter(css));
   // only run the first attribute case
   amsi::initCase(mdl,css[0]);
   pACase pd = (pACase)AttNode_childByType((pANode)css[0],"problem definition");
+  (void)pd;
   apf::Mesh * msh =  apf::createMesh(sm_msh);
   apf::Field * u = apf::createLagrangeField(msh,"displacement",apf::VECTOR,1);
   apf::Numbering * nm = apf::createNumbering(u);
-  int dofs = apf::NaiveOrder(nm);
+  apf::NaiveOrder(nm);
   /*
   amsi::DummyLAS las(dofs);
   int tps[] = {amsi::NeuBCType::traction,amsi::NeuBCType::pressure};

### 2017-09-07 doc/articles/amsi/docsiamart.tex

\documentclass[final]{siamart1116}

%% ------------------------------------------------------------------
%% Code used in examples, needed to reproduce 
%% ------------------------------------------------------------------
%% Used for \set, used in an example below
\usepackage{braket,amsfonts}

%% Used in table example below
\usepackage{array}

%% Used in table and figure examples below
\usepackage[caption=false]{subfig}
%% Used for papers with subtables created with the subfig package
\captionsetup[subtable]{position=bottom}
\captionsetup[table]{position=bottom}

%% Used for PgfPlots example, shown in the "Figures" section below.
\usepackage{pgfplots}

%% Used for creating new theorems, remarks
\newsiamthm{claim}{Claim}
\newsiamremark{rem}{Remark}
\newsiamremark{expl}{Example}
\newsiamremark{hypothesis}{Hypothesis}
\crefname{hypothesis}{Hypothesis}{Hypotheses}

%% Algorithm style, could alternatively use algpseudocode
\usepackage{algorithmic}

%% For figures
\usepackage{graphicx,epstopdf}

%% For referencing line numbers
\Crefname{ALC@unique}{Line}{Lines}

%% For creating math operators
\usepackage{amsopn}
\DeclareMathOperator{\Range}{Range}

%strongly recommended
\numberwithin{theorem}{section}

%% ------------------------------------------------------------------
%% Macros for in-document examples. These are not meant to reused for
%% SIAM journal papers.
%% ------------------------------------------------------------------
\usepackage{xspace}
\usepackage{bold-extra}
\usepackage[most]{tcolorbox}
\newcommand{\BibTeX}{{\scshape Bib}\TeX\xspace}
\newcounter{example}
\colorlet{texcscolor}{blue!50!black}
\colorlet{texemcolor}{red!70!black}
\colorlet{texpreamble}{red!70!black}
\colorlet{codebackground}{black!25!white!25}

\newcommand\bs{\symbol{'134}} % print backslash in typewriter OT1/T1
\newcommand{\preamble}[2][\small]{\textcolor{texpreamble}{#1\texttt{#2 \emph{\% <- Preamble}}}}

\lstdefinestyle{siamlatex}{%
  style=tcblatex,
  texcsstyle=*\color{texcscolor},
  texcsstyle=[2]\color{texemcolor},
  keywordstyle=[2]\color{texemcolor},
  moretexcs={cref,Cref,maketitle,mathcal,text,headers,email,url},
}

\tcbset{%
  colframe=black!75!white!75,
  coltitle=white,
  colback=codebackground, % bottom/left side
  colbacklower=white, % top/right side
  fonttitle=\bfseries,
  arc=0pt,outer arc=0pt,
  top=1pt,bottom=1pt,left=1mm,right=1mm,middle=1mm,boxsep=1mm,
  leftrule=0.3mm,rightrule=0.3mm,toprule=0.3mm,bottomrule=0.3mm,
  listing options={style=siamlatex}
}

\newtcblisting[use counter=example]{example}[2][]{%
  title={Example~\thetcbcounter: #2},#1}

\newtcbinputlisting[use counter=example]{\examplefile}[3][]{%
  title={Example~\thetcbcounter: #2},listing file={#3},#1}

\DeclareTotalTCBox{\code}{ v O{} }
{ %fontupper=\ttfamily\color{texemcolor},
  fontupper=\ttfamily\color{black},
  nobeforeafter,
  tcbox raise base,
  colback=codebackground,colframe=white,
  top=0pt,bottom=0pt,left=0mm,right=0mm,
  leftrule=0pt,rightrule=0pt,toprule=0mm,bottomrule=0mm,
  boxsep=0.5mm,
  #2}{#1}

% Stretch the pages
\patchcmd\newpage{\vfil}{}{}{}
\flushbottom

%% ------------------------------------------------------------------
%% End of macros for in-document examples. 
%% ------------------------------------------------------------------

%% ------------------------------------------------------------------
%% HEADING INFORMATION
%% ------------------------------------------------------------------
\begin{tcbverbatimwrite}{tmp_\jobname_header.tex}
\title{Guide to Using SIAM's \LaTeX\ Style%
  \thanks{Submitted to the editors DATE.
\funding{Funding information goes here.}}}

\author{Dianne Doe%
  \thanks{Imagination Corp., Chicago, IL (\email{ddoe@imag.com}).}%
  \and
  Paul T. Frank%
  \thanks{Department of Applied Math, Fictional University, Boise, ID
    (\email{ptfrank@fictional.edu}, \email{jesmith@fictional.edu}).}
  \and
  Jane E. Smith%
  \footnotemark[3]
}

% Custom SIAM macro to insert headers
\headers{Guide to Using  SIAM'S \LaTeX\ Style}
{Dianne Doe, Paul T. Frank, and Jane E. Smith}
\end{tcbverbatimwrite}
\input{tmp_\jobname_header.tex}

% Optional: Set up PDF title and authors
\ifpdf
\hypersetup{ pdftitle={Guide to Using  SIAM'S \LaTeX\ Style} }
\fi

%% ------------------------------------------------------------------
%% END HEADING INFORMATION
%% ------------------------------------------------------------------

%% ------------------------------------------------------------------
%% MAIN Document
%% ------------------------------------------------------------------
\begin{document}
\maketitle

%% ------------------------------------------------------------------
%% ABSTRACT
%% ------------------------------------------------------------------
\begin{tcbverbatimwrite}{tmp_\jobname_abstract.tex}
\begin{abstract}
  Documentation is given for use of the SIAM standard \LaTeX\ and \BibTeX\
  macros.  Instructions and suggestions for compliance with SIAM style
  standards are also included. Familiarity with standard \LaTeX\
  commands is assumed.
\end{abstract}

\begin{keywords}
  \LaTeX, \BibTeX, SIAM Journals, Documentation 
\end{keywords}

\begin{AMS}
  00A20 
\end{AMS}
\end{tcbverbatimwrite}
\input{tmp_\jobname_abstract.tex}
%% ------------------------------------------------------------------
%% END HEADER
%% ------------------------------------------------------------------

\section{Introduction}
\label{sec:intro}

This file is documentation for the SIAM \LaTeX\ style, including how
to typeset the main document, the \BibTeX\ file, and any supplementary
material. More information
about SIAM's editorial style can be found in the style manual, available
at \url{https://www.siam.org/journals/pdf/stylemanual.pdf}.
%
The major changes in the SIAM standard class are summarized in \cref{sec:changes}.
%
The SIAM \LaTeX\@ files can be found at
\url{https://www.siam.org/journals/auth-info.php}. The files that
are distributed for the standard macros are given below. 
\begin{itemize}
\item \texttt{siamart1116.cls} (required): Main SIAM standard \LaTeX\ class file.
\item \texttt{siamplain.bst} (required): Bibliographic style file for
  \BibTeX.
\item \texttt{docsiamart.tex}: Produces this documentation.
\item \texttt{references.bib}: \BibTeX\ database for this
  documentation and examples.
\item \texttt{ex\_article.tex}: Template for article.
\item \texttt{ex\_supplement.tex}: Template for supplement.
\item \texttt{ex\_shared.tex}: Template for shared information for
  article and supplement.
\end{itemize}
To use these files, put \texttt{siamart1116.cls} and
\texttt{siamplain.bst} in the directory with your
paper or, alternatively, into your \LaTeX\@ and \BibTeX\@ paths,
respectively. 
%
The outline of a SIAM \LaTeX\ article is shown in
\cref{ex:outline}. Templates are provided and discussed in more detail
in \cref{sec:template}.

\begin{example}[label={ex:outline},listing only,%
  listing options={style=siamlatex,{morekeywords=[1]{maketitle},
      morekeywords=[2]{siamart1116}},}]%
  {Document outline}
\documentclass{siamart1116}
% Preamble consisting of:
% Packages and macro definitions go here.
% Define title, authors, headers here.
\begin{document}
\maketitle
% Other front matter goes here: abstract, keywords, AMS subjects.
% Main body goes here.
% Appendices goes here (optional).
% Acknowledgements go here (optional).
% Bibliography goes here.
\end{document}
\end{example}

\section{Class options}
\label{sec:class-options}

Class options can be included in the bracketed argument of the
command, separated by commas. The possible class options are:
\begin{itemize}
\item \code{review} --- Recommended for submitting your manuscript to
  a SIAM journal.  Adds line numbers as well as the statement ``This
  manuscript is for review purposes only'' to the bottom of each page.
\item \code{final} --- Turns off the black boxes that help authors
  identify lines that are too long. The final published version will
  have this option on.
\item \code{supplement} --- Specifies that the file is a supplement
  and not the main document, causing changes in the appearance of the
  title and numbering; see \cref{sec:supplement} for details.
\item \code{hidelinks} --- Turns off colors on hyperlinks; see
  \cref{sec:cr+hyp}. The hyperlinks still exist, but there is no color
  to differentiate them.
  The final published version will have this option on.
\end{itemize}

\section{Front matter}
\label{sec:front}
The title and author parts are formatted using the standard
\code{\title}, \code{\author}, and \code{\maketitle} commands as
described in Lamport \cite{La86}. The title and author should be
declared in the preamble. The title and author names are automatically
converted to uppercase in the document.
%
If there is more than one author, each additional author should be preceded by
the \code{\and} command.  
%
The addresses and support acknowledgments are added via
\code{\thanks}. Each author's thanks should specify their address.
The support acknowledgment should be put in the title thanks,
unless specific support needs to be specified for individual authors,
in which case it should follow the author address.  
%
The header for this file was produced by the code in \cref{ex:header}, including an
example of a shared footnote. Each thanks produces a footnote, so the
footnote of the second author is \#3.
%
The command
\code{\headers{title}{authors}} command, with
the title (possibly shortened to fit) and the authors' names, creates
the page headers, automatically converted to uppercase.

\examplefile[label={ex:header},listing only,%
listing options={style=siamlatex,%
deletetexcs={and,thanks,title,author},%
{moretexcs=[2]{and,thanks,title,author,maketitle,headers,email}}}
]{Title and authors in preamble}{tmp_\jobname_header.tex}

\newpage
Following the author and title is the abstract, key words listing, and AMS subject
classifications, designated using the \code{abstract}, \code{keywords}, and \code{AMS}
environments. 
Authors are responsible for providing AMS numbers which can be found
on the AMS web site \cite{AMSMSC2010}. The abstract, keywords, and AMS subject classifications
for this document are specified in \cref{ex:abstract}.

\examplefile[label={ex:abstract},%
before upper={\preamble{\bs newcommand\{\bs BibTeX\}\{\{\bs scshape Bib\}\bs TeX\bs xspace\}}},
listing only,%
listing options={style=siamlatex,%
{morekeywords=[2]{abstract,keywords,AMS}}}
]{Abstract, keywords, and AMS classifications}{tmp_\jobname_abstract.tex}

A more complete example, including a PDF supplement, that uses the
included files \texttt{ex\_article.tex}, \texttt{ex\_supplement.tex},
and \texttt{ex\_shared.tex} is discussed in
\cref{sec:template}. The example files can be used as a starting point
for producing a document.

\section{Cross references and hyperlinks}
\label{sec:cr+hyp}

SIAM now supports cross references and hyperlinks via the
\texttt{cleveref} and \texttt{hyperef} packages, which are loaded by
the class file.

\subsection{Cleveref}
\label{sec:cleveref}

SIAM strongly recommends using the commands provided by 
the \texttt{cleveref} package for cross referencing.
The package is automatically loaded and already customized
to adhere to SIAM's style guidelines. 
%
To create a cross reference, use the command \code{\cref} (inside
sentence) or \code{\Cref} (beginning of a sentence) in place of the
object name and \code{\ref}. 
%
The \texttt{cleveref}
package enhances \LaTeX's cross-referencing features, allowing
the format of cross references to be determined automatically
according to the ``type" of cross reference (equation, section, etc.)
and the context in which the cross reference is used.
%
So, the package
\emph{automatically} inserts the object name as well as the
appropriate hyperlink; see \cref{ex:cref}.
It may require two \LaTeX\@ compilations for the references to show up
correctly.
Additional examples are shown in the sections below
for equations, tables, figures, sections, etc.

\begin{example}[label=ex:cref,bicolor,listing options={style=siamlatex,%
{morekeywords=[2]{cref,ref}}}]{Advantage of using cleveref}
The normal way to get a cross reference with a hyperlink requires a
lot of typing: \hyperref[thm:mvt]{Theorem~\ref*{thm:mvt}}.
The \texttt{cleveref} package gets both the name and hyperlink
automatically using a single macro: \cref{thm:mvt}.
It also handles multiple references with the same macro, such as
\cref{thm:mvt,fig:pgfplots,fig:testfig}.
\end{example}

%

%\newpage
\subsection{Hyperef}
\label{sec:hyperef}

Hyperlinks are created with the \code{\href} and \code{\url} commands,
as shown in \cref{ex:href}.
SIAM has also defined the \code{\email} command, as shown in
\cref{ex:header}.
You can hide links (i.e., turn off link colors) with the \code{hidelinks} option.

\begin{example}[label={ex:href},bicolor,%
listing options={style=siamlatex,%
{morekeywords=[2]{href,url}}}]{Creating hyperlinks}
The \href{https://www.siam.org}{SIAM homepage} has general information.
Note that the colored text will \emph{not} appear in the print version
nor will the hyperlink be active, so the writer may want to specify
the location explicitly instead by using \url{https://www.siam.org}.
\end{example}

Note that homepage links via \code{\url} in the \code{\thanks} environment require
special formatting for the tilde (\string~) character. The formatting
is used in the template and shown in \cref{ex:shared}.

\section{Math and equations}
\label{sec:math}
Here we show some example equations, with numbering, and examples of
referencing the equations. SIAM now includes the package 
\texttt{amsmath} by default, and we include some of its features as
well, although the reader should consult the package user manual for
further guidance \cite{amsmath,shortmath}.
 Several of the example are adapted 
from Mittlebach and Goossen's guide to \LaTeX~\cite{MiGo04}.

\Cref{ex:textmath} is a straightforward example of inline mathematics
equations that does not use any special packages or features.

\begin{example}[label={ex:textmath},bicolor]{Inline math}
The following shows an example of math in text: 
Let $S=[s_{ij}]$ ($1\leq i,j\leq n$) be a $(0,1,-1)$-matrix of order $n$.
\end{example}

In \cref{ex:bbm}, we show the recommended method for getting
blackboard fonts using the \texttt{amsfonts} package. This is not
loaded by default and must be included in the preamble.

\begin{example}[label={ex:bbm},bicolor,before upper={\preamble{\bs
      usepackage\{amsfonts\}}},%
listing options={style=siamlatex,%
{morekeywords=[2]{mathbb}}}]{Blackboard math}
Blackboard bold characters, such as $\mathbb{C}$ and $\mathbb{R}$,
should be created with the \texttt{amsfonts} package, although this
is not included by default.
\end{example}

%\newpage
\Cref{ex:smallmatrix} shows the \code{smallmatrix} environment for an
inline matrix from the \texttt{amsmath} package, which is included by
default.

\begin{example}[label={ex:smallmatrix},bicolor,%
listing options={style=siamlatex,%
{morekeywords=[2]{smallmatrix}}}]{Inline matrix}
Matrices of no more than two rows appearing in text can be created
as shown in the next example:
$B = \bigl[ \begin{smallmatrix} B_{11} & B_{12} \\ 
  B_{21} & B_{22} \end{smallmatrix} \bigr]$.
\end{example}

Bigger matrices can be rendered with environments from
the \texttt{amsmath} package, such as \code{bmatrix} and
\code{pmatrix} used in
\cref{ex:matrices}.

\begin{example}[label={ex:matrices},bicolor,%
listing options={style=siamlatex,%
{morekeywords=[2]{bmatrix,pmatrix}}}]{Creating matrices}
Display matrices can be rendered using environments from \texttt{amsmath}:
\begin{equation}\label{eq:matrices}
  S=\begin{bmatrix}1&0\\0&0\end{bmatrix}
  \quad\text{and}\quad
  C=\begin{pmatrix}1&1&0\\1&1&0\\0&0&0\end{pmatrix}.
\end{equation}
\Cref{eq:matrices} shows some example matrices.
\end{example}

\newpage
\Cref{ex:dmo} shows how to use the \code{\DeclareMathOperator} command
from the \texttt{amsopn} package to declare the \code{\Range} macro. 
(This example also uses the \texttt{braket} package for the
\code{\set} macro, but this is not necessarily recommended by SIAM.)
%
\begin{example}[label={ex:dmo},%
before upper={\preamble{\bs usepackage\{braket,amsfonts,amsopn\}}\\
\noindent\preamble{\bs DeclareMathOperator\{\bs Range\}\{Range\}}},%
bicolor,%
listing options={style=siamlatex,%
{moretexcs=[2]{Range}}}
]{Declaring math operators}
An example of a math operator:
\begin{equation}\label{eq:range}
  \Range(A) = \set{ y \in \mathbb{R}^n | y = Ax }.
\end{equation}
\end{example}

%\newpage
\Cref{ex:foo} shows how to use the \code{align} environment from
\texttt{amsmath} to easily align multiple equations.

\begin{example}[label={ex:foo},bicolor,%
listing options={style=siamlatex,%
{morekeywords=[2]{align}}}]{Aligned equations}
\Cref{eq:a,eq:b,eq:c} show three aligned equations.
\begin{align} 
  f &= g, \label{eq:a} \\
  f' &= g', \quad\text{and} \label{eq:b} \\
  \mathcal{L}f &= \mathcal{L}g \label{eq:c}.
\end{align}
\end{example}

Another way to number a set of equations is the
\code{subequations} environment from \texttt{amsmath}, as shown in \cref{ex:aligned}.

\begin{example}[label={ex:aligned},bicolor,%
listing options={style=siamlatex,%
{morekeywords=[2]{subequations}}}]{Subequations}
We calculate the Fr\'{e}chet derivative of $F$ as follows:
\begin{subequations}
\begin{align}
  F'(U,V)(H,K) 
  &= \langle R(U,V),H\Sigma V^{T} + U\Sigma K^{T} -
  P(H\Sigma V^{T} + U\Sigma K^{T})\rangle \label{eq:aa} \\
  &= \langle R(U,V),H\Sigma V^{T} + U\Sigma K^{T}\rangle 
  \nonumber \\
  &= \langle R(U,V)V\Sigma^{T},H\rangle + 
  \langle \Sigma^{T}U^{T}R(U,V),K^{T}\rangle. \label{eq:bb}
\end{align}
\end{subequations}
\Cref{eq:aa} is the first line, and \cref{eq:bb} is the last line.
\end{example}

%\newpage
~

For an equation split over multiple lines, \cref{ex:ml} shows the
usage of the \code{multline} environment provided by \texttt{amsmath}.

~

\begin{example}[label={ex:ml},bicolor,%
listing options={style=siamlatex,%
{morekeywords=[2]{multline}}}]{Equation split across lines}
We claim that the projection $g(U,V)$ is given by the pair of matrices:
\begin{multline} \label{eq:ml}
  g(U,V) = \biggl( \frac{R(U,V)V\Sigma^{T}U^{T} 
    - U\Sigma V^{T}R(U,V)^{T}}{2}U,\\
  \frac{R(U,V)^{T}U\Sigma V^{T}-V \Sigma^{T}U^{T}R(U,V)}{2}V \biggr).
\end{multline}
\end{example}

\section{Theorem-like environments}
\label{sec:thm}
SIAM loads \texttt{ntheorem} package and uses it to define the following
theorem-like environments:
\code{theorem}, 
\code{lemma}, 
\code{corollary},
\code{definition}, and
\code{proposition}.
SIAM also defines a \code{proof} environment that automatically
inserts the symbol ``$\,\proofbox\,$'' at the end of any proof, even if it ends in an
equation environment. \emph{Note that the document may need to be
  compiled twice for the mark to appear.}
Some of the calculus examples were adapted from \cite{CalcI}. 
%
% An example theorem and lemma declaration is shown in \cref{ex:theorem}.  These
% are all numbered in the same sequence and produce labels in small caps
% with an italic body. Use the command \code{\textup} to produce Roman
% type for numbers and parentheses.
%
\Cref{ex:theorem} shows usage of the \code{theorem} environment. 
An optional argument can be used to name the theorem.
\Cref{ex:cor} illustrates a corollary, without a name, and the
proof environment.

~

\begin{example}[label=ex:theorem,bicolor,parbox=false,%
listing options={style=siamlatex,%
{morekeywords=[2]{theorem}}}]{Theorem}
\begin{theorem}[Mean Value Theorem]\label{thm:mvt}
  Suppose $f$ is a function that is continuous on the closed interval
  $[a,b]$.  and differentiable on the open interval $(a,b)$.
  Then there exists a number $c$ such that $a < c < b$ and
  \begin{displaymath}
    f'(c) = \frac{f(b)-f(a)}{b-a}.
  \end{displaymath}
  In other words, $f(b)-f(a) = f'(c)(b-a)$.
\end{theorem}
\end{example}

\begin{example}[label=ex:cor,bicolor,parbox=false,%
listing options={style=siamlatex,%
{morekeywords=[2]{corollary,proof}}}]%
{Corollary and proof}
\begin{corollary}
  Let $f(x)$ be continuous and differentiable everywhere. If $f(x)$
  has at least two roots, then $f'(x)$ must have at least one root.
\end{corollary}
\begin{proof}
  Let $a$ and $b$ be two distinct roots of $f$.
  By \cref{thm:mvt}, there exists a number $c$ such that
  \begin{displaymath}
    f'(c) = \frac{f(b)-f(a)}{b-a} = \frac{0-0}{b-a} = 0.
  \end{displaymath}
\end{proof}
\end{example}

SIAM also defines commands to create your own theorem- and
remark-like environments:
\begin{itemize}
\item \code{newsiamthm} --- Small caps header, italized body. 
\item \code{newsiamremark} --- Italics header, roman body.
\end{itemize}
Each command takes two arguments. The first is the environment name,
and the second is the name to show in the document. These commands should
be used instead of \code{\newtheorem}.
\Cref{ex:claim,ex:ref} shows how to use the commands above, including how to
specify the plural version for \texttt{cleveref} if it is unusual.

\begin{example}[label=ex:claim,bicolor,%
before upper={\preamble{\bs newsiamthm\{claim\}\{Claim\}}\\
\noindent\preamble{\bs newsiamremark\{hypothesis\}\{Hypothesis\}}\\
\noindent\preamble{\bs crefname\{hypothesis\}\{Hypothesis\}\{Hypotheses\}}},%
parbox=false,%
listing options={style=siamlatex,%
{morekeywords=[2]{claim,proof,hypothesis}}}]{New theorem-like environment}
\begin{claim}\label{cl:constant}
  If $f'(x) = 0$ for all $x \in (a,b)$ then $f(x)$ is constant on $(a,b)$.
\end{claim}
\begin{hypothesis}\label{hyp1}
The function $f$ is continuously differentiable.
\end{hypothesis}
\begin{hypothesis}\label{hyp2}
The random variable is normally distributed.
\end{hypothesis}
\end{example}

% \begin{example}[label=ex:remark,bicolor,%
% before upper={\preamble{\bs newsiamremark\{expl\}\{Example\}}},%
% parbox=false,%
% listing options={style=siamlatex,%
% {morekeywords=[2]{expl}}}]{New remark-like environment}
% \begin{expl}[Trivial note]\label{ex:a}
%   Let $f(x) = 2$. Since $f'(x) = 0$ for all $x$, $f$ is constant
%   everywhere.
% \end{expl}
% \end{example}

\begin{example}[label=ex:ref,bicolor,listing options={style=siamlatex,%
{morekeywords=[2]{cref}}}]{References}
We can reference multiple types of objects with a single reference:
\cref{cl:constant,thm:mvt,hyp1,hyp2}.
\end{example}

\section{Tables}
\label{sec:tab}
Table captions should go above the tables.
\Cref{ex:simpletable} shows the code to generate a
\cref{tab:simpletable}.
A more complicated example is shown in 
\cref{ex:table}, which generates \cref{tab:KoMa14}. This
example uses subfloats via the \texttt{subfig} package, as well as
special column options from the \texttt{array} package.

\begin{tcbverbatimwrite}{tmp_\jobname_simpletable.tex}
\begin{table}[tbhp]
  \caption{Example table}
  \label{tab:simpletable}
  \centering
  \begin{tabular}{|c|c|c|} \hline
   Species & \bf Mean & \bf Std.~Dev. \\ \hline
    1 & 3.4 & 1.2 \\
    2 & 5.4 & 0.6 \\ \hline
  \end{tabular}
\end{table}
\end{tcbverbatimwrite}

\examplefile[label={ex:simpletable},%
listing only, listing options={style=siamlatex}]%
{Example table.}{tmp_\jobname_simpletable.tex}

\input{tmp_\jobname_simpletable.tex}

\begin{tcbverbatimwrite}{tmp_\jobname_table.tex}
\newcolumntype{R}{>{$}r<{$}}  %
\newcolumntype{V}[1]{>{[\;}*{#1}{R@{\;\;}}R<{\;]}} %
\begin{table}[tbhp]
 \captionsetup{position=top} %<- Needed for using subtables created with the subfig package 
  \caption{Example table adapted from Kolda and Mayo \rm{\cite{KoMa14}}.}
  \label{tab:KoMa14}
  \centering
  \subfloat[$\beta=1$]{
    \begin{tabular}{|r|R|V{3}|c|r@{\,$\pm$\,}l|} \hline
occ. & \multicolumn{1}{c|}{$\lambda$} & \multicolumn{4}{c|}{$\mathbf{x}$} & 
fevals &  \multicolumn{2}{c|}{time (sec.)}\\ \hline
 718 & 11.3476 &  0.5544 &  0.3155 &  1.2018 &  0.0977 &   45 & 0.17 & 0.06 \\ \hline 
 134 &  3.7394 &  0.2642 & -1.1056 &  0.2657 & -0.3160 &   31 & 0.12 & 0.05 \\ \hline 
   4 & \multicolumn{6}{c|}{\emph{--- Failed to converge ---}} & 0.21 & 0.10 \\ \hline 
 \end{tabular}}

  \subfloat[$\beta=-1$]{
    \begin{tabular}{|r|R|V{3}|c|r@{\,$\pm$\,}l|} \hline
occ. & \multicolumn{1}{c|}{$\lambda$} & \multicolumn{4}{c|}{$\mathbf{x}$} & 
fevals &  \multicolumn{2}{c|}{time (sec.)}\\ \hline
  72 & -1.1507 &  0.2291 &  0.6444 &  0.3540 & -0.8990 &   34 & 0.14 & 0.06 \\ \hline 
 624 & -6.3985 &  0.1003 &  0.1840 &  0.5305 &  1.2438 &   48 & 0.19 & 0.08 \\ \hline 
   2 & \multicolumn{6}{c|}{\emph{--- Failed to converge ---}} & 0.23 & 0.02 \\ \hline 
 \end{tabular}}
\end{table}
\end{tcbverbatimwrite}

\examplefile[label={ex:table},%
before upper={\preamble[\scriptsize]{\bs usepackage\{array\}}\\[-0.4em]
\noindent\preamble[\scriptsize]{\bs usepackage[caption=false]\{subfig\}}},%
listing only, listing options={%
  style=siamlatex,basicstyle=\ttfamily\scriptsize}]%
{Example table with subtables.}{tmp_\jobname_table.tex}

\input{tmp_\jobname_table.tex}

\section{Figures}
\label{sec:fig}

It is recommended that all figures be generated in high resolution.
In the past, SIAM has required encapsulated
postscript (EPS) format for final production. This is still an
acceptable format, but SIAM also now allows high-resolution 
PDF, JPEG, and PNG figures. 
%
If working with EPS images and using \texttt{pdflatex}, we
recommend the package \texttt{epstopdf} to automatically convert EPS
images to PDF for inclusion in PDF documents created by
\texttt{pdflatex}.
%
\Cref{ex:fig} shows the code to generate \cref{fig:testfig}.  This
example uses the \texttt{graphicx} package for the
\code{\includegraphics} command.

\begin{tcbverbatimwrite}{tmp_\jobname_fig.tex}
\begin{figure}[tbhp]
  \centering
  \subfloat[$\epsilon_{\max}=5$]{\label{fig:a}\includegraphics{lexample_fig1}}
  \subfloat[$\epsilon_{\max}=0.5$]{\label{fig:b}\includegraphics{lexample_fig2}}
  \caption{Example figure using external image files.}
  \label{fig:testfig}
\end{figure}
\end{tcbverbatimwrite}

\examplefile[label={ex:fig},%
before upper={\preamble[\scriptsize]{\bs usepackage\{graphicx,epstopdf\}}\\[-0.4em]
\noindent\preamble[\scriptsize]{\bs usepackage[caption=false]\{subfig\}}},%
listing only, listing options={%
  style=siamlatex,basicstyle=\ttfamily\scriptsize}]%
{Example figure with subfigures and external files}{tmp_\jobname_fig.tex}

\input{tmp_\jobname_fig.tex}

%\newpage
Another option for figures is a graphics-generator that is platform- and
format-independent.
PGF is a TeX macro package for generating  such graphics and works together with the most important TeX backend drivers, including pdftex and dvips. 
The user-friedly syntax layer called TikZ.
Here we show an example using \texttt{PGFPLOTS}, useful for drawing high-quality 
plots directly in \LaTeX.
\Cref{ex:data} and \cref{ex:pgfplots} shows the data and code, respectively, to generate \cref{fig:pgfplots}, adapted from \cite{pgfplots}.

\examplefile[label={ex:data},listing only,
listing options={style=siamlatex,basicstyle=\ttfamily\scriptsize}]%
{Example data file (data.dat)}{data.dat}

\begin{tcbverbatimwrite}{tmp_\jobname_tikz.tex}
\begin{figure}[tbhp]
  \centering
  \begin{tikzpicture}
    \begin{loglogaxis}[height=2.75in, grid=major,
      xlabel={Degrees of Freedom}, ylabel={$L_2$ Error},
      legend entries={$d=2$,$d=3$}]
      \addplot table [x=d2_dof,y=d2_l2_err] {data.dat};
      \addplot table [x=d3_dof,y=d3_l2_err] {data.dat};
    \end{loglogaxis}
  \end{tikzpicture}
  \caption{Example \texttt{PGFPLOTS} figure.}
  \label{fig:pgfplots}
\end{figure}
\end{tcbverbatimwrite}

\examplefile[label={ex:pgfplots},%
before upper={\preamble[\scriptsize]{\bs usepackage\{pgfplots\}}},%
listing only, listing options={%
  style=siamlatex}]%
{Example TikZ/PGF for platform-independent graphics.}{tmp_\jobname_tikz.tex}

\input{tmp_\jobname_tikz.tex}

%\newpage  
\section{Algorithms}
\label{sec:algs}

SIAM automatically includes the \texttt{algorithm} package in the
class definition. This provides the float environment.
Users have the choice of \texttt{algpseudocode},
\texttt{algorithmic}, and other packages for actually formatting the algorithm.
For example, \cref{alg:buildtree} is produced by the code in \cref{ex:alg}.
In order to reference lines within the algorithm, we need to tell the
\texttt{cleveref} package how to do the referencing, which is the
second line of \cref{ex:alg}. Then we can use the code
\code{\cref{line3}} to produce \cref{line3}.

\begin{tcbverbatimwrite}{tmp_\jobname_alg.tex}
\begin{algorithm}
\caption{Build tree}
\label{alg:buildtree}
\begin{algorithmic}[1]
\STATE{Define $P:=T:=\{ \{1\},\ldots,\{d\}$\}}
\WHILE{$\#P > 1$}
\STATE\label{line3}{Choose $C^\prime\in\mathcal{C}_p(P)$ with $C^\prime := \operatorname{argmin}_{C\in\mathcal{C}_p(P)} \varrho(C)$}
\STATE{Find an optimal partition tree $T_{C^\prime}$ }
\STATE{Update $P := (P{\setminus} C^\prime) \cup \{ \bigcup_{t\in C^\prime} t \}$}
\STATE{Update $T := T \cup \{ \bigcup_{t\in\tau} t : \tau\in T_{C^\prime}{\setminus} \mathcal{L}(T_{C^\prime})\}$}
\ENDWHILE
\RETURN $T$
\end{algorithmic}
\end{algorithm}
\end{tcbverbatimwrite}

\examplefile[float=htpb,label={ex:alg},%
before upper={\preamble[\scriptsize]{\bs usepackage\{algorithmic\}}\\[-0.4em]
\preamble[\scriptsize]{\bs Crefname\{ALC@unique\}\{Line\}\{Lines\}}},%
listing only, listing options={%
  style=siamlatex,basicstyle=\ttfamily\scriptsize}]%
{Example algorithm}{tmp_\jobname_alg.tex}

\input{tmp_\jobname_alg.tex}

\section{Sections}
\label{sec:sec}
Sections are denoted using standard \LaTeX\ section commands, i.e.,
\code{\section}, \code{\subsection}, etc. 
%
If you wish to end the section title with something other that a period (the default), you have to add the command \code{\nopunct} at the end of the title.

Appendices are created with the normal sectioning commands, following
the command \code{\appendix}. 
%
Titles of appendices created with
\code{\section} are preceded by the word ``Appendix,'' but not the
subsections or appendices created with \code{\section*}.
%
Unlike normal sections, appendix sections may be sensitive to blank lines
following the declaration, causing a new paragraph rather than the
text immediately following the appendix title. This can be corrected
by removing and blank lines.
%
Any numbered, labeled sections can be referenced using \code{\cref},
including those without a title.
%
Section titles are automatically inserted into the table of contents
and converted to bookmarks; see \cref{sec:spec-instr-pdf} for handling
special characters.

The acknowledgments section comes
immediately before the references and after any appendices. It should
be declared by \code{\section*{Acknowledgments}}.

%
% \Cref{ex:secref} shows how to reference sections.

% \begin{example}[label={ex:secref},lefthand ratio=0.4,bicolor,sidebyside,%
% listing options={style=siamlatex,basicstyle=\ttfamily\scriptsize,%
% deletetexcs={cref,Cref},{moretexcs=[2]{cref,Cref}}}]
% {Right and wrong ways to reference a section}
% Inside a sentence\dots\\
% Single: \cref{sec:intro}\\
% Range: \cref{sec:intro,sec:front,%
% sec:sec}\\
% Multiple: \cref{sec:intro,sec:sec,%
% sec:tab,sec:math,sec:thm}\\
% Appendix: \cref{sec:changes}\\

% Beginning of a sentence\dots\\
% Single: \Cref{sec:intro}\\
% Range: \Cref{sec:intro,sec:front,%
% sec:sec}\\
% Multiple: \Cref{sec:intro,sec:sec,%
% sec:tab,sec:math,sec:thm}\\
% Appendix: \Cref{sec:changes}\\

% Just don't do it this way\dots\\
% Section~\ref{sec:intro}  
% \end{example}

\section{Supplemental material}
\label{sec:supplement}

For several SIAM journals, authors are encouraged to submit
Supplementary Materials to complement their articles. This might
include additional figures or examples, animations, data sets used in
the paper, computer code used to generate figures or tables, or other
materials that are necessary to fully document the research contained
in the paper or to facilitate the readers' ability to understand and
extend the work.

\newpage
The class option \code{supplement} should be used in the
supplemental \LaTeX\ file provided for creating PDF supplemental material.
The supplement should have the same title and authors as the main
document.  The title is modified automatically by the SIAM
class file so that it is preceded by the text ``Supplementary
Materials'', followed by a colon.  
%
The numbering is modified so that
all sections, equations, figures, tables, algorithms, %
and so on to start with ``SM''. 
%
A supplement does have sections but does not have an
abstract, keywords, AMS classifications, or appendices.  The main
document and supplement can cross reference sections, equations,
theorem-like declarations, figures, tables, algorithms, etc.  However,
there is no sharing of references. The references are optional for a
supplement.
A template is provide, as discussed in \cref{sec:template}.

\section{Template}
\label{sec:template}

The files \texttt{ex\_article.tex}, \texttt{ex\_shared.tex},
and \texttt{ex\_supplement} \texttt{.tex} provide a template that can be used
for creating a \LaTeX\ document with an optional supplement.
%
\Cref{ex:artoutline,ex:suppoutline} give the outline of an article and
the supplement. In this case we
assume that the title and authors are defined in the
\texttt{ex\_shared.tex} file, shown in \cref{ex:shared}.
%
Cross referencing between the main document and the supplement is
enabled using the \texttt{xr-hyperref} package (included by the
SIAM class file). Use \code{\externaldocument}
to specify the external document to search for external
references.
%

\begin{example}[label={ex:artoutline},listing only,%
  listing options={style=siamlatex,%
    {morekeywords=[2]{externaldocument}}}]%
  {Document outline with supplement}
\documentclass{siamart1116}
\input{ex_shared}
\externaldocument{ex_supplement}
\begin{document}
\maketitle
% Other front matter: abstract, keywords, subject classifications.
% Main body goes here.
% Appendices and/or acknowledgments.
% Bibliography
\end{document}
\end{example}

\begin{example}[label={ex:suppoutline},listing only,%
  listing options={style=siamlatex,%
    {morekeywords=[2]{supplement,siampretitle,externaldocument}}}]%
  {Supplement document outline}
\documentclass[supplement]{siamart1116}
\input{ex_shared}
\externaldocument{ex_article}
\begin{document}
\maketitle
% No abstract, keywords, subject classifications.
% Main body goes here.
% No appendices or acknowledgments.
% Optional bibliography goes here.
\end{document}
\end{example}

%\newpage
\Cref{ex:shared} (from \texttt{ex\_shared.tex}) shows how the
``shared'' title and authors may be defined across the main document
and a supplement. 
%
Note the use of the \code{\string} command in the URL for the tilde;
this is only necessary inside the \code{\thanks} command.
%

\examplefile[label=ex:shared,listing only,%
listing options={style=siamlatex,firstline=19,lastline=40,%
{morekeywords=[2]{siampretitle,MakeUppercase,string}}}
]{Example of shared title and author macros}{ex_shared.tex}

\section{Bibliography}
\label{sec:bib}

The SIAM \BibTeX style file, now called \texttt{siamplain.bst}, has been updated to include the new keys listed below:
\begin{itemize}
\item \code{doi}: Digital object identifier, a unique alphanumeric string
\item \code{url}: Web address, usually impermanent
\item \code{urldate}: Date that the web address was last accessed
\item \code{eprint}: Archive identifier, a unique alphanumeric string
\item \code{eprintclass}: Archive class
\item \code{archive}: Archive URL, defaults to \nolinkurl{https://arXiv.org/abs}
\item \code{archivepreprint}: Archive name, defaults to ``arXiv''.
\item \code{eid}: Article ID, if there are no page numbers
\item \code{pagetotal}: Total number of pages, for use with article ID
\end{itemize}
Every entry type has been modified to include an optional link to a
DOI, a URL, and/or an archive preprint reference.
%
Additionally, the \code{article} entry now supports an Article ID,
\code{eid}, and number of pages, \code{pagetotal}.
%
To use this, include the following code in your \LaTeX\ source code:
\code{\bibliographystyle{siamplain}}.

\subsection{DOI}

A digital object identifier (DOI) is a unique alphanumeric string that provides a persistent link to its location on the Internet. The publisher assigns a DOI when your article is published and made available electronically. Using the \code{doi} field in \BibTeX to specify it, as shown for \cite{KoMa14} in \cref{ex:article}; observe the new \code{doi} field which produces a hyperlink in the citation.
Do not include the full URL, i.e., \nolinkurl{https://doi.org/} preceding the DOI.
Authors are highly advised to provide DOIs; if they are not provided, they
will be requested during the copyediting process.

\examplefile[label=ex:article,listing only,%
listing options={style=siamlatex,firstline=1,lastline=12,{morekeywords=[2]{doi}}}
]{Example article in \BibTeX}{references.bib}

\subsection{URL}
There is also now support for the  \code{url} field. Generally, the DOI is preferred to the URL, since the DOIs should be permanent references. For that reason, it is good practice to specify the last date that the URL was accessed, which is specified by the optional \code{urldate} field. Reference \cite{Hi14} produced by \cref{ex:url} shows an example of using these fields.

\examplefile[label=ex:url,listing only,%
listing options={style=siamlatex,firstline=20,lastline=28,%
{morekeywords=[2]{url,urldate}}}
]{Example with the URL field in \BibTeX}{references.bib}

\subsection{Preprint servers such as arXiv}

More and more manuscripts are available on preprint servers. In fact,
SIAM's publication policy explicitly allows the final accepted version
of any article to be posted on a preprint server such as arXiv.

For an arXiv paper,
the \code{eprint} field is used to specify the identifier. The optional \code{eprintclass} field specifies the class. \Cref{ex:arxiv} shows the \BibTeX for \cite{PeKoPi14}.

\examplefile[label=ex:arxiv,listing only,%
listing options={style=siamlatex,firstline=30,lastline=38,%
{morekeywords=[2]{eprint,eprintclass}}}
]{Example arXiv reference in \BibTeX}{references.bib}

\newpage
Other preprint servers are supported as well, but these require
specification of the fields \code{archive} and \code{archiveprefix}. In this case, the target URL is formed by concatenating the \code{archive}, a forward slash (\texttt{/}), and the \code{eprint}; and the text for the hyperlink is formed by concatenating the \code{archiveprevix}, a colon (:), and the \code{eprint}.
%
\Cref{ex:pubmed} shows the code to generate \cite{WoZhMeSh05}, including the preprint from PubMed.
Note that this example has both the journal citation as well as the link for the preprint.

\examplefile[label=ex:pubmed,listing only,%
listing options={style=siamlatex,firstline=40,lastline=55,%
{morekeywords=[2]{eprint,eprintclass,archive,archiveprefix}}}
]{Example PubMed reference in \BibTeX}{references.bib}

\subsection{Article ID}
Some journals use an article ID rather than page numbers. 
The field \texttt{eid} specifies the article ID. The optional field \texttt{pagetotal} can say the number
of pages in the document. An example of an article using these fields is shown in \cref{ex:eid} for citation \cite{Ne03}.

\examplefile[label=ex:eid,listing only,%
listing options={style=siamlatex,firstline=57,lastline=67,%
{morekeywords=[2]{eid,pagetotal}}}
]{Example article ID reference in \BibTeX}{references.bib}

\subsection{Software citations}

SIAM encourages software citations, both related technical publications as well as the software itself. A citation to a software package may look something like what is shown in \cref{ex:clawpack} for citation \cite{clawpack}. Notice the double braces around the \texttt{author} key; else, it would appear as ``C. D. Team''.

\examplefile[label=ex:clawpack,listing only,%
listing options={style=siamlatex,firstline=69,lastline=76,%
{morekeywords=[2]{key}}}
]{Example software reference in \BibTeX}{references.bib}

\appendix
\section{Summary of major changes}
\label{sec:changes}

Here we summarize the major changes in the latest version of
the SIAM standard \LaTeX\@ and \BibTeX\ classes:
\begin{itemize}
\item Change in file names: \texttt{siamltex1213.cls} is replaced by
  \texttt{siamart1116.cls}, and the \texttt{siam.bst} bibliography style
  file is replaced by \texttt{siamplain.bst}; see \cref{sec:intro}.
\item Hyperlinking in cross references via the
  \texttt{cleveref} package, including customizations to adhere to
  SIAM conventions. Automatic PDF bookmarks enabled for sections,
  appendices, and references.
  See \cref{sec:cleveref}.
\item Colored hyperlinks (red for external, green for internal). Can
  be disabled with class options \code{hidelinks}; see \cref{sec:hyperef}.
\item New fields for \BibTeX, as listed in \cref{sec:bib}.
\item  Support for supplemental PDF files, including
  cross references between the supplement and the main document; see
  \cref{sec:supplement}. 
\item New command for setting headers: \code{\headings}; see
  \cref{sec:front}. 
\item Updated theorem-like and proof environments using
  \texttt{ntheorem} package; see \cref{sec:thm}.
\item Fixed handling of appendices to adhere to SIAM style guidelines;
  see \cref{sec:sec}.
\item Added \code{review} class option; see \cref{sec:class-options}.
\item Includes the following packages by default:
  \texttt{algorithm}
  \texttt{amsmath}, 
  \texttt{breakurl}, 
  \texttt{cleveref}, 
  \texttt{hypcap}, 
  \texttt{hyperef},
  \texttt{ifpdf},
  \texttt{ntheorem}, 
  \texttt{xcolor}, and
  \texttt{xr-hyperef}.
  Adds the \texttt{lineno}
  package if the \code{review} class option is enabled.
\end{itemize}

\section{Special modifications}

\subsection{Special macros}
In past versions, 
the SIAM standard \LaTeX\ class defined the following macros:
\code{\const},
\code{\diag},
\code{\grad},
\code{\Range},
\code{\rank}, and
\code{\supp}.
These are no longer declared; however, they can be redefined using the
\code{\DeclareMathOperator} command from the \texttt{amsopn}
package as demonstrated in \cref{ex:dmo}.

\subsection{Labeling objects according to the section number}
SIAM recommends numbering  objects sequentially. 
However, if you prefer to number objects by section, e.g., Figure 5.1
would indicate the first figure in Section 5, put the code from
\cref{ex:numbering} into the document preamble.

\begin{example}[label=ex:numbering,listing only,%
listing options={style=siamlatex,{morekeywords=[2]{makeatletter,makeatother}},}]%
{Labeling objects by section}
\renewcommand{\theequation}{\thesection.\arabic{equation}} 
\renewcommand{\thefigure}{\thesection.\arabic{figure}}
\renewcommand{\thetable}{\thesection.\arabic{table}}
\renewcommand{\thealgorithm}{\thesection.\arabic{algorithm}}
\end{example}

\subsection{Appendices with no title}
The SIAM style manual \cite{siam} allows for an appendix that is
numbered (by a letter) but has no title. We have a special command to
create such an appendix: \code{\appendixnotitle}. This is equivalent
to a \code{\section} command in the appendix except that it has no
arguments.

\subsection{Changing the font of algorithm titles}
\label{sec:chang-font-algor}

Currently, the algorithm title font matches the figure and table title
fonts. To make the algorithm title font appear as small caps, insert
the code in \cref{ex:algfont} into the preamble.

\begin{example}[label=ex:algfont,listing only,%
listing options={style=siamlatex}]{Changing algorithm title font}
\makeatletter
\renewcommand{\ALG@name}{\sc Algorithm}
\makeatother
\end{example}

\subsection{Changing title of proof}
To change the title of a proof to, say, ``Proof of main theorem,''
simply use an optional argument as shown in
\cref{ex:newproof}.

\begin{example}[label=ex:newproof,listing only,%
listing options={style=siamlatex}]{Altered proof environment}
\begin{proof}[Proof of main theorem]
We now show ...
\end{proof}
\end{example}

\subsection{Special instructions for PDF bookmarks}
\label{sec:spec-instr-pdf}

Section titles are automatically inserted into the table of contents
and therefore used as PDF bookmarks. Some special symbols may not
correctly render. In that case, you can define alternate text as shown
in \cref{ex:specialtitle}, using the \code{\texorpdfstring} command
provided by the \texttt{hypertex} package. In this example, the
default bookmark would be ``Discussion of Z = X Y''; instead, it is
replaced by ``Discussion of Z = X union Y.''

\begin{example}[label=ex:specialtitle,listing only,%
listing options={style=siamlatex}]{PDF bookmarks for symbols in section titles}
\section{Discussion of \texorpdfstring{{\boldmath$Z=X \cup Y$}}{Z = X union Y}}
\end{example}

\bibliographystyle{siamplain}
\bibliography{references}

\end{document}

### 2017-12-11 doc/articles/amsi_overview/amsi_overview.tex
working on the amsi paper

@@
-\title{The Adaptive Multiscale Simulation Infrastructure -- Core Features and Biological Tissue Simulation Feature Utilization}
-\author{W.R. Tobin, D. Fovargue, V.W.L. Chan, M.S. Shephard}
+\title{The Adaptive Multiscale Simulation Infrastructure -- Core Features and Multi-scale Soft Tissue Simulation}
+\author{W.R. Tobin, V.W.L. Chan, M.S. Shephard}
 \date{\today}
@@
 \label{amsi_scales}
-Individual scales in a simulation are declared during the initialization of the simulation, follow by the determination and asignment of RankSets to scales, guided by a supplied configuration file or user-implemented algorithm.
+Individual scales in a simulation are declared during the initialization of the simulation, follow by the determination and assignment of RankSets to scales, guided by a supplied configuration file or user-implemented algorithm.
 Scale-coupling relations are also declared at initialization time, though this decleration is limited to the indication that two scale may interact at some point during the simulation, and nothing more.

### 2018-01-05 doc/articles/amsi/amsi.tex

@@
 
-\title{The Adaptive Multiscale Simulation Infrastructure - Methods and Soft-tissue Example Simulation}
-\author{W.R. Tobin, V.W.L. Chan, M.S. Shephard}
-\date{\today}
 \documentclass[review]{siamart1116}
 
+\newcommand{\blue}[1]{\textcolor{blue}{[#1]}}
+\newcommand{\red}[1]{\textcolor{red}{[#1]}}
+\newcommand{\green}[1]{\textcolor{green}{[#1]}}
+
 \begin{document}
+
+\title{The Adaptive Multiscale Simulation Infrastructure - Methods and Soft-tissue Example Simulation}
+
+% Authors: full names plus addresses.
+\author{
+  William R. Tobin \thanks{Scientific Computational Research Center, Rensselaer Polytechnic Institute, Low Center for Industrial Innovation, CII-4011, 110 8th Street, Troy, NY 12180.} 
+  \and
+  Victor W.L. Chan\footnotemark[1]
+  \and
+  Catalin R.
+Picu\footnotemark[1] \thanks{Department of Mechanical, Aerospace and Nuclear Engineering, Rensselaer Polytechnic Institute, Jonsson Engineering Center, Rm.\ 2049, 110 8th Street, Troy, NY 12180.}
+  \and
+  Victor H. Barocas\thanks{Department of Biomedical Engineering, University of Minnesota, 7-105 Nils Hasselmo Hall, 312 Church Street SE, Minneapolis, MN 55455.}
+  \and
+  Mark S. Shephard\footnotemark[1].
+}
+\date{\today}
 \maketitle
@@
 \begin{abstract}
-Discuss multi-scale and frameworks, show how AMSI does multi-scale, and how it is different than others.
+\red{Discuss multi-scale and frameworks, show how AMSI does multi-scale, and how it is different than others.
 Introduce the soft-tissue problem.
 Discuss hierarchical multi-scale and relate it to well-known multi-scale models, and what requirements it imposes.
-Discuss how AMSI works, and how to use AMSI to address the requirements of a hierarchical multi-scale problem.
-Show how AMSI was used to construct the soft-tissue problem.
-Show results somehow...
+Discuss AMSI design decisions, the implementation of AMSI, and how to use AMSI to implement the soft tissue simulation.
+Show some performance measures relevant to the implementation of this sort of simulation, and finally produce a roadmap on how to apply AMSI to other simulations (hierarchical and concurrent).}
 
 \end{abstract}
-
-Targeting the SIAM Journal On Scientific Computing
 
@@
 
-Boilerplate about why multi-scale simulation is required, what issues it attempts to confront.
+The development of new numerical simulations has proliferated rapidly in the last several decades.
+The massive increases in computational power of massively parallel machines has facilitated the implementation and execution of numerical simulations of unprecedented scale and fidelity.
+However, when it is the case that the governing physical mechanism of interest in a simulation occurs at a scale orders removed from the desired resolution of the solution, even the massive increases in computational power of the last decades is insufficient in many cases.
+Multi-scale simulation techniques offer a method of achieving accurate results on a domain orders of magnitude removed from the underlying phenomena guiding the physical action of the simulation.
 
-Introduce concept that multi-scale is difficult from an implemenetation and computational efficiency standpoint, while there has been a proliferation in the development of actual physical coupling models through largely ad-hoc implementations.
+Multi-scale simulation broadly falls into two categories: concurrent (partitioned-domain) and hierarchical problems.
+These categories are most readily distinguished by the relationships between the interacting scales in a problem.
 
-Briefly discuss hierarchical multi-scale vs concurrent multi-scale, mention that AMSI is currently focused on providing support for hierarchical multi-scale problems, though leaving development paths open for concurrent.
+In concurrent problems the domain of each scale is modeled using similar physical dimensions: the standard units used to describe the scale will typically fall within several orders of magnitude of each other.
+Interface surfaces or regions between the scale domains are used to transfer relevant physical tensor field information between scales.
+The transfer of relevant tensor field values allow the scales to influence one another and gives rise to multi-scale behaviors in the problem simulation.
+Often values in concurrent simulations will not require the use of up- or down-scaling operations to transfer tensor values.
 
-Discuss 2-3 other multi-scale frameworks (UINTAH, MPCCI, one more..?)
+The Domain Decomposition Methods (DDM) \cite{toselli2005domain} and the Quasicontinuum Methods (QM) \cite{miller2002quasicontinuum} both allow the implementation of concurrent multi-scale problems.
+DDM allows the implementation of multi-physics continuum-to-continuum coupling while QM is used to couple atomistic models to continuum regions in a macro-scale domain.
 
-\subsection{AMSI}\label{sec:amsi}
+%Most continuum/continuum coupling problems fall into this category when the scale separation is sufficiently small.
+This class of problem also includes atomistic/continuum couplings \cite{miller2009unified}, in which the scales may often be separated by more than several orders of magnitude.
 
-Introduce AMSI, show how it intends to adress the issues of difficult implementation as a top priority while attempting to maintain computational efficiency as well.
+Hierarchical problems are characterized by strongly-separated scales in which the physical dimensions of the scale are typically 6 or more orders of magnitude separated.
+The domain relationship between two such scales is often reduced to the existence of micro-scale simulations occurring at various points throughout the larger-scale simulation.
+The tensor field couplings which give rise to multi-scale behavior occur at each point in the macro-scale simulation at which a micro-scale simulation is embedded.
+The local value of tensor field data is used to establish conditions at the micro-scale, which in turn produces a tensor field value relevant to the evolution of the macro-scale simulation.
+Values in hierarchical simulations are more likely to require up- and down-scaling operations to account for  differences between the coupled scales.
 
-Show how the AMSI approach is different from the other multi-scale frameworks discussed above.
+Hierarchical problems can be implemented with the use of information passing schemes in which a well-defined set of parameters is transmitted between the simulation scales \cite{weinan2011principles,buehler2006concurrent,fish2010multiscale}.
+This approach allows the replacement of a constitutive model commonly used in continuum mechanics problems with a full single-scale simulation conducted at a distinct points in the macro-scale problem.
 
-\subsection{Soft-Tissue Problem}\label{sec:bio}
+The Heterogeneous Multiscale Methods (HMM) \cite{weinan2003heterogenous,weinan2007heterogeneous} are methods used to solve multi-scale problems with either localized regions of macroscopic model failure or problems without a known macroscopic model entirely.
+HMM is well-suited for application to hierarchical problems as it does not require the existence of a macro-scale constitutive model, allowing the micro-scale simulations to fully guide system behavior.
 
-Brief overview of the multi-scale structure of AMSI with reference back to papers covering the physics and some results particular to the physics in greater detail (the neuron paper).
+%The Soft Biological Tissue problem discussed in section \ref{sec:bio-implementation} is an instance of this case of an HMM problem.
+%Continuum/continuum problems with strongly-separated scales fall into the hierarchical category,
 
-\section{Methods}\label{sec:methods}
+An example of scale separation maps \cite{borgdorff2013foundations} associated with these multi-scale problem categories can be seen in figure \ref{fig:scale-map}.
+Further discussion of the taxonomy of multi-scale modeling in \cite{tadmor2011modeling} is a useful primer on the various categories of multi-scale models, with accompanying examples for each category to further elucidate the concepts.
 
-Maybe add a section before \ref{sec:hierarchic} taking the most generic parts of \ref{sec:amsi-construction} and discussing them first, before moving into hierarchical in particular?
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\columnwidth]{scale-separation-map}
+  \caption{Scale separation maps for hierarchical and concurrent multi-scale problems, respectively.}
+  \label{fig:scale-map}
+\end{figure}
 
-\subsection{General Hierarchical Multi-scale}\label{sec:hierarchic}
+In the rest of this paper we will discuss currently existing multi-scale frameworks including particulars of their implementations and use cases.
+We will then discuss the Adaptive Multi-scale Simulation Infrastructure (AMSI) tools we have developed including our intent in development and targeted problem types, the design of our system, and the subsequent implementation of that design.
+We will further discuss the application of our infrastructure to the development of a multi-scale soft biological tissue simulation.
+We then discuss the performance of the soft tissue simulation and how this relates to the AMSI multi-scale coupling approach.
+Finally we will discuss our plans for future development and extension of the capabilities of our infrastructure.
 
-Dig a little deeper into hierarchical multi-scale, discuss the computational aspects of this type of multi-scale, and reference some well-known multi-scale models that fall into this category.
+\section{Multi-scale Frameworks}\label{sec:frameworks}
 
-\subsection{AMSI Simulation Construction}\label{sec:amsi-construction}
+There are numerous multi-scale frameworks currently in development intended to facilitate the implementation of multi-scale simulations on HPC machines.
+We do not intend to provide a full survey of such frameworks in this section but rather to discuss several frameworks which exhibit features common to many frameworks existing in the multi-scale space at this moment.
 
-Discuss how AMSI operates in general, then go in to how to use AMSI to address the needs of a hierarchical multi-scale problem.
+Software intended to support multi-scale simulations needs to ”enable the introduction of new models, algorithms, and data structures \cite{keyes2013multiphysics}.
+Further, a set of well-defined modules with distinct interfaces must be implemented, especially as this concerns data ownership and simulation state.
+State data must be query-able through interfaces in order for coupled codes to maintain consistency throughout the simulation.
+Further, implementing general code, or providing mechanisms to extract important underlying data in general formats is necessary.
+This forces multi-scale application developers to adhere to specific data structures and algorithms, or write their own wrapper code in order to translate specialized data structures into more general formats.
+These issues and rationales for multi-scale software development are further discussed in\cite{keyes2013multiphysics}.
 
-\subsection{Soft-Tissue Implementation}\label{sec:bio-construction}
+The Uintah Problem Solving Environment (PSE) \cite{berzins2010uintah} is a framework developed at the University of Utah intended to allow the intuitive combination of parallel simulation components into novel scientific applications, including but not limited to multi-scale simulations.
+Each component in the Uintah system is designed to solve a PDE on a structured AMR grid, and must adhere to a C++ API used to establish connections between components.
+All components are decomposed into a set of discrete tasks, each of which has a well-defined set of required input and output data.
+The information about these tasks is expressed in the form of a directed acyclic graph (DAG) \cite{berzins2015extending}, and is used by the Uintah system to schedule task execution across the HPC system.
+The parallel communication necessary to provide task input and output occurs through a data warehousing structure which controls access to local and global data.
+
+The Mesh-Based Parallel Code Coupling Interface (MpCCI) \cite{joppich2006mpcci} is a library which facilitates the coupling of simulation codes based on the underlying domain grid over which simulation data (specifically tensor field values) are distributed.
+Support for different grid formats is provided by MpCCI, as well as different methods of data interpolation when needed to communicate data between simulation scales with mismatched grids.
+MpCCI uses a code-coupler component object in order for two single-scale application codes to communicate with eachother \cite{delalondre2010collaborative}.
+A coupler object is implemented to transform values from a producing single-scale code into values that can be used by a consuming single-scale code.
+
+In the time since MpCCI was developed over a decade ago it has seen sufficent adoption to warrant the organization of several user forums and is currently used to support the implementation of multi-physics codes.
+There have also been coupler objects implemented to allow many industry-standard simulation tools to be coupled through MpCCI including Abaqus, ANSYS Fluent, MATLAB, and others \cite{mpcci-website}.
+
+The \verb=MUSCLE-HPC= \cite{belgacem2017muscle} code is a partial implementation of the \verb=MUSCLE-2= toolkit \cite{borgdorff2014distributed} for use on HPC machines.
+The multi-scale simulation in \verb=MUSCLE-HPC= is constructed from a set of submodels, each of which is implemented by inheriting from an abstract C++ class which provides an API for initialization, boundary condition updates, numerical solution iteration, and several data query functions.
+Coupling together of submodels into a multi-scale model is done through the use of an interface of Conduits which the class API has methods to express the set of input and output Conduits for the submodel.
+Further the \verb=MUSCLE-HPC= runtime environment requires that a machine core acting as a server is devoted to running a special Manager class to register client submodels and facilitate the creation of parallel communication channels between the submodels.
+
+%\red{there should be some technical depth on each of the frameworks, a clear indication of the goals of the system design and summary of capabilities}
+
+%\subsection{Multi-scale Methods}\label{sec:multi-scale-methods}
+%\red{Dig a little deeper into multi-scale, discuss the computational aspects of this type of multi-scale, and reference some well-known multi-scale models that fall into this category.}
+
+%As noted in section \ref{sec:multiscale}, hierarchical multi-scale problems are characterized by the strongly-separated nature of the single-scale components interacting in the problem.
+This allows the scales to be coupled with minimal-to-no inter-domain introspection operations.
+Thus these problems can be implemented with the use of information passing schemes in which a well-defined set of parameters is transmitted between the simulation scales \cite{weinan2011principles,buehler2006concurrent,fish2010multiscale}.
+This approach allows the replacement of a constitutive model commonly used in continuum mechanics problems with a full single-scale simulation conducted at a distinct points in the macro-scale problem.
+
+%The Heterogeneous Multiscale Methods (HMM) \cite{weinan2003heterogenous,weinan2007heterogeneous} is a model used to solve multi-scale problems with either localized regions of macroscopic model failure or problems without a known macroscopic model entirely.
+The HMM has a microscopic process acting on physical field \verb-u- and a macroscopic process acting on field \verb+U+, defined over domains \verb-d- and \verb+D+ respectively.
+The physical processes at each scale acting on their respective fields (state variables is the term HMM uses) are related through compression and reconstruction operations used to map values between scales.
+
+%HMM does not require the existence of a macro-scale constitutive model, and instead allows for the possibility that the micro-scale process can directly contribute to the construction of tangent stiffness matrix for the macro-scale problem.
+The Soft Biological Tissue problem discussed in section \ref{sec:bio-implementation} is an instance of this case of an HMM problem.
+
+%%This approach is most often used when a micro-scale simulation is executing on a domain many orders of magnitude spatially separated from the primary macro-scale simulation.
+The micro-scale is sent relevant tensor field values dependent on the location it is embedded in the macro-scale simulation, and produces values used to guide the macro-scale simulation.
+
+%%Information passing is unsuitable for multi-scale simulations where the data shared between scales is heavily coupled to structure and discretization of the problem domain, as is the case when physical fields at foreign scales must be reconciled often (or even rarely) during the lifetime of a simulation.
+For problems with these characteristics, where the spatial scales of the problem are not so strongly separated, the concurrent multi-scale coupling method is more suitable.
+
+%% Hierarchical domain coupling implies the physics to be coupled occur at very strongly separated scales, typically 6 or more orders of magnitude separated.
+Due to this strong separation and the precision of floating point operations it is common to consider the simulation taking place at the finer scale as a point in the larger scale, and couple the simulations together through derived tensor field values at that single point, rather than through a common interface or overlap region.
+
+\section{AMSI}\label{sec:amsi}
+
+%\red{Introduce AMSI, show how it intends to address the issues of difficult implementation as a top priority while attempting to maintain computational efficiency as well.}
+
+The Adaptive Multi-scale Simulation Infrastructure (AMSI) a set of libraries providing support for the implementation and execution of dynamic numerically multi-scale simulations on massively-parallel HPC machines.
+AMSI is designed to ease the incorporation of legacy single-scale simulation codes into new multi-scale simulations.
+This is accomplished by minimizing the insertion of calls to the AMSI systems in a legacy simulation code to only those locations associated with the desired multi-scale coupling.
+
+In contrast to the approach taken in MpCCI, the single-scale codes used in AMSI need not possess a spatial discretization of the problem domain.
+Nor is there any requirement that AMSI directly query the domain data structures.
+Thus AMSI does not require that scale-coupling data be associated with any spatial discretization as is required in \cite{joppich2006mpcci}.
+This protects the capabilities provided by AMSI against a shift in community adoption of new simulation models or different domain representations.
+
+\red{contrast uintah, muscle}
+
+An AMSI user can integrate their codes using AMSI without the need to refactor their data structures or interfaces in order to accommodate the AMSI runtime, as is the case when using multi-scale frameworks designed to operate on a well-specified API the user must expose to the framework.
+AMSI requires only minimal placement of multi-scale coupling calls at those locations in legacy codes associated with the coupling operations used for multi-scale interaction.
+All interfaces to the AMSI system operate on built-in data types or are implemented to accept generic data types, limiting any need to refactor existing codebases to contend with specialized data types and inheritance schemes \cite{plimpton2013developing}.
+
+AMSI is not tied to any particular physics model, domain, or fields implementation or representation, providing operations and data structures to model coupling on an abstract level, while facilitating and supplying less abstract operations (or hooks for such operations) where possible.
+
+There are many open questions in the development and verification of multi-scale modeling approaches \cite{hoekstra2014multiscale}.
+AMSI attempts to provide the tools to approach these questions through the development of new multi-scale models and simulations from existing single-scale simulations without placing assumptions on the multi-scale model or simulation implementation that would preclude exploration of an open question.
+
+\section{Model Design}\label{sec:model-design}
+
+The design of the simulation coupling operations utilizing the AMSI multiscale library is informed by two abstract models.
+The first is the Abstract Scale Model (ASM), which describes a generalized single-scale simulation.
+The second is the Abstract Coupling Model (ACM), which describes a generic coupling of two such simulations.
+These models provide a workflow through which a simulation developer can produce a new multi-scale simulation on an HPC machine using the facilities provided by the AMSI libraries.
+Further, these models guided many development decisions during the implementation of the AMSI multiscale library.
+
+\red{expand:}
+By instantiating these models for a particular problem, a developer will consider all selection decisions required by the proposed multi-scale taxonomy discussed in \cite{gravemeier2008towards}.
+
+ The AMSI model workflow additionally includes considerations for the hierarchy of implementation operations that go into the production of a new multi-scale code.
+
+The design of a Multiscale Modeling Language (MML) is discussed in \cite{hoekstra2007toward}.
+Our approach is compatible with such a language though we do not directly necessitate the use of any specialized languages in the construction of multi-scale simulations with AMSI.
+
+\red{work in citation \cite{ingram2004classification}}
+
+\subsection{Abstract Scale Model}\label{sec:scale-model}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\columnwidth]{scale-model-lines}
+  \caption{The abstract single-scale model used to inform the development of a multi-scale coupling model.}
+  \label{fig:scale-model}
+\end{figure}
+
+The Abstract Scale Model (ASM) (illustrated in figure \ref{fig:scale-model}) provides a hierarchical model of the implementation specifics of a single-scale simulation.
+The model is oriented on the principle of most abstraction to least abstraction along the vertical axis.
+Each level in the abstraction hierarchy requires information in three different areas: the physics being simulated, the domain (type) over which the simulation is to take place, and the physical tensor fields required to describe the problem.
+By specifying or implementing the information required at each node in the ASM a developer will fully describe and produce a single-scale simulation.
+
+The conceptual physical model is often the starting point in the development of an ASM, as it informs many of the decisions made during the specification of the rest of the model.
+This high-level description is often specified in terms of the primary physical property(ies) being modeled, e.g.
+displacements, temperature, trajectories, etc, as well as (typically) one or more conservative laws.
+The specification of the physical model most directly effects the conceptual domain and conceptual tensor fields.
+The conceptual domain incorporates both the temporal and spatial domains of the problem, which provides sufficient information to plot the scale on a scale-separation graph.
+\red{Additionally the conceptual domain includes the specification of e.g.
+continuum vs.
+atomistic approaches.} The conceptual tensor fields are directly influenced by the physical model and includes every primary and supporting field required to simulate the physical model as specified by the chosen conservative laws.
+
+\red{add in some references:}
+The mathematical physical derivation is a specification of the chosen form of the conservative law(s) from the conceptual physical model and accompanying mathematical requirements.
+The mathematical geometric description of the domain incorporates the specification of the method of mathematical representation of a domain for the problem, e.g.
+for a continuum problem a CAD format or grid assumptions.
+The mathematical tensor field structure for the required fields is specified, including symmetries which allow for tensor order reductions to simplify the mathematical formulations in the physical derivation.
+
+The numerical approach includes the approach to solving the mathematical physical derivation this includes the construction and solution of the discretized model.
+The discretized geometry specifies the representation of the geometric domain the numerical approach will query and operate on, this includes for instance meshes and grids for continuum problems and discrete particles for MD problems.
+The tensor field structures and APIs are specified over the discretized geometry and are queried for use in the numerical approach to produce a solution which is also expressed in terms of a tensor field structure.
+
+The computational implementation of an ASM requires the combination of all numerical specifications in order to produce a final simulation.
+
+\subsection{Abstract Coupling Model}\label{sec:coupling-model}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\columnwidth]{coupling-model-lines}
+  \caption{The abstract multi-scale coupling model used to guide the design of AMSI.}
+  \label{fig:coupling-model}
+\end{figure}
+
+The ACM is derived from a model initially discussed in \cite{shephard2010component}.
+
+The Abstract Coupling Model (ACM) (seen in figure \ref{fig:coupling-model}) provides a model of the coupling of two distinct realizations of an ASM (fig \ref{fig:scale-model}).
+The hierarchy of each coupled ASM must be taken into account during the specification of the primary nodes in the ACM.
+It is convenient to view the ACM as a top-down view of two ASMs and the coupling layer between them.
+Each of the conceptual, mathematical, and numerical layers must still be considered when specifying nodes in the ACM.
+Specifying and implementing the ACM fully will result in a valid multi-scale coupling operating between two single-scale simulations.
+
+The physical model compatibility requires the specification of the relationship between the two physical models.
+This includes any laws relating the physical quantities modeled in the two ASMs, and may require intermediate operations including derived values in order for the two physical models to influence each other.
+Proceeding down the ASM hierarchy this further requires a particular mathematical derivation, as well as the specification of any numerical operations required to produce all quantities needed to satisfy the mathematical derivation.
+
+The geometric interaction of problem domains includes the specification of each domain on a scale-separation map, as well as the nature of the domain relationship with respect to e.g.
+overlapping regions/interfaces or full locality of sub-domains for strongly-separated scales.
+This specifies whether the multi-scale problem is concurrent or hierarchic as discussed in sections \ref{sec:multiscale}.
+Going down the hierarchy this also includes any mathematical operations required to map spatially between the domains which can include e.g.
+inverting mappings, interpolation operations, tensor field queries at selected locations.
+At the numerical level the operations necessitated by the mathematical relationships must be implemented.
+
+The field transforms in a coupling require the specification of all operations required for influences from a field associated with a scale to be incorporated into a field on the other scale(s) in the ACM.
+At the conceptual level this includes the full set of tensor fields on each scale along with known relevant relationships between the fields.
+The specification of these operations is primarily mathematical and numerical in nature, and is influence heavily by the physical model compatibility and geometric interactions.
+
+%\red{cite other coupling models?}
+
+\section{AMSI Model Implementation}\label{sec:amsi-mdl-impl}
+
+\section{Parallel Design}\label{sec:amsi-par-design}
+
+\subsection{AMSI Design}\label{sec:amsi-design}
+
+%\cite{belgacem2017future} (this approach is similar to that taken in \cite{} without requiring the inheritence of/restricution to a particular class interface and the addition of a managing process on a single node in the parallel execution space of the simulation)
+
+The AMSI multiscale library provides the facilities to instantiate, manage, and execute dynamic scale coupling operations on otherwise uncoupled single-scale simulation codes.
+AMSI specifically targets the use-case of a developer taking existing single-scale simulation codes and combining them to produce a new multi-scale code while minimizing intervention in the legacy codebases.
+
+The core of the AMSI multiscale systems operates at the numerical and implementation levels of the ASM and ACM abstraction hierarchy.
+While developers use a top-down approach to specify and construct a simulation, initial development on AMSI multiscale has focused on providing functionality in a bottom-up fashion.
+The intent of this approach is to provide a useful set of functionality for developers before working up the abstraction hierarchy to provide facilities for operations required at those levels.
+
+At the lowest-level, the AMSI multiscale library is concerned with the instantiation and management of the individual single-scale codes used in a multi-scale simulation.
+The parallel structure of HPC machines requires that in order to couple multiple single-scale codes into a multi-scale code, the relationship of those individual single-scale codes to the machine must be modeled by AMSI multiscale.
+
+The set of threads of execution in a given simulation and their relationship with the physical hardware of the machine is called the parallel execution space (roughly equivalent to MPI\_COMM\_WORLD) of the simulation.
+AMSI multiscale models the parallel execution space of the whole multi-scale simulation as well as each single-scale code in the multi-scale system, through the use of a Scale.
+
+\subsection{AMSI Scales}\label{sec:scale}
+
+The Scale is a data structure modeling the relationship between the computational level of a single ASM used in the construction of a multi-scale simulation and the parallel execution space of the HPC machine the simulation is executing on.
+The relationship between the Scale and the parallel execution space is specified by the user using an AMSI multiscale configuration file, further discussed in section \ref{sec:amsi-par-impl}.
+All Scales present in a multi-scale simulation are instantiated and configured at the beginning of a multi-scale simulation.
+
+Several methods for determining the partitioning of the parallel execution space and assigning the partitions to Scales in a simulation are provided.
+Scale allocations may be determined by total process count, process ratios between multiple Scales, and additionaly may take advantage of machine introspection operations such as using the \verb-hwloc- \cite{broquedis2010hwloc} hardware locality library in order to take advantage of machine structure for parallel communication operations.
+
+Also during initialization, a main-like function is associated with an individual Scale.
+A developer may simply reuse the main function of the existing single-scale simulation, however the function is additionally passed an MPI\_COMM associated with the Scale on which parallel operations for the single-scale code modeled by the Scale are to take place.
+
+\subsection{AMSI Couplings}\label{sec:coupling}
+
+The Coupling is a data structure modeling the relationship of units of coupling data between two distinct Scales.
+The Coupling in AMSI multiscale is not assumed to be symmetric so in order to establish a two-way coupling between Scales a Coupling must be produced for each direction of communication.
+While it is often the case that a coupling is bi-directional, this is not always the case \cite{gravemeier2008towards}, so this form is preferred.
+Each Coupling exists on the union of the parallel execution spaces associated with both Scales being coupled, and specifies on Scale as the producer and the other as the consumer.
+
+The Coupling operates at the computational level of the ACM, but is directly informed by specifications at the numerical level.
+Individual units of scale-coupling data are supplied to the Coupling, e.g.
+tensor field values located at points as specified by the domain relationship in the ACM.
+These may be raw values or values derived locally, i.e.
+the tensor field transformation operations may by applied to the coupling data on either end of the coupling.
+This allows for minimization of communication overhead whenever possible.
+
+A unit of coupling data generically refers to any data type as the precise implementation or data structure of the data is not assumed by the Coupling or AMSI multiscale until the communication phase of the coupling operation.
+
+The parallel discretization of the numerical domains of the multi-scale problem informs the parallel locality of both the origin and the destination of each unit of scale-coupling data.
+At present this mapping is taken care of by the developer.
+Facilitating the automation of the construction of this mapping is the next step in advancing the development of AMSI multiscale up the ASM/ACM abstraction hierarchy.
+
+\section{AMSI Parallel Implementation}\label{sec:amsi-par-impl}
+
+\red{cite the in-progress neuron paper as the most up-to-date reference for biotissue}
+
+Software intended to suppport multi-physics simulations needs to "enable the introduction of new models, algorithms, and data structures" \cite{keyes2013multiphysics}.
+Further, a set of well-defined modules with distinct interfaces must be implemented, especially as this concerns data ownership and simulation state.
+State data must be query-able through interfaces in order for coupled codes to maintain consistency throughout the simulation.
+Further, implementing general code, or providing mechanisms to extract important underlying data in general formats is necessary.
+This forces multi-physics application developers to adhere to specific data structures and algorithms, or write their own wrapper code in order to translate specialized data structures into more general formats.
+These issues and rationales for multi-scale software development are further discussed in \cite{keyes2013multiphysics}.
+
+Toward this end AMSI provides a simple API used for multi-scale coupling and scale-sensitive load-balancing operations, and makes no requirement on the application code providing any interfaces for AMSI usage.
+This allows intervention in legacy code only in locations where multi-scale operations are to be implemented, and where replacement of global collective operations with scale-collective operations is required in order to prevent parallel deadlock.
+
+%As AMSI is designed around an abstract simulation model, the implementation is not dependent on a particular API or data structure for the discrete domain representation, tensor fields, and numerical implementation of the physical derivation.
+All computational operations operating on user data are implemented generically to allow the use or arbitrary data types for coupling operations.
+The user may need to provide a description of the datatype to the underlying parallel communication library for data types requiring a deep copy for serialization.
+
+\red{Show a multi-scale simulation initialization in AMSI and how construct and execute a Coupling using code snippets.}
+
+\section{Soft-Tissue Implementation}\label{sec:bio-implementation}
 
 Show how AMSI was used with the existing single-scale simulations to construct the biotissue problem.
+Discuss the tensor fields involved in the coupling and the transformations required to coupled the physical models together (deformation gradient and stress/stress derivatives), the geometric coupling which is known a-priori due to the strongly-separated nature of this multi-scale simulation since it is hierarchical.
 
-\section{Results}\label{sec:results}
+Discuss the control flow of the single-scale macro and micro simulations. Discuss the modifications to the control flow using AMSI to couple the simulations together into a multi-scale simulation.
+
+Include control flow diagrams for both individual scales as well as the combined multi-scale simulation.
+
+\section{AMSI Performance Measures}\label{sec:amsi-metrics}
+
+Discuss that currently scales must operate on non-overlapping partitions in the parallel execution space, requiring that one scale is idle while the other executes.
+Address how to minimize this idle time.
 
@@
 
-\section{Future Work}\label{sec:future-work}
+\section{Development of Additional Multi-scale Simulations}\label{sec:new-simulations}
 
-Brief overview of obstacles being worked on towards concurrent multi-scale.
+\subsection{Hierarchical Simulations}\label{sec:info-passing-examples}
+
+This should be relatively straightforward, just need problem to discuss.
+
+The production of CouplingData objects for hierarchical multi-scale problems is particularly straightforward as the domain relationship is well-established a-priori.
+
+\subsection{Coupled-Domain Simulations}\label{sec:concurrent-examples}
+
+Discussion of obstacles being worked on towards concurrent multi-scale with discussion of at least one particular problem.
+
+The production of a CouplingData object derived from the relationship between the domains in a concurrent partitioned-domain problem is a complex operation currently undergoing development.
 
 \bibliographystyle{siamplain}
-\bibliography{amsi}
+\bibliography{multiscale_theory,multiscale_frameworks,software}
 \end{document}
